### 2014-11-25 tex/short_2015/body.tex
EG short paper initial version

% ---------------------------------------------------------------------
% EG author guidelines plus sample file for EG publication using LaTeX2e input
% D.Fellner, v1.17, Sep 23, 2010

\title[Automatic Proxy Geometry]%
      {Automatic Proxy Geometry}

% for anonymous conference submission please enter your SUBMISSION ID
% instead of the author's name (and leave the affiliation blank) !!
\author[J.\,O. Nygaard \& J. Mikkelsen]
       {J.\,O. Nygaard$^{1}$
        and J. Mikkelsen$^{1}$
        and Husk Anonymisering Hvis Paakrevet!$^{1}$
%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
        \\
% For Computer Graphics Forum: Please use the abbreviation of your first name.
         $^1$SINTEF ICT, applied mathematics, Norway
%        $^2$ Another Department to illustrate the use in papers from authors
%             with different affiliations
       }

% ------------------------------------------------------------------------

% if the Editors-in-Chief have given you the data, you may uncomment
% the following five lines and insert it here
%
% \volume{27}   % the volume in which the issue will be published;
% \issue{1}     % the issue number of the publication
% \pStartPage{1}      % set starting page

%-------------------------------------------------------------------------
\begin{document}

% \teaser{
%  \includegraphics[width=\linewidth]{eg_new}
%  \centering
%   \caption{New EG Logo}
% \label{fig:teaser}
% }

\maketitle

\begin{abstract}
   We describe an algorithm, and implementation, of an automatic proxy geometry
   system for a server-client remote rendering setup. The client may be a thin
   client in the form of a hand-held pad, for our implementation we only assume
   the availability of WebGL. For our rendering server, we make use of a
   GL-capable web server. In situations where the rate of received images
   deteriorate due to low bandwith, high latency or long rendering times, the
   client displays an automatically generated proxy geometry. This proxy
   geometry is computed from depth buffers bundled with rendered images from the
   server.
   % Husk: Skal vaere en blank linje nedenfor

\begin{classification} % according to http://www.acm.org/class/1998/
\CCScat{FIX THIS Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
\end{classification}

\end{abstract}

%-------------------------------------------------------------------------
\section{Introduction}

With increasingly powerful rendering engines, in forms ranging from mobile and
hand-held devices to traditional stationary computers, one might be led to
believe that rendering frame rates could only increase and visualization could
only become more interactive. Alas, hand in hand with this development we also
find ever increasing requirements on computation accuracy, power efficiency,
data sizes and so on.

One result of this is that server-side rendering is as relevant as ever
before. And, with a client-server model for rendering, the problems with
transferring images from the server to the model is of course also still
present.

In order to obtain the frame rates on the client suitable for interactive
visualizations, demands are put on bandwidth, latency and rendering speed. If
one or more of these demands are not met, according to a suitable measure, the
frame rate will deteriorate together with the user experince.

One way to deal with such situations is for the client to have a so-called proxy
model, which is defined as a temporary model to be shown and manipulated locally
on the client while waiting for the appropriate image from the server.

A particularly unwelcomed chore is that of building proxy geometries for various
data sets. The typical ``hello world'' proxy geometry is the familiar bounding
box wire frame, but this will degrade the user experience unacceptably for most
applications. One alternative is to hand-build such simpler proxy
geometries. They have to be simple both for the building cost, and also for the
actual rendering on the possibly very thin client. Another alternative is to
automate this process, a large triangulation may for instance be represented by
a much coarser approximating triangulation that can be built automatically. One
disadvantage of this approach is that this is not easy to do for many kinds of
data.

Our suggested solution is to pass along depth information from the server along
with ordinary rendered images. From this depth image, a rudimentary 3D model can
be built, and using the ordinary image as a texture, this model can be
transformed and rendered on the client while waiting for the next update from
the server. If the server does not take too long to respond, and/or the client
does not change the position or orientation of the model too much, it is to be
hoped that this display of a proxy model integrates seemlessly with the slower
stream of server-rendered frames.

[Kanskje ha inn en foreløpig figur-skisse her, med dybdefelt vist som en
  høydeflate...?]

[Trenger referanser til state of the art o.l. her]

%-------------------------------------------------------------------------

\section{The auto proxy algorithm}

The first main idea is to automatically construct a proxy geometry on the client
side from depth data provided by the server. For a new frame to be displayed on
the client, the server renders an image in a frame buffer, and it is assumed
that it in this process also renders a corresponding depth buffer. This depth
buffer may be regarded as a height map over a viewing plane relative to the
observing client. Hence, this depth map does not contain information about scene
elements not in view for the set of view parameters at the time of server side
rendering. Our approach is based on the assumption that small transformations of
this height map still will give good approximations to the scene being
rendered. These renderings are exactly what the client will compute and show
while waiting for the next update from the server. See
Figure~\ref{fig:2Dheightmap} and Figure~\ref{fig:2DheightmapRoated} for an
illustration of this heightmap computed on the server, and a recovered 3D model
on the client.

\begin{figure}[htb]
  \centering
%  \includegraphics[width=.8\linewidth]{test.jpg}
%  \vspace{-5cm}
  \input{fig2.tex}
  \caption{\label{fig:2Dheightmap}
           The rendered image (of \eg, a cube) as a height map seen from above,
           for a certain transformation.}
\end{figure}

\begin{figure}[htb]
  \centering
  \subfigure[Low frame rate]{
    \input{fig3a.tex}
  }
  \subfigure[Initial proxy model]{
    \input{fig3b.tex}
  }
  \subfigure[Client-generated frame]{
    \input{fig3c.tex}
  }
  \subfigure[With texturing]{
    \input{fig3d.tex}
  }
  \subfigure[Screen-space-sized splats]{
    \input{fig3e.tex}
  }
  \subfigure[With intra-splat depths]{
    \input{fig3f.tex}
  }
  \caption{\label{fig:2DheightmapRotated}
           The depth map and image sent from the server enables the client to
           recover a 3D model approximating the one rendered on the server. This
           client-side model is shown here with a small additional rotation, again
           seen from above.}
\end{figure}

Note that the depth map from the server does not allow the client to recover all
information in the server side model. For instance, the solid surface shown in
Figure~\ref{fig:2Dheightmap} cannot be deduced, hence we show the dots in
Figure~\ref{fig:2DheightmapRotated}. Still, there are several things we can do
to improve the client-side computed approximation of the server-side image, and
these will now be described.

%-------------------------------------------------------------------------
\subsection{The server}

The server keeps a 3D model and accompanying viewing parameters, and will render
an image for delivery to the client either on its own accord or because the
client requests it. As long as the server renders a picture into a framebuffer
with a connected depth buffer, it has a negligible cost for the server to bundle
the resulting depth map with the rendered image upon return of the render
result. In practice, this is easily handled by a few extra lines of computer
code where the image is extracted from its framebuffer and prepared for sending
to the client.

The cost must of course be considered if the operation of packaging images for
sending to the client is significant, which may be the case if, \eg, images must
be compressed or converted to some special format. An example of the latter
might be {\texttt base64}-encoding for converting the image to a purely ASCII
format. Together with the rendered image (which we will call ``rgb image'' to
avoid confusion with the ``depth image'') and the depth image, we will also send
the current view transformations, back to the client.

%-------------------------------------------------------------------------
\subsection{The client}

When the client receieves an rgb image, a depth image and view transformations,
it builds a ``proxy model'' from this. This model can then be rendered directly,
if there is no available image from the server for a new transformation provided
by the user, or it can be combined with other proxy models the client already
has. The latter option will be discussed in
Section~\ref{sec:proxyModelReplacement} below.

%-------------------------------------------------------------------------
%\subsubsection{Proxy model building / Splatting}

As indicated by Figure~\ref{fig:2DheightmapRotated}~(a), the received
height map does not really allow us more than concluding where a set of points
belong to the 3D model, these points giving rise to exactly the depth map from
the server. Given these points, the client may build a model approximating the
actual scene being rendered on the server. Since the information from such a
depth map can only contain the foremost point along any ray from the observer
toward the scene, it is often said to be in 2.5D, as opposed to purely 2D or
fully 3D.

The most simple thing the client can do, is just to transform and render this
set of 3D points with the color sampled from the corresponding location in the
server-rendered image, as illustrated in
Figure~\ref{fig:2DheightmapRotated}~(b).

We may set up the server to send a depth buffer of different resolution than the
actual image, and the client can render at most one 3D point for each depth
fragment available. This may be too much for a thin client, and one may want to
operate with a smaller number of 3D points, and instead render each of them with
a larger number of pixels on the client side. This opens up for problems and
possibilities connected to {\em splatting}, \eg, what size and shape of
geometries to render for each 3D point, how to color them, how to compute depth
fragments and so on.

The most simple solution is to render each 3D point as a fixed geometry, for
instance a 2D sphere or rectangle, with the correspondig color from the image,
see Figure~\ref{fig:2DheightmapRotated}~(c). We will briefly describe some
improvements to this. We also note that other possibilities exist apart
from this ``splat-based'' proxy model, \eg, building and maintaining a 3D
occupancy mesh, a computing a distance field from which iso-surfaces can be
extracted, and so on.

In the following, we will refer to a set of reconstructed 3D points from a depth
map as a {\em proxy model}, and the reconstructed points as {\em splats}.

%-------------------------------------------------------------------------
\subsubsection{Splat sizing}

Note that as soon as a given proxy model is transformed, each splat should
ideally be rendered into a number of client pixels according to the new splats
3D position. A splat transformed into view closer to the observer, should result
in more client fragments, and vice versa.

...

%-------------------------------------------------------------------------
\subsubsection{Texturing}

From the previous section, we note that the splats may occupy more client
fragments than just one, even if the should be one to one correspondence between
server depth fragments and client pixels. This will of course also be the case
if the client's capabilities dictate a realtively small number of splats.

In any case, a splat may result in many client fragments, which necessitates an
``intra-splat'' fragment coloring or texturing. This is not entirely straight
forward, since the client really do not know how a finitely-sized splat should
map to the server's model, having only a discrete set of 3D points on that
model.

A first approximation is for the client to assume that the corresponding part of
the server's model is planar in a region around the given point. If this is the
case, a local 2D texture transformation will provide a good approximation to the
intra-splat texturing to be performed on the client. For the splat centered in
$\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be centered on the client's canvas at
$\pv_{i, j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j}$, the texture coordinate
transformation to be used is,
\[
  T = \ldots,
\]
such that the client fragment $\pv_{i, j} + \delta_{i, j}$ gets texel
\[
  I_{p, q},
\]
with $(p, q) = \ldots$.

When the assumption does not hold, \eg, if the splats are very large, or they
originates from a non-planar part of the server's model, this may look rather
odd, see Figure~\ref{fig:LargeSplatsOnCorners}.

\begin{figure}[htb]
  \centering
  \framebox{\includegraphics[width=.8\linewidth]{test.jpg}}
  \caption{\label{fig:LargeSplatsOnCorners}
           The rendered image (of \eg, a cube) as a height map seen from above,
           for a certain transformation.}
\end{figure}

%-------------------------------------------------------------------------
\subsubsection{Splat depth fragments}

%-------------------------------------------------------------------------

%------------------------------------------------------------------------
\subsection{Splat set replacement algorithms}
\label{sec:proxyModelReplacement}

In the case of splat sets: Replacement algorithms

%-------------------------------------------------------------------------
\subsection{Dynamic/heuristic method selection}

%-------------------------------------------------------------------------
\section{Results and discussion}

%-------------------------------------------------------------------------
\section{Conclusion and future work}

%-------------------------------------------------------------------------
\subsection{References}

List all bibliographical references in 9-point Times, single-spaced, at the
end of your paper in alphabetical order. When referenced in the text, enclose
the citation index in square brackets, for example~\cite{Lous90}. Where
appropriate, include the name(s) of editors of referenced books.

For your references please use the following algorithm:
\begin{itemize} 
\item \textbf{one} author: first 3 chars plus year -- 
      e.g.\ \cite{Lous90}
\item \textbf{two}, \textbf{three} or \textbf{four} authors: first char
      of each family name plus year --  e.g.\ \cite{Fellner-Helmberg93} 
      or \cite{Kobbelt97-USHDR} or \cite{Lafortune97-NARF}
\item \textbf{more than 4} authors: first char of family name from 
      first 3 authors followed by a '*' followed by the year -- 
      e.g.\ \cite{Buhmann:1998:DCQ} or \cite{FolDamFeiHug.etal93} 
\end{itemize}

For BibTeX users a style file \ \texttt{eg-alpha.bst} \ is available which
uses the above algorithm.

%-------------------------------------------------------------------------
\subsection{Conclusions}

Please direct any questions to the production editor in charge of
these proceedings.

%-------------------------------------------------------------------------

%\bibliographystyle{eg-alpha}
\bibliographystyle{eg-alpha-doi}

\bibliography{egbibsample}

%-------------------------------------------------------------------------

\end{document}

### 2014-11-25 include/tinia/qtcontroller/impl/ServerThread.hpp
Fixed bugs in the protocol buffer creation:
- Data in Snapshot<...>Fetcher are written in destructor, hence all fetcher usage need to be in empty scope.
- Now operates on a referanse to a QByteString buffer, so that we get the protocol buffer to come out again from the function calls.
- Added a lot of debugging text. This should be cleaned later.

@@ -35,7 +35,7 @@ private:
      *  contain depth and transformation data (optional proto buffer fields).
      *  getSnapshotBundleTxt() seems not to be used, and we therefore do not need a replica of it.
      */
-    void getSnapshotBytes( QByteArray* protoBytes, const QString &request,
+    void getSnapshotBytes( QByteArray* &protoBytes, const QString &request,
                            tinia::jobcontroller::Job* job,
                            tinia::qtcontroller::impl::OpenGLServerGrabber* grabber,
                            const bool with_depth );
@@ -55,7 +55,7 @@ private:
     /** Handles non-static content where we want to sent raw bytes of images, if applicable.
      * @returns true if the file is non-static, false otherwise.
      */
-    bool handleNonStatic(QByteArray *protoBytes, QTextStream& os, const QString& file,
+    bool handleNonStatic(QByteArray* &protoBytes, QTextStream& os, const QString& file,
                          const QString& request);
 
     void updateState(const QString& request);

### 2014-11-25 src/qtcontroller/ServerThread.cpp

@@ -196,6 +196,7 @@ public:
         //m_gl_grabber_locker.unlock();
 
         QByteArray str( qBuffer.data(), int(qBuffer.size()) );
+        std::cout << "SnapshotAsBytesFetcher::~SnapshotAsBytesFetcher - returning " << qBuffer.size() << " bytes." << std::endl;
         m_reply << str;
     }
 
@@ -284,11 +285,13 @@ void ServerThread::run()
 
             QString requestURI = getRequestURI(request); // Should return a filename with .txt or .xml
             if (requestURI.split('.').last() == "txt") {
+                std::cout << "\n\nServerThread::run() - Starting on txt-request" << std::endl;
 
                 // Still need a QTextStream in case handleNonStatic throws an error, but it should not be connected on socket yet!
                 QString errorString;
                 QTextStream textStream(&errorString);
                 if (!handleNonStatic(responseBuffer, textStream, requestURI, request)) {
+                    std::cout << "ServerThread::run() - Not able to handleNonStatic..." << std::endl;
                     //QTextStream os(&socket);
                     textStream.setAutoDetectUnicode(true);
                     textStream << getStaticContent(getRequestURI(request)) << "\r\n";
@@ -296,14 +299,25 @@ void ServerThread::run()
                     os << errorString;
                 }
                 else {
+                    std::cout << "ServerThread::run() - Handled nonStatic" << std::endl;
+                    std::cout << "ServerThread::run() - responseBuffer is at addresse " << responseBuffer << std::endl;
+                    std::cout << "ServerThread::run() - responseBuffer is size " << responseBuffer->size() << std::endl;
+
                     QDataStream ds(&socket);
                     ds << httpHeader(getMimeType("file.bin"), 200, QString("utf-8"), (unsigned int)(responseBuffer->size())).toLocal8Bit();
+                    //std::cerr << httpHeader(getMimeType("file.bin"), 200, QString("utf-8"), (unsigned int)(responseBuffer->size())); //.toLocal8Bit();
                     ds << QString("\r\n").toLocal8Bit();
+                    //ds << QString("\r\n").toAscii();
+                    //std::cerr << QString("\r\n");
                     ds << responseBuffer;
+                    //std::cerr << "trying to dump response buffer" << std::endl;
+                    //std::cerr << responseBuffer;
[…]

### 2014-11-26 js/gui/Canvas.js
Unsuccedeed java script side support for protocol buffers.
Untested support for protocol buffers is added in the java script at the subscription of "updateSendPartialComplete" where we expected the binary data to come as params.response. The binary blob containing the protobuf never entered this piece of code, even though it is requested successfully from the browser.

I have not been able to find out where our protobuf blob flows through the java script, and we now decide to tackle the problem from a different angle. This is therefore likely to be the last commit on this branch, and the branc

@@ -19,6 +19,9 @@
 dojo.provide("gui.Canvas");
 
 dojo.require("3rdparty.webgl-utils");
+//dojo.require("3rdparty.ByteBuffer");
+//dojo.require("3rdparty.Long\.min");
+//dojo.require("3rdparty.ProtoBuf\.min");
 dojo.require("renderlist.RenderList");
 dojo.require("renderlist.RenderListParser" );
 dojo.require("renderlist.RenderListRenderer" );
@@ -194,7 +197,8 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
         // We need to fetch a new image.
         dojo.subscribe("/model/updateParsed", dojo.hitch(this, function (params) {
             if (!this._imageLoading) {
-                console.log("Getting new image");
+                console.log("Getting new image - UpdateParsed");
+                console.log(params);
                 dojo.xhrGet({ // Here we explicitly ask for a new image in a new HTTP connection.
                     url: this._urlHandler.getURL(),
                     preventCache: true,
@@ -207,6 +211,7 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                     })
                 });
             }
+            console.log("Got new image?");
         }));
 
         // We are initiating a new update TO the server.
@@ -220,6 +225,11 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
         // The "partial" bit says that we _may_ have another update to send after this as well.
         // Either way, we should show the image we just got from the server (it's newer than the one we have!).
         dojo.subscribe("/model/updateSendPartialComplete", dojo.hitch(this, function (params) {
+
+            if ( params.ioArgs.args.contentType.match(/octet/) ) {
+                log("--------- OCTET STREAM!!! --------------             <-- <-- <--");
+            }
+
             if (params.response.match(/\"rgb\"\:/)) { // For the time being, we assume this to be an image.
 //                console.log("/model/updateSendPartialComplete: response = " + params.response);
                 var response_obj = eval( '(' + params.response + ')' );
@@ -230,12 +240,40 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                 }
             } else {
                 console.log("This was not a snapshot. Why are we here at all?");
+                if (!params.response.match(/html/)) {
+                    console.log("Because we recieved a protocal buffer!");
+                    console.log(params);
+                    //console.log(params.response);
+                    console.log("params.ioArgs.handleAs: " + params.ioArgs.handleAs);
[…]

### 2014-11-26 js/gui/ProxyModel.js

@@ -98,5 +98,71 @@ dojo.declare("gui.ProxyModel", null, {
 
     },
 
+    setAllFromBytes: function(depthBufferAsBytes, imageAsBytes, viewMatAsText, projMatAsText) {
+        console.log("Beginning of setAllFromBytes in ProxyModel.js");
+        if ( this.state == 1 ) {
+            throw "Trying to set data for a proxy model being processed!";
+        }
+
+        if ( (!depthBufferAsBytes) || (!viewMatAsText) || (!projMatAsText) ) {
+            // then we have probably been passed a snapshot made when autoProxy was disabled. We cannot and should not use this, so we just return.
+            return;
+        }
+
+        this.state = 1;
+        var imagesLoaded = 0;
+        var depth_t0 = 0;
+        var rgb_t0 = 0;
+
+        var depthImage = new Image();
+        depthImage.onload = dojo.hitch(this, function() {
+            this._gl.bindTexture(this._gl.TEXTURE_2D, this.depthTexture);
+            this._gl.texImage2D(this._gl.TEXTURE_2D, 0, this._gl.RGB, this._gl.RGB, this._gl.UNSIGNED_BYTE, depthImage);
+            this._gl.texParameteri(this._gl.TEXTURE_2D, this._gl.TEXTURE_MAG_FILTER, this._gl.NEAREST);
+            this._gl.texParameteri(this._gl.TEXTURE_2D, this._gl.TEXTURE_MIN_FILTER, this._gl.NEAREST);
+            this._gl.bindTexture(this._gl.TEXTURE_2D, null);
+//            console.log("Depth image (" + depthBufferAsText.length + " bytes) loaded in " + (Date.now()-depth_t0) + " ms. (" + Math.floor((depthBufferAsText.length/(Date.now()-depth_t0))) + " bytes/ms)");
+            imagesLoaded = imagesLoaded + 1;
+            if (imagesLoaded==2) {
+                this.projection         = projMatAsText.split(/ /);
+                this.projection_inverse = mat4.inverse(mat4.create( this.projection ));
+                this.from_world         = viewMatAsText.split(/ /);
+                this.to_world           = mat4.inverse(mat4.create( this.from_world ));
+                this.dir = vec3.create( [-this.to_world[8], -this.to_world[9], -this.to_world[10]] );
+                vec3.normalize(this.dir); // Should probably already be normalized...
[…]

### 2014-11-26 js/gui/ProxyRenderer.js

@@ -418,6 +418,23 @@ dojo.declare("gui.ProxyRenderer", null, {
         // console.log("setDepthData: done");
     },
 
+    setDepthDataRaw: function(imageAsBytes, depthBufferAsBytes, viewMatAsText, projMatAsText) {
+        console.log("setDepthDataRaw: Starting load proceess for images-as-bytes data");
+        if (this._proxyModelBeingProcessed.state!=0) {
+            console.log("A depth buffer is already being processed, discarding the new one just received! (state=" + this._proxyModelBeingProcessed.state + ")");
+        } else {
+            if (!this._lock2) {
+                // lock2 not set, we shall update the ring buffer as usual
+                this._proxyModelBeingProcessed.setAllFromBytes(depthBufferAsBytes, imageAsBytes, viewMatAsText, projMatAsText);
+            }
+            if (this._lock) {
+                // All models have been reset and we just inserted a proxy model in the ring buffer, after this, we shall not insert any more, hence we set lock2.
+                this._lock2 = true;
+            }
+        }
+        console.log("setDepthDataRaw: End of load proceess for images-as-bytes data");
+
+    },
 
     // Some useful wrappers
     _setUniform1i: function(prog, name, value) {

### 2014-11-26 js/model/ExposedModelSender.js

@@ -104,6 +104,7 @@ dojo.declare("model.ExposedModelSender", null, {
             
             // Reset keys we have seen.
             this._keys = {};
+            console.log("ExposedModelSender.js - updateComplete - Does this show?");
             dojo.publish("/model/updateSendComplete", [{"response": response, "ioArgs" : ioArgs}]);
         }
     },

### 2014-11-27 examples/cuberenderer/CubeJob.cpp
Sets up Freeglut, GLdebugMessages and calls into cubeJob :)

@@ -88,13 +88,15 @@ bool CubeJob::renderFrame(const std::string &session, const std::string &key, un
     glClear(GL_COLOR_BUFFER_BIT|GL_DEPTH_BUFFER_BIT);
 
     glViewport(0, 0, width, height);
-    tinia::model::Viewer viewer;
-    m_model->getElementValue( key, viewer);
+   // tinia::model::Viewer viewer;
+    //m_model->getElementValue( key, viewer);
     glLoadIdentity();
     glMatrixMode(GL_PROJECTION);
-    glLoadMatrixf( viewer.projectionMatrix.data() );
+    glLoadIdentity();
+    //glLoadMatrixf( viewer.projectionMatrix.data() );
     glMatrixMode(GL_MODELVIEW);
-    glLoadMatrixf( viewer.modelviewMatrix.data() );
+    //glLoadMatrixf( viewer.modelviewMatrix.data() );
+    glLoadIdentity();
 
     glBegin(GL_POLYGON);
     glColor3f(   1.0,  0.0, 0.0 );

### 2014-11-27 include/tinia/trell/FRVGLJobController.hpp

@@ -67,20 +67,19 @@ private:
     jobcontroller::OpenGLJob*                           m_openGLJob;
 
     int                                                 m_quality;
-    struct RenderEnvironment {
+    
         GLuint                                          m_fbo;
         GLuint                                          m_renderbuffer_rgba;
         GLuint                                          m_renderbuffer_depth;
         GLsizei                                         m_width;
         GLsizei                                         m_height;
         GLsizei                                         m_samples;
-    };
-
-    std::list<RenderEnvironment*>                       m_environments;
+    
 
     GLsizei                                             m_max_samples;
 
-
+    int                                                 m_argc;
+    char**                                               m_argv;
     
     bool
     checkFramebufferCompleteness() const;

### 2014-11-27 include/tinia/trell/GLDebugMessages.hpp

@@ -1,0 +1,190 @@
+/************************************************************************
+ *
+ *  File: GLDebugMessages.hpp
+ *
+ *  Created: 5. November 2012
+ *
+ *  Authors: Erik W. Bjønnes <erik.bjonnes@sintef.no>
+ *
+ *  Copyright (C) 2012 by SINTEF.  All rights reserved.
+ *
+ *  SINTEF, Pb 124 Blindern, N-0314 Oslo, Norway
+ *  http://www.sintef.no
+ *********************************************************************/
+#pragma once
+
+#include <gl/glew.h>
+#include <iostream>
+#include <ostream>
+#include <string>
+
+#ifndef APIENTRY
+#define APIENTRY
+#define TINIA_APIENTRY
+#endif
+
+
+/** Helper header for setting up OpenGL debug messages.
+ *  
+ * If GL version does not support KHR_DEBUG will try with older ARB extension.
+ * If neither is present, will give a lot of compiler errors!
+ */
+namespace tinia { namespace trell { namespace GLDebugMessages{
+
+namespace impl {
+#ifdef WIN32
+#include "Windows.h"
+#endif
+
+std::string getTypeString( GLenum type )
+{
+  switch(type) {
+  case GL_DEBUG_TYPE_ERROR: 
+    return("Error");
+  case GL_DEBUG_TYPE_DEPRECATED_BEHAVIOR:
+    return("Deprecated Behaviour");
+  case GL_DEBUG_TYPE_UNDEFINED_BEHAVIOR:
+    return("Undefined Behaviour");
+  case GL_DEBUG_TYPE_PORTABILITY:
+    return("Portability Issue");
+  case GL_DEBUG_TYPE_PERFORMANCE:
+    return("Performance Issue");
+  case GL_DEBUG_TYPE_OTHER:
+    return("Other");
+  default:
+    return("");
+  }
+}
+
+std::string getSeverityString( GLenum severity )
+{
+  switch(severity) {
+  case GL_DEBUG_SEVERITY_HIGH: 
+    return("High severity");
+  case GL_DEBUG_SEVERITY_MEDIUM:
+    return("Medium severity");
+  case GL_DEBUG_SEVERITY_LOW:
+    return("Low severity");
+  case GL_DEBUG_SEVERITY_NOTIFICATION:
+	  return("Notification");
+  default:
+    return("Unknown severity");
+  }
+}
+
+
[…]

### 2014-11-27 src/trell/FRVGLJobController.cpp

@@ -20,72 +20,12 @@
 #include <sstream>
 #include <tinia/renderlist/XMLWriter.hpp>
 #include "tinia/trell/FRVGLJobController.hpp"
-
-#include <GL/glut.h>
+#include "tinia/trell/GLDebugMessages.hpp"
+#include <GL/freeglut.h>
 
 namespace {
 
 static const std::string package = "FRVGLJobController";
-
-struct GLDebugLogWrapperData
-{
-    void  (*m_logger_callback)( void* logger_data, int level, const char* who, const char* msg, ... );
-    void*   m_logger_data;
-};
-
-
-#ifdef GLEW_khr_DEBUG // make sure the glew version is new enough
-static
-void
-GLDebugLogWrapper( GLenum source,
-                   GLenum type,
-                   GLuint id,
-                   GLenum severity,
-                   GLsizei length,
-                   const GLchar* message,
-                   void* data )
-{
-
-    const char* source_str = "???";
-    switch( source ) {
-    case GL_DEBUG_SOURCE_API:             source_str = "api"; break;
-    case GL_DEBUG_SOURCE_WINDOW_SYSTEM:   source_str = "wsy"; break;
-    case GL_DEBUG_SOURCE_SHADER_COMPILER: source_str = "cmp"; break;
-    case GL_DEBUG_SOURCE_THIRD_PARTY:     source_str = "3py"; break;
-    case GL_DEBUG_SOURCE_APPLICATION:     source_str = "app"; break;
-    case GL_DEBUG_SOURCE_OTHER:           source_str = "oth"; break;
-    default: break;
-    }
-
-    const char* type_str = "???";
-    switch( type ) {
-    case GL_DEBUG_TYPE_ERROR:               type_str = "error"; break;
-    case GL_DEBUG_TYPE_DEPRECATED_BEHAVIOR: type_str = "deprecated"; break;
-    case GL_DEBUG_TYPE_UNDEFINED_BEHAVIOR:  type_str = "undef"; break;
-    case GL_DEBUG_TYPE_PORTABILITY:         type_str = "portability"; break;
-    case GL_DEBUG_TYPE_PERFORMANCE:         type_str = "performance"; break;
-    case GL_DEBUG_TYPE_OTHER:               type_str = "other"; break;
-    default: break;
-    }
-
-    int level = 2;
-    switch( severity ) {
[…]

### 2014-11-27 tex/short_2015/body.tex
tex additions and comment in autoProxy.vs

@@
 
+\vspace{10cm}
+
 A first approximation is for the client to assume that the corresponding part of
@@
 $\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be centered on the client's canvas at
-$\pv_{i, j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j}$, the texture coordinate
-transformation to be used is,
+screen coordinate $\pv_{i, j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i,
+j}$, the texture coordinate transformation to be used is,
 \[
-  T = \ldots,
+  T =
+  \begin{pmatrix} 1/\text{splats}_x & 0 \\ 0 & 1/\text{splats}_y \end{pmatrix}
+  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
+  \begin{pmatrix} \text{vpwidth}/\text{splats}_x & 0 \\ 0 & \text{vpheight}/\text{splats}_y \end{pmatrix}
+  \cdot \text{overlap},
 \]
+where $(\ev_1 \, \, \, \ev_2)$ maps the ``client's splat region'' (in $[0,
+1]^2$) to the corresponding texture area, $(\sv_{dx} \, \, \, \sv_{dy})^{-1}$
+maps the ``client's screen space splat area'' to $[0, 1]^2$, and $\text{splatSize}_x
+= \text{vp}_\text{width}/\text{splats}_x$ and $\text{splatSize}_y
+= \text{vp}_\text{height}/\text{splats}_y$ together with the ``overlap'' just
+provides a scaling factor so that ``overlap=1'' produces non-overlapping but
+covering splats when $M_c=M_s$, \ie, the client does no transformation of the
+proxy model.
+
+We compute $\sv_{dx}$ and $\sv_{dy}$ by sampling proxy model positions and
+performing the perspective division and transformation into window coordinates
+in the vertex shader.
+Let
+\[
+  \text{pos} =
+  U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
+  \text{pos}_{\Delta x} =
+  U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
+  \text{and}\, \, \, 
+  \text{pos}_{\Delta y} =
+  U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
+\]
+where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
+2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
+y+\Delta y) - 1$ are depths sampled from the received depth buffer and transformed
+to $[-1, 1]$.
+
+We use $\Delta = \Delta x = \Delta y = ...$, so that ...
+
+Then, for the 2D vectors spanning the splats in client window coordinates, we
+get
+\[
+  \sv_{\Delta x} =
+    \frac{\text{viewport width}}{2\Delta x} \left(
+        \frac{\text{pos}_{\Delta x}.xy}{\text{pos}_{\Delta x}.w} -
+        \frac{\text{pos}.xy}{\text{pos}.w}
+    \right),
+\]
+and
+\[
+  \sv_{\Delta y} =
+    \frac{\text{viewport width}}{2\Delta y} \left(
+        \frac{\text{pos}_{\Delta y}.xy}{\text{pos}_{\Delta y}.w} -
+        \frac{\text{pos}.xy}{\text{pos}.w}
+    \right),
+\]
+where we have used the ``shader notation'' to extract 2D vectors and scalars
+from the 4D vectors.
+
+Then, ...
+
 such that the client fragment $\pv_{i, j} + \delta_{i, j}$ gets texel

### 2014-11-27 examples/jpg_proxy_test/JPCJob.cpp
Testing depth buffer resolution deviating from rgb image resolution

@@ -35,19 +35,8 @@
 #include "tinia/model/GUILayout.hpp"
 #include <iostream>
 #include "tinia/model/File.hpp"
+#include "tinia/utils/ProxyDebugGUI.hpp"
 
-
-namespace {
-
-    // NB! This list must match the one in ProxyRenderer.js, exactly!
-    const char *allowed_auto_proxy_algos[] = { "0) AngleCoverage-5",
-                                               "1) AngleCoverage-2",
-                                               "2) OnlyMostRecent",
-                                               "3) ReplaceOldestWhenDifferent-5",
-                                               "4) ReplaceOldest-5",
-                                               NULL };
-
-}
 
 
 namespace tinia {
@@ -63,142 +52,26 @@ bool JPCJob::init()
     m_model->addElement("viewer", viewer);
     m_model->addElement<std::string>( "boundingbox", "-2.0 -2.0 -2.0 2.0 2.0 2.0" );
 
-    // Adding variables to the model
-    // Note that these values are not communicated to the ProxyRenderer until they are actually changed, due to the use
-    // of listeners. (Should maybe fix this, by some initialization routine.)
+    // The constructor of proxyGUI sets up "magical" elements in m_model.
+    tinia::utils::ProxyDebugGUI proxyGUI( m_model,
+                                          true, // autoProxy
+                                          true, // autoProxy-debugging
+                                          true, // jpgProxy
+                                          true,  // autoSelect proxy method
+                                          true  // depth buffer manipulation
+                                          );
+
+    // And this method returns a corresponding grid containing GUI elements for manipulating these elements.
+    tinia::model::gui::Grid *grid = proxyGUI.getGrid();
+
+    // Setting up root consisting of canvas + grid
+    tinia::model::gui::HorizontalLayout *rootLayout = new tinia::model::gui::HorizontalLayout();
     {
-        m_model->addElement<bool>( "ap_autoSelect", false );            // Selects whatever proxy method works fastest.
-        m_model->addElement<bool>( "ap_autoSelectSampleAll", false );
[…]

### 2014-11-27 include/tinia/utils/ProxyDebugGUI.hpp

@@ -41,12 +41,13 @@ class ProxyDebugGUI //  : public Item
 public:
 
     ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
-                   const bool with_ap, const bool with_ap_debugging, const bool with_jpg, const bool with_auto_select );
+                   const bool with_ap, const bool with_ap_debugging, const bool with_jpg, const bool with_auto_select,
+                   const bool with_depth_buffer_manipulation = false );
     tinia::model::gui::Grid *getGrid();
     
 private:
 
-    bool m_w_ap, m_w_apd, m_w_jpg, m_w_as;
+    bool m_w_ap, m_w_apd, m_w_jpg, m_w_as, m_with_depth_buffer_manipulation;
 
 };
 

### 2014-11-27 js/gui/Canvas.js

@@ -88,6 +88,9 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                 "jpeg_quality": 50
             });
         }
+        this._urlHandler.updateParams({
+            "depth_w": this._width, "depth_h": this._height
+        });
 
         // This is not in use
         // this._onLoadFunction = dojo.hitch(this, this._loadComplete);
@@ -332,6 +335,12 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
         }) );
         this._modelLib.addLocalListener( "ap_jpgQuality", dojo.hitch(this, function(event) {
             this._urlHandler.updateParams( { "jpeg_quality": this._modelLib.getElementValue("ap_jpgQuality") } );
+        }) );
+        this._modelLib.addLocalListener( "ap_depthWidth", dojo.hitch(this, function(event) {
+            this._urlHandler.updateParams( { "depth_w": this._modelLib.getElementValue("ap_depthWidth") } );
+        }) );
+        this._modelLib.addLocalListener( "ap_depthHeight", dojo.hitch(this, function(event) {
+            this._urlHandler.updateParams( { "depth_h": this._modelLib.getElementValue("ap_depthHeight") } );
         }) );
 
         this._modelLib.addLocalListener( "ap_autoSelectSampleAll", dojo.hitch(this, function(event) {

### 2014-11-27 src/qtcontroller/ServerThread.cpp

@@ -73,7 +73,8 @@ public:
                                     tinia::qtcontroller::impl::OpenGLServerGrabber* gl_grabber,
                                     const bool getRBGsnapshot,
                                     const bool pngMode, // Could have used TrellRequest, but then we would have to drag in mod_trell.h...
-                                    const int jpg_quality )
+                                    const int jpg_quality,
+                                    const unsigned depth_w = 0, const unsigned depth_h = 0 ) // Default values will result in canvas size being used.
         : m_reply( reply ),
           m_request( request ),
           m_job( NULL ),
@@ -81,7 +82,10 @@ public:
           m_gl_grabber_locker( gl_grabber->exclusiveAccessMutex() ),
           m_getRGBsnapshot( getRBGsnapshot ),
           m_pngMode( pngMode ),
-          m_jpg_quality( jpg_quality )
+          m_jpg_quality( jpg_quality ), // Could get this (and other) parameter from parsing below, like is done for m_ket etc. Same thing is done by the caller.
+                                        // (Better done by the caller, now that we (may) have multiple viewers...)
+          m_depth_w( depth_w ),
+          m_depth_h( depth_h )
     {
         using namespace tinia::qtcontroller::impl;
         
@@ -95,6 +99,12 @@ public:
                                                  "width height key" );
         m_width  = arguments.get<0>();
         m_height = arguments.get<1>();
+        if (depth_w==0) {
+            m_depth_w = m_width;
+        }
+        if (depth_h==0) {
+            m_depth_h = m_height;
+        }
         m_key    = ( proper_key_to_use != "" ? proper_key_to_use : arguments.get<2>() );
     }
     
@@ -112,6 +122,9 @@ public:
                                       0, -1,
                                       0, m_height);
         img = img.transformed(flipTransformation);
+        if ( (m_depth_w!=m_width) || (m_depth_h!=m_height) ) {
+            img = img.scaled(m_depth_w, m_depth_h);
+        }
         QBuffer qBuffer;
         if (m_pngMode) {
             img.save(&qBuffer, "png");
@@ -145,6 +158,7 @@ protected:
     bool                                            m_getRGBsnapshot;
     bool                                            m_pngMode;
     int                                             m_jpg_quality;
+    unsigned                                        m_depth_w, m_depth_h;
 };
 
 
[…]

### 2014-11-27 src/utils/ProxyDebugGUI.cpp

@@ -46,8 +46,10 @@ namespace utils {
 
 
 ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
-                              const bool with_ap, const bool with_ap_debugging, const bool with_jpg, const bool with_auto_select )
-    : m_w_ap(with_ap), m_w_apd(with_ap_debugging), m_w_jpg(with_jpg), m_w_as(with_auto_select)
+                              const bool with_ap, const bool with_ap_debugging, const bool with_jpg, const bool with_auto_select,
+                              const bool with_depth_buffer_manipulation /* = false */ )
+    : m_w_ap(with_ap), m_w_apd(with_ap_debugging), m_w_jpg(with_jpg), m_w_as(with_auto_select),
+      m_with_depth_buffer_manipulation(with_depth_buffer_manipulation)
 {
     if ( m_w_ap && m_w_jpg && m_w_as ) {
         model->addElement<bool>( "ap_autoSelect", false );            // Selects whatever proxy method works fastest.
@@ -110,6 +112,13 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addElement( "ap_fragExtStatus", "---" );
         model->addElement( "ap_consoleLog", "---" );
         model->addElement<int>( "ap_cntr", 0 );
+    }
+
+    if (with_depth_buffer_manipulation) {
+        model->addConstrainedElement<int>("ap_depthWidth", 512, 4, 1024);
+        model->addAnnotation("ap_depthWidth", "Depth buffer Width)");
+        model->addConstrainedElement<int>("ap_depthHeight", 512, 4, 1024);
+        model->addAnnotation("ap_depthHeight", "Depth buffer height)");
     }
 }
 
@@ -199,6 +208,17 @@ tinia::model::gui::Grid *ProxyDebugGUI::getGrid()
         mainGrid->setChild(row, 1, new tinia::model::gui::Label("ap_consoleLog", true));
         row++;
         mainGrid->setChild(row, 0, new tinia::model::gui::VerticalExpandingSpace());
+    }
+
+    if (m_with_depth_buffer_manipulation) {
+        mainGrid->setChild(row, 0, new tinia::model::gui::HorizontalSlider("ap_depthWidth"));
+        mainGrid->setChild(row, 1, new tinia::model::gui::Label("ap_depthWidth", false));
[…]

### 2014-11-28 examples/cuberenderer/CubeJob.cpp
CubeJob now has a rotate function that can be used instead of exposed model.
Not ideal tying the library so close to the application, but since this is just to a proof of concept...

@@ -39,6 +39,8 @@
 namespace tinia {
 namespace example {
 CubeJob::CubeJob()
+    :
+    m_rotation(0.0f)
 {
 }
 
@@ -97,6 +99,7 @@ bool CubeJob::renderFrame(const std::string &session, const std::string &key, un
     glMatrixMode(GL_MODELVIEW);
     //glLoadMatrixf( viewer.modelviewMatrix.data() );
     glLoadIdentity();
+    glRotatef( m_rotation * 57.2957795, 0.001, 0.0, 1.0 );
 
     glBegin(GL_POLYGON);
     glColor3f(   1.0,  0.0, 0.0 );
@@ -155,5 +158,13 @@ CubeJob::getRenderList( const std::string& session, const std::string& key )
 {
     return &m_renderlist_db;
 }
+
+float CubeJob::rotate( float r )
+{
+    float f = m_rotation;
+    m_rotation += r;
+    return f;
+}
+
 }
 }

### 2014-11-28 examples/cuberenderer/CubeJob.hpp

@@ -43,10 +43,13 @@ public:
     const tinia::renderlist::DataBase*
     getRenderList(const std::string &session, const std::string &key);
 
+    float rotate( float r );
+
 private:
     tinia::renderlist::DataBase     m_renderlist_db;
     GLuint                      m_gpgpu_quad_vertex_array;
     GLuint                      m_gpgpu_quad_buffer;
+    float                       m_rotation;
 };
 }
 }

### 2014-11-28 include/tinia/trell/FRVGLJobController.hpp

@@ -56,6 +56,8 @@ protected:
     void
     cleanup();
 
+
+
 bool
 FRVGLJobController::onGetSnapshot( char*               buffer,
                                    const size_t        width,

### 2014-11-28 src/trell/FRVGLJobController.cpp

@@ -23,6 +23,10 @@
 #include "tinia/trell/GLDebugMessages.hpp"
 #include <GL/freeglut.h>
 
+
+//not the way it should be, but for proof of concept and bypassing exposedModel, here we go
+#include "../../examples/cuberenderer/CubeJob.hpp"
+
 namespace {
 
 static const std::string package = "FRVGLJobController";
@@ -313,16 +317,28 @@ int FRVGLJobController::run( int argc, char** argv )
     m_argv = argv;
     init();
     m_openGLJob->initGL();
-    //call init
-    //call initGL
     //setup polling of websocket messages
     //ie update/idle function
-    //set job->renderFrame as glutRenderFunction
+    
     //call glutMainloop.
+
     m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
     glutSwapBuffers();
+    
+    float rotation = 0.001, prev = 0;
+    tinia::example::CubeJob* cj = (tinia::example::CubeJob*) m_openGLJob;
+    for (int i = 0; i < 150; i++ )
+    {
+        prev = cj->rotate( rotation * i );
+
+        m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
+        glutSwapBuffers();
+    }
+
     return 1;
 }
+
+
 
 } // of namespace trell
 } // of namespace tinia

### 2014-11-28 include/tinia/trell/FRVGLJobController.hpp
Splitting up into FRVQtController and GLController to cope with Qt MOC'ing.

@@ -20,16 +20,20 @@
 #include <GL/glew.h>
 #include <unordered_map>
 
+
 #include "tinia/jobcontroller/Controller.hpp"
 #include "tinia/jobcontroller/OpenGLJob.hpp"
+
+
 namespace tinia {
 namespace trell {
 
 class FRVGLJobController : public jobcontroller::Controller
 {
+
 public:
 
-    FRVGLJobController( bool is_master = false );
+    explicit FRVGLJobController( bool is_master = false );
     void setJob( jobcontroller::Job *job);
 
    ////from Job.hpp
@@ -43,7 +47,7 @@ public:
     void finish();
     void fail();
     void addScript( const std::string& script) {}
-    //fiks dette, du maa legge til FRVGLJobController til et lib....
+    
 protected:
     /** \copydoc MessageBox::init */
     virtual
@@ -66,18 +70,15 @@ FRVGLJobController::onGetSnapshot( char*               buffer,
                                    const std::string&  key );
 
 private:
+
     jobcontroller::OpenGLJob*                           m_openGLJob;
-
-    int                                                 m_quality;
-    
-        GLuint                                          m_fbo;
-        GLuint                                          m_renderbuffer_rgba;
-        GLuint                                          m_renderbuffer_depth;
-        GLsizei                                         m_width;
-        GLsizei                                         m_height;
-        GLsizei                                         m_samples;
-    
-
+    int                                                 m_quality;    
+    GLuint                                              m_fbo;
+    GLuint                                              m_renderbuffer_rgba;
+    GLuint                                              m_renderbuffer_depth;
+    GLsizei                                             m_width;
+    GLsizei                                             m_height;
+    GLsizei                                             m_samples;
     GLsizei                                             m_max_samples;
 
     int                                                 m_argc;

### 2014-11-28 include/tinia/trell/FRVQtController.hpp

@@ -1,0 +1,34 @@
+#pragma once
+
+#include <QObject>
+
+QT_FORWARD_DECLARE_CLASS(QWebSocket);
+
+namespace tinia{ namespace trell {
+
+class FRVGLJobController;
+
+
+class FRVQtController : public QObject
+{
+    Q_OBJECT
+public:
+    //FRVQtController();
+    explicit FRVQtController( tinia::trell::FRVGLJobController* glJob, QObject* parent = Q_NULLPTR );
+    ~FRVQtController();
+
+Q_SIGNALS:
+    void closed();
+    
+private Q_SLOTS:
+    void onConnect();
+    void onMessageReceived();
+
+private:
+    QWebSocket* m_socket;
+    FRVGLJobController* m_glJob;
+    
+};
+
+
+}} //end tinia::trell

### 2014-11-28 src/trell/FRVQtController.cpp

@@ -1,0 +1,35 @@
+#include "tinia/trell/FRVQtController.hpp"
+#include "tinia/trell/FRVGLJobController.hpp"
+#include <QtWebSockets/QWebSocket>
+#include <QObject>
+
+
+tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, QObject* parent /*= Q_NULLPTR */ )
+    :
+    m_glJob( glJob )
+{
+
+}
+
+tinia::trell::FRVQtController::~FRVQtController()
+{
+    m_glJob->finish();
+    m_socket->close();
+}
+
+void tinia::trell::FRVQtController::closed()
+{
+
+}
+
+void tinia::trell::FRVQtController::onConnect()
+{
+
+}
+
+void tinia::trell::FRVQtController::onMessageReceived()
+{
+
+}
+
+

### 2014-11-28 include/tinia/qtcontroller/moc/OpenGLServerGrabber.hpp
qtcontroller and qimage-scaling

@@ -42,7 +42,7 @@ public:
     grabRGB( tinia::jobcontroller::OpenGLJob* job,
              unsigned int width,
              unsigned int height,
-             const std::string &key);
+             const std::string &key );
 
     /** Grabs the depth buffer of a view
      *
@@ -55,7 +55,9 @@ public:
     grabDepth( tinia::jobcontroller::OpenGLJob* job,
                unsigned int width,
                unsigned int height,
-               const std::string &key);
+               const std::string &key,
+               const unsigned depth_w = 0, // The default value 0 means that the size of the canvas (i.e., width x height) will be used for the depth buffer also.
+               const unsigned depth_h = 0 );
 
 private:
     void setupOpenGL();

### 2014-11-28 js/gui/ProxyRenderer.js

@@ -176,6 +176,37 @@ dojo.declare("gui.ProxyRenderer", null, {
         //         Not adding a test for construction-time setting of all of these, just a subset.
         if (true) { // if (this._debugging) {
             //-------------------------------------------------------
+            this.exposedModel.addLocalListener("ap_set_depth_size_128", dojo.hitch(this, function(event) {
+                if(this.exposedModel.getElementValue("ap_set_depth_size_128")) {
+                    this._loadShaders();
+                    this.exposedModel.updateElement("ap_depthWidth", 128);
+                    this.exposedModel.updateElement("ap_depthHeight", 128);
+                } else {
+                    this.exposedModel.updateElement("ap_set_depth_size_128", false);
+                }
+            }));
+            //-------------------------------------------------------
+            this.exposedModel.addLocalListener("ap_set_depth_size_256", dojo.hitch(this, function(event) {
+                if(this.exposedModel.getElementValue("ap_set_depth_size_256")) {
+                    this._loadShaders();
+                    this.exposedModel.updateElement("ap_depthWidth", 256);
+                    this.exposedModel.updateElement("ap_depthHeight", 256);
+                } else {
+                    this.exposedModel.updateElement("ap_set_depth_size_256", false);
+                }
+            }));
+            //-------------------------------------------------------
+            this.exposedModel.addLocalListener("ap_set_depth_size_512", dojo.hitch(this, function(event) {
+                if(this.exposedModel.getElementValue("ap_set_depth_size_512")) {
+                    this._loadShaders();
+                    this.exposedModel.updateElement("ap_depthWidth", 512);
+                    this.exposedModel.updateElement("ap_depthHeight", 512);
+                } else {
+                    this.exposedModel.updateElement("ap_set_depth_size_512", false);
+                }
+            }));
+
+            //-------------------------------------------------------
             this.exposedModel.addLocalListener("ap_reloadShader", dojo.hitch(this, function(event) {
                 if(this.exposedModel.getElementValue("ap_reloadShader")) {
                     this._loadShaders();
@@ -194,7 +225,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
             //-------------------------------------------------------
[…]

### 2014-11-28 src/qtcontroller/OpenGLServerGrabber.cpp

@@ -53,7 +53,8 @@ OpenGLServerGrabber::grabRGB( jobcontroller::OpenGLJob *job,
     glPixelStorei( GL_PACK_ALIGNMENT, 4 );
 
     // make sure that buffer is large enough to hold raw image
-    size_t req_buffer_size = scanline_size*height*3;
+    size_t req_buffer_size = scanline_size*height*3; // Why *3 here?!
+    std::cout << "scanline_size=" << scanline_size << ", width=" << width << ", height=" << height << ", req_buffer_size=" << req_buffer_size << ", w*h*3=" << width*height*3 << std::endl;
     if( (m_buffer == NULL) || (m_buffer_size < req_buffer_size) ) {
         if( m_buffer != NULL ) {
             delete m_buffer;
@@ -84,7 +85,9 @@ void
 OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
                                 unsigned int width,
                                 unsigned int height,
-                                const std::string& key)
+                                const std::string& key,
+                                const unsigned depth_w, /* = 0 */
+                                const unsigned depth_h ) /* = 0 */
 {
     if( !m_openglIsReady ) {
         setupOpenGL();
@@ -92,6 +95,17 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
     if(m_width != width || m_height != height) {
         resize(width, height);
     }
+
+    if ( (depth_w!=0) || (depth_h!=0) ) {
+
+        std::cout << "er her" << std::endl;
+        throw "not implemented";
+        
+
+    } else {
+        
+        // Old QImage path
+    
 
     glBindFramebuffer( GL_FRAMEBUFFER, m_fbo );
     glViewport( 0, 0, width, height );
@@ -101,11 +115,12 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
     // job->renderFrame( "session", key, m_fbo, width, height );
 
     // QImage requires scanline size to be a multiple of 32 bits.
-    size_t scanline_size = 4*width;
+    size_t scanline_size = 4*width; // 4 bytes per fragment, since each fragment gets a float
     glPixelStorei( GL_PACK_ALIGNMENT, 1 ); // 4 and 1 equally good, in this case?
 
     // make sure that buffer is large enough to hold raw image
-    size_t req_buffer_size = scanline_size*height*4;
[…]

### 2014-11-28 src/qtcontroller/ServerThread.cpp

@@ -107,33 +107,61 @@ public:
         }
         m_key    = ( proper_key_to_use != "" ? proper_key_to_use : arguments.get<2>() );
     }
-    
+
     ~SnapshotAsTextFetcher()
     {
         using namespace tinia::qtcontroller::impl;
-        QImage img( m_gl_grabber->imageBuffer(),
-                    m_width,
-                    m_height,
-                    QImage::Format_RGB888 );
-    
-        // This is a temporary fix. The image is reflected through the horizontal
-        // line y=height ((x, y) |--> (x, h-y) ).
-        QTransform flipTransformation(1, 0,
-                                      0, -1,
-                                      0, m_height);
-        img = img.transformed(flipTransformation);
-        if ( (m_depth_w!=m_width) || (m_depth_h!=m_height) ) {
-            img = img.scaled(m_depth_w, m_depth_h);
-        }
-        QBuffer qBuffer;
-        if (m_pngMode) {
-            img.save(&qBuffer, "png");
+
+        bool use_qt_scaling;
+        m_job->getExposedModel()->getElementValue( "ap_use_qt_img_scaling", use_qt_scaling );
+        if (use_qt_scaling) {
+
+            QImage img( m_gl_grabber->imageBuffer(),
+                        m_width,
+                        m_height,
+                        QImage::Format_RGB888 );
+            // This is a temporary fix. The image is reflected through the horizontal
+            // line y=height ((x, y) |--> (x, h-y) ).
+            QTransform flipTransformation(1, 0,
+                                          0, -1,
+                                          0, m_height);
+            img = img.transformed(flipTransformation);
+            if ( (m_depth_w!=m_width) || (m_depth_h!=m_height) ) {
+                img = img.scaled(m_depth_w, m_depth_h); // Should ignore aspect ratio, and do no (bi-)linear filtering, according to the man pages.
+            }
+            QBuffer qBuffer;
+            if (m_pngMode) {
+                img.save(&qBuffer, "png");
+            } else {
+                img.save(&qBuffer, "jpg", m_jpg_quality);
+            }
+            
+            QString str( QByteArray( qBuffer.data(), int(qBuffer.size()) ).toBase64() );
+            m_reply << str;
+
         } else {
-            img.save(&qBuffer, "jpg", m_jpg_quality);
+            
+            QImage img( m_gl_grabber->imageBuffer(),
+                        m_depth_w,
+                        m_depth_h,
[…]

### 2014-11-28 src/utils/ProxyDebugGUI.cpp

@@ -91,19 +91,19 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addAnnotation("ap_decayMode", "Splats decaying from center");
         model->addElement<bool>( "ap_roundSplats", false );
         model->addAnnotation("ap_roundSplats", "Circular splats");
-        model->addElement<bool>( "ap_screenSpaceSized", true );
+        model->addElement<bool>( "ap_screenSpaceSized", false );
         model->addAnnotation("ap_screenSpaceSized", "Screen-space-sized splats");
-        model->addConstrainedElement<int>("ap_overlap", 200, 1, 300);
+        model->addConstrainedElement<int>("ap_overlap", 100, 1, 300);
         model->addAnnotation("ap_overlap", "Overlap factor)");
         model->addElement<bool>( "ap_alwaysShowMostRecent", true );
         model->addAnnotation("ap_alwaysShowMostRecent", "Most recent model in front");
-        model->addConstrainedElement<int>("ap_splats", 64, 2, 512);
+        model->addConstrainedElement<int>("ap_splats", 12, 2, 512);
         model->addAnnotation("ap_splats", "Number of splats)");
         model->addElement<bool>( "ap_resetAllModels", false );
         model->addAnnotation("ap_resetAllModels", "Remove all models, update now");
         model->addElement<bool>( "ap_useISTC", true );
         model->addAnnotation("ap_useISTC", "Use intra-splat texcoo");
-        model->addElement<bool>( "ap_splatOutline", false );
+        model->addElement<bool>( "ap_splatOutline", true );
         model->addAnnotation("ap_splatOutline", "Square splat outline");
         model->addElement<bool>( "ap_reloadShader", false );
         model->addAnnotation("ap_reloadShader", "Reload shader");
@@ -115,10 +115,20 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
     }
 
     if (with_depth_buffer_manipulation) {
-        model->addConstrainedElement<int>("ap_depthWidth", 512, 4, 1024);
+        model->addConstrainedElement<int>("ap_depthWidth", 512, 64, 1024);
         model->addAnnotation("ap_depthWidth", "Depth buffer Width)");
-        model->addConstrainedElement<int>("ap_depthHeight", 512, 4, 1024);
[…]

### 2014-11-28 js/gui/ProxyRenderer.js
Minor bugfix

@@ -194,7 +194,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
             //-------------------------------------------------------
             if ( this.exposedModel.hasKey("ap_overlap") ) {
-                this._alwaysShowMostRecent = this.exposedModel.getElementValue("ap_overlap") / 100.0;
+                this._splatOverlap = this.exposedModel.getElementValue("ap_overlap") / 100.0;
             }
             this.exposedModel.addLocalListener( "ap_overlap", dojo.hitch(this, function(event) {
                 this._splatOverlap = this.exposedModel.getElementValue("ap_overlap") / 100.0;

### 2014-11-28 examples/cuberenderer/sockets_main.cpp
Now sets up a server to listen for websocket connections.
Initialisation order is very specific and should be refactored later.

@@ -18,6 +18,10 @@
 
 #include "CubeJob.hpp"
 #include "tinia/trell/FRVGLJobController.hpp"
+#include "tinia/trell/FRVQtController.hpp"
+#include <QCoreApplication>
+
+
 int main(int argc, char** argv)
 {
     using namespace tinia::example;
@@ -27,5 +31,11 @@ int main(int argc, char** argv)
     trellController->setJob(testJob);
 
     trellController->run(argc, argv);
-    exit(EXIT_SUCCESS);
+    tinia::trell::FRVQtController* server  = new tinia::trell::FRVQtController( trellController );
+
+    QCoreApplication blah(argc, argv );
+    QObject::connect( server, &tinia::trell::FRVQtController::closed, &blah, &QCoreApplication::quit );
+    auto retVal = blah.exec();
+
+    exit( retVal );
 }

### 2014-11-28 include/tinia/trell/FRVGLJobController.hpp

@@ -47,12 +47,14 @@ public:
     void finish();
     void fail();
     void addScript( const std::string& script) {}
-    
-protected:
-    /** \copydoc MessageBox::init */
+
     virtual
     bool
     init( );
+
+protected:
+    /** \copydoc MessageBox::init */
+    
 
 
     /** \copydoc MessageBox::cleanup */

### 2014-11-28 include/tinia/trell/FRVQtController.hpp

@@ -3,6 +3,7 @@
 #include <QObject>
 
 QT_FORWARD_DECLARE_CLASS(QWebSocket);
+QT_FORWARD_DECLARE_CLASS(QWebSocketServer);
 
 namespace tinia{ namespace trell {
 
@@ -21,13 +22,17 @@ Q_SIGNALS:
     void closed();
     
 private Q_SLOTS:
-    void onConnect();
-    void onMessageReceived();
+     void onNewConnection();
+    void processTextMessage(QString message);
+    void processBinaryMessage(QByteArray message);
+    void socketDisconnected();
 
 private:
-    QWebSocket* m_socket;
+    QWebSocketServer* m_pWebSocketServer;
+    QList<QWebSocket*> m_clients;
+
     FRVGLJobController* m_glJob;
-    
+    unsigned int m_port;
 };
 
 

### 2014-11-28 src/trell/FRVGLJobController.cpp

@@ -92,8 +92,8 @@ FRVGLJobController::init()
 
    
     
-
-    return m_openGLJob->initGL();
+    //initialise OpenGL job
+    return m_openGLJob->init() && m_openGLJob->initGL();
 
 }
 
@@ -303,7 +303,7 @@ FRVGLJobController::checkForGLError() const
 
 void FRVGLJobController::finish()
 {
-    std::cerr << "will never ever be called, as this prototype does not support this call" << std::endl;
+    m_openGLJob->quit();
 }
 
 void FRVGLJobController::fail()
@@ -315,17 +315,21 @@ int FRVGLJobController::run( int argc, char** argv )
 {
     m_argc = argc;
     m_argv = argv;
-    init();
-    m_openGLJob->initGL();
     //setup polling of websocket messages
     //ie update/idle function
-    
+    if( !init() ){
+        glDebugMessageInsert( GL_DEBUG_SOURCE_APPLICATION, GL_DEBUG_TYPE_ERROR, 113, GL_DEBUG_SEVERITY_HIGH, -1, "Could not initialise OpenGLJob, terminating");
+        exit(113);
+    }
     //call glutMainloop.
+
+
 
     m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
     glutSwapBuffers();
     
-    float rotation = 0.001, prev = 0;
+    static float rotation = 0.001;
+    float prev = 0;
     tinia::example::CubeJob* cj = (tinia::example::CubeJob*) m_openGLJob;
     for (int i = 0; i < 150; i++ )
     {

### 2014-11-28 src/trell/FRVQtController.cpp

@@ -1,35 +1,79 @@
 #include "tinia/trell/FRVQtController.hpp"
 #include "tinia/trell/FRVGLJobController.hpp"
 #include <QtWebSockets/QWebSocket>
+#include <QtWebSockets/QWebSocketServer>
 #include <QObject>
 
 
 tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, QObject* parent /*= Q_NULLPTR */ )
     :
-    m_glJob( glJob )
+    QObject(parent),
+    m_glJob( glJob ),
+    m_port(8081)
 {
+    m_pWebSocketServer = new QWebSocketServer(QStringLiteral("Echo Server"),
+        QWebSocketServer::NonSecureMode, this);
+    // Listen on the specified port in the local network:
+   
 
+    if(m_pWebSocketServer->listen(QHostAddress::Any, m_port)) {
+        // if listening successful:
+        qDebug() << "SocketServer listening on port " << m_port;
+        connect(m_pWebSocketServer, &QWebSocketServer::newConnection,
+            this, &FRVQtController::onNewConnection);
+        connect(m_pWebSocketServer, &QWebSocketServer::closed, this, &FRVQtController::closed);
+    }
 }
 
 tinia::trell::FRVQtController::~FRVQtController()
 {
     m_glJob->finish();
-    m_socket->close();
+    for( int i = 0 ; i < m_clients.length(); ++i ){
+        m_clients[i]->close();
+    }
+    m_pWebSocketServer->close();
+    qDeleteAll(m_clients.begin(), m_clients.end());
 }
 
-void tinia::trell::FRVQtController::closed()
+//void tinia::trell::FRVQtController::closed()
+//{
+//
+//}
+
+void tinia::trell::FRVQtController::onNewConnection()
 {
+    QWebSocket *pSocket = m_pWebSocketServer->nextPendingConnection();
+
+    connect(pSocket, &QWebSocket::textMessageReceived, this, &FRVQtController::processTextMessage);
+    connect(pSocket, &QWebSocket::binaryMessageReceived, this, &FRVQtController::processBinaryMessage);
+    connect(pSocket, &QWebSocket::disconnected, this, &FRVQtController::socketDisconnected);
+
+    m_clients << pSocket;
 
 }
 
-void tinia::trell::FRVQtController::onConnect()
+void tinia::trell::FRVQtController::processTextMessage(QString message)
 {
-
[…]

### 2014-11-28 include/tinia/trell/FRVGLJobController.hpp
slight refactoring for testing with a web-page.

@@ -52,6 +52,10 @@ public:
     bool
     init( );
 
+
+    void render();
+
+
 protected:
     /** \copydoc MessageBox::init */
     
@@ -91,6 +95,8 @@ private:
 
     bool
     checkForGLError() const;
+
+
 
 };
 

### 2014-11-28 src/trell/FRVGLJobController.cpp

@@ -322,12 +322,15 @@ int FRVGLJobController::run( int argc, char** argv )
         exit(113);
     }
     //call glutMainloop.
-
-
-
     m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
     glutSwapBuffers();
-    
+    render();
+
+    return 1;
+}
+
+void FRVGLJobController::render()
+{    
     static float rotation = 0.001;
     float prev = 0;
     tinia::example::CubeJob* cj = (tinia::example::CubeJob*) m_openGLJob;
@@ -337,9 +340,7 @@ int FRVGLJobController::run( int argc, char** argv )
 
         m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
         glutSwapBuffers();
-    }
-
-    return 1;
+    }   
 }
 
 

### 2014-11-28 src/trell/FRVQtController.cpp

@@ -47,6 +47,7 @@ void tinia::trell::FRVQtController::onNewConnection()
     connect(pSocket, &QWebSocket::textMessageReceived, this, &FRVQtController::processTextMessage);
     connect(pSocket, &QWebSocket::binaryMessageReceived, this, &FRVQtController::processBinaryMessage);
     connect(pSocket, &QWebSocket::disconnected, this, &FRVQtController::socketDisconnected);
+    m_glJob->render();
 
     m_clients << pSocket;
 
@@ -54,15 +55,12 @@ void tinia::trell::FRVQtController::onNewConnection()
 
 void tinia::trell::FRVQtController::processTextMessage(QString message)
 {
-    static char* testing = "testing";
-    m_glJob->run(1, &testing );
+     m_glJob->render();
 }
 
 void tinia::trell::FRVQtController::processBinaryMessage(QByteArray message)
 {
-    static char* testing = "testing";
-    m_glJob->run(1, &testing );
-
+    m_glJob->render();
 }
 
 void tinia::trell::FRVQtController::socketDisconnected()

### 2014-11-28 js/gui/Canvas.js
on the way to reduced depth img size

@@ -427,7 +427,7 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                     if (response_obj[this._key].snaptype == "jpg") {
                         snaptype = snaptype + parseInt(this._modelLib.getElementValue("ap_jpgQuality")/10);
                     }
-                    console.log("new snaptype = " + snaptype);
+                    // console.log("new snaptype = " + snaptype);
                     this._snapshotTimings.update( snaptype, (tmp - response_obj[this._key].timestamp) );
                     this._snapshotTimings.print();
                     this._autoSelectSnapshotType(this._snapshotTimings);

### 2014-11-28 js/gui/ProxyRenderer.js

@@ -300,6 +300,30 @@ dojo.declare("gui.ProxyRenderer", null, {
             }) );
 
             //-------------------------------------------------------
+            if ( this.exposedModel.hasKey("ap_mid_splat_sampling") ) {
+                this._loadShaders();
+            }
+            this.exposedModel.addLocalListener( "ap_mid_splat_sampling", dojo.hitch(this, function(event) {
+                this._loadShaders();
+            }) );
+
+            //-------------------------------------------------------
+            if ( this.exposedModel.hasKey("ap_small_delta_sampling") ) {
+                this._loadShaders();
+            }
+            this.exposedModel.addLocalListener( "ap_small_delta_sampling", dojo.hitch(this, function(event) {
+                this._loadShaders();
+            }) );
+
+            //-------------------------------------------------------
+            if ( this.exposedModel.hasKey("ap_larger_delta_sampling") ) {
+                this._loadShaders();
+            }
+            this.exposedModel.addLocalListener( "ap_larger_delta_sampling", dojo.hitch(this, function(event) {
+                this._loadShaders();
+            }) );
+
+            //-------------------------------------------------------
             if ( this.exposedModel.hasKey("ap_autoProxyAlgo") ) {
                 console.log("new algo: " + this.exposedModel.getElementValue("ap_autoProxyAlgo") );
                 this._initProxyCoverage( this.exposedModel.getElementValue("ap_autoProxyAlgo"), this.gl );
@@ -421,6 +445,36 @@ dojo.declare("gui.ProxyRenderer", null, {
             splat_vs_src = "// #define MID_TEXEL_SAMPLING\n" + splat_vs_src;
             splat_fs_src = "// #define MID_TEXEL_SAMPLING\n" + splat_fs_src;
             console.log("// #define MID_TEXEL_SAMPLING");
+        }
+
+        if ( (this.exposedModel.hasKey("ap_mid_splat_sampling")) && (this.exposedModel.getElementValue("ap_mid_splat_sampling")) ) {
+            splat_vs_src = "#define MID_SPLAT_SAMPLING\n" + splat_vs_src;
+            splat_fs_src = "#define MID_SPLAT_SAMPLING\n" + splat_fs_src;
+            console.log("#define MID_SPLAT_SAMPLING");
[…]

### 2014-11-28 src/qtcontroller/OpenGLServerGrabber.cpp

@@ -53,8 +53,9 @@ OpenGLServerGrabber::grabRGB( jobcontroller::OpenGLJob *job,
     glPixelStorei( GL_PACK_ALIGNMENT, 4 );
 
     // make sure that buffer is large enough to hold raw image
-    size_t req_buffer_size = scanline_size*height*3; // Why *3 here?!
-    std::cout << "scanline_size=" << scanline_size << ", width=" << width << ", height=" << height << ", req_buffer_size=" << req_buffer_size << ", w*h*3=" << width*height*3 << std::endl;
+//    size_t req_buffer_size = scanline_size*height*3; // Why *3 here?!
+    size_t req_buffer_size = scanline_size*height;
+    std::cout << "rgb scanline_size=" << scanline_size << ", width=" << width << ", height=" << height << ", req_buffer_size=" << req_buffer_size << ", w*h*3=" << width*height*3 << std::endl;
     if( (m_buffer == NULL) || (m_buffer_size < req_buffer_size) ) {
         if( m_buffer != NULL ) {
             delete m_buffer;
@@ -96,17 +97,6 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
         resize(width, height);
     }
 
-    if ( (depth_w!=0) || (depth_h!=0) ) {
-
-        std::cout << "er her" << std::endl;
-        throw "not implemented";
-        
-
-    } else {
-        
-        // Old QImage path
-    
-
     glBindFramebuffer( GL_FRAMEBUFFER, m_fbo );
     glViewport( 0, 0, width, height );
     // Should not be necessary to render, we will always have rendered for the RGB-grab before doing a depth-grab!
@@ -119,8 +109,9 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
     glPixelStorei( GL_PACK_ALIGNMENT, 1 ); // 4 and 1 equally good, in this case?
 
     // make sure that buffer is large enough to hold raw image
-    size_t req_buffer_size = scanline_size*height*4; // Why 4 here?!
-    std::cout << "scanline_size=" << scanline_size << ", width=" << width << ", height=" << height << ", req_buffer_size=" << req_buffer_size << ", w*h*4=" << width*height*4 << std::endl;
+//        size_t req_buffer_size = scanline_size*height*4; // Why 4 here?!
+    size_t req_buffer_size = scanline_size*height;
[…]

### 2014-11-28 src/utils/ProxyDebugGUI.cpp

@@ -119,7 +119,7 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addAnnotation("ap_depthWidth", "Depth buffer Width)");
         model->addConstrainedElement<int>("ap_depthHeight", 512, 64, 1024);
         model->addAnnotation("ap_depthHeight", "Depth buffer height)");
-        model->addElement<bool>( "ap_mid_texel_sampling", false );
+        model->addElement<bool>( "ap_mid_texel_sampling", false );              // false best
         model->addAnnotation("ap_mid_texel_sampling", "Sample mid-texel");
         model->addElement<bool>( "ap_use_qt_img_scaling", true );
         model->addAnnotation("ap_use_qt_img_scaling", "Use QImage::scaled()");
@@ -129,6 +129,12 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addAnnotation("ap_set_depth_size_256", "256");
         model->addElement<bool>( "ap_set_depth_size_512", false );
         model->addAnnotation("ap_set_depth_size_512", "512");
+        model->addElement<bool>( "ap_small_delta_sampling", true );             // true best
+        model->addAnnotation("ap_small_delta_sampling", "Small delta");
+        model->addElement<bool>( "ap_larger_delta_sampling", false );           // cannot see any difference
+        model->addAnnotation("ap_larger_delta_sampling", "Larger delta");
+        model->addElement<bool>( "ap_mid_splat_sampling", false );
+        model->addAnnotation("ap_mid_splat_sampling", "Sample mid-splat");
     }
 }
 
@@ -234,8 +240,11 @@ tinia::model::gui::Grid *ProxyDebugGUI::getGrid()
         mainGrid->setChild(row, 2, new tinia::model::gui::Button("ap_set_depth_size_512"));
         row++;
         mainGrid->setChild(row, 0, new tinia::model::gui::CheckBox("ap_mid_texel_sampling"));
+        mainGrid->setChild(row, 1, new tinia::model::gui::CheckBox("ap_small_delta_sampling"));
+        mainGrid->setChild(row, 2, new tinia::model::gui::CheckBox("ap_larger_delta_sampling"));
         row++;
         mainGrid->setChild(row, 0, new tinia::model::gui::CheckBox("ap_use_qt_img_scaling"));
[…]

### 2014-11-30 include/tinia/trell/IPCGLJobController.hpp
First trell attempt

@@ -62,6 +62,8 @@ protected:
                    TrellPixelFormat    pixel_format,
                    const size_t        width,
                    const size_t        height,
+                   const size_t        depth_width,
+                   const size_t        depth_height,
                    const std::string&  session,
                    const std::string&  key );
 

### 2014-11-30 include/tinia/trell/IPCJobController.hpp

@@ -122,6 +122,8 @@ protected:
                    TrellPixelFormat    pixel_format,
                    const size_t        width,
                    const size_t        height,
+                   const size_t        depth_width,
+                   const size_t        depth_height,
                    const std::string&  session,
                    const std::string&  key );
     virtual

### 2014-11-30 src/mod_trell/mod_trell_url_decode.c

@@ -243,6 +243,8 @@ trell_decode_path_info( trell_dispatch_info_t* dispatch_info, request_rec *r )
     dispatch_info->m_base64 = 0;
     dispatch_info->m_width = 0;
     dispatch_info->m_height = 0;
+    dispatch_info->m_depth_w = 0;
+    dispatch_info->m_depth_h = 0;
 
     int i;
     int o = -1;
@@ -438,8 +440,10 @@ trell_decode_path_info( trell_dispatch_info_t* dispatch_info, request_rec *r )
         dispatch_info->m_pixel_format = TRELL_PIXEL_FORMAT_RGB_CUSTOM_DEPTH;
         dispatch_info->m_base64 = 1;
         if( (trell_hash_strncpy( r, dispatch_info->m_key, form, "key", TRELL_KEYID_MAXLENGTH-1 ) == 0 )
-                || (trell_hash_atoi( r, component, request, &dispatch_info->m_width, form, "width", 1 ) == 0 )
-                || (trell_hash_atoi( r, component, request, &dispatch_info->m_height, form, "height", 1 ) == 0 ) )
+            || (trell_hash_atoi( r, component, request, &dispatch_info->m_width, form, "width", 1 ) == 0 )
+            || (trell_hash_atoi( r, component, request, &dispatch_info->m_height, form, "height", 1 ) == 0 )
+            || (trell_hash_atoi( r, component, request, &dispatch_info->m_depth_w, form, "depth_w", 1 ) == 0 )
+            || (trell_hash_atoi( r, component, request, &dispatch_info->m_depth_h, form, "depth_h", 1 ) == 0 ) )
         {
             ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, r, "%s: parsing %s failed.", r->handler, request );
             return HTTP_BAD_REQUEST;

### 2014-11-30 src/mod_trell/pass_reply_png.c

@@ -235,7 +235,7 @@ static int trell_pass_reply_png_bundle( void*          data,
     }
 
     static int num_of_keys = 0;
-    size_t buffer_img_size=0, padded_img_size=0, matrix_size = sizeof(float)*16*w_depth, canvas_size=0;
+    size_t buffer_img_size=0, depth_img_size=0, padded_img_size=0, padded_depth_size=0, matrix_size = sizeof(float)*16*w_depth, canvas_size=0;
 
     size_t offset = 0;
     if( part == 0 ) {
@@ -254,6 +254,9 @@ static int trell_pass_reply_png_bundle( void*          data,
             }
         }
 
+        // Some (all?) of this doesn't make (the same kind of) sense as before, now that we send two images from this routine, and they may
+        // even differ in size...
+
         encoder_state->dispatch_info->m_png_entry = apr_time_now();
         // first invocation
         tinia_msg_image_t* msg = (tinia_msg_image_t*)buffer;
@@ -264,8 +267,11 @@ static int trell_pass_reply_png_bundle( void*          data,
         encoder_state->width  = msg->width;
         encoder_state->height = msg->height;
         buffer_img_size       = 3 * encoder_state->width * encoder_state->height;
+        depth_img_size        = 3 * msg->depth_width * msg->depth_height;
+        ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, encoder_state->r, "depth_img_size = %lu.", depth_img_size );
         padded_img_size       = 4*( (buffer_img_size+3)/4 );
-        canvas_size           = padded_img_size + w_depth*( padded_img_size + 2*matrix_size );
+        padded_depth_size     = 4*( (depth_img_size+3)/4 );
+        canvas_size           = padded_img_size + w_depth*( padded_depth_size + 2*matrix_size );
         encoder_state->buffer = apr_palloc( encoder_state->r->pool, num_of_keys * canvas_size );
         const size_t filtered_img_size_bound = (3*encoder_state->width+1) * encoder_state->height; // The +1 is for the png filter flag
         encoder_state->filtered = apr_palloc( encoder_state->r->pool, filtered_img_size_bound + w_depth*2*matrix_size ); // Just in case the image is smaller than 4*16 bytes!
@@ -343,6 +349,22 @@ static int trell_pass_reply_png_bundle( void*          data,
                 BB_APPEND_STRING( encoder_state->r->pool, bb, ", \"depth\": \"" );
                 {
[…]

### 2014-11-30 src/trell/IPCGLJobController.cpp

@@ -389,6 +389,8 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
                                    TrellPixelFormat    pixel_format,
                                    const size_t        width,
                                    const size_t        height,
+                                   const size_t        depth_width,
+                                   const size_t        depth_height,
                                    const std::string&  session,
                                    const std::string&  key )
 {
@@ -502,35 +504,69 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
         glReadPixels( 0, 0, width, height, GL_RGB, GL_UNSIGNED_BYTE, buffer_pos );
         buffer_pos += 4*((width*height*3 + 3)/4); // As long as GL_PACK_ALIGNMENT is set to 1 above, this should be ok. (I.e., no padding for single scan lines.)
         // NB! We read four bytes per pixel, then convert to three, meaning that the buffer must be large enough for four!!!
+        // (This should be taken care of further up the stack. The caller of this method is IPCJobController::handle().)
         glReadPixels( 0, 0, width, height, GL_DEPTH_COMPONENT, GL_FLOAT, buffer_pos );
-        // Depth encoded as 24 bit fixed point values.
-        for (size_t i=0; i<width*height; i++) {
-            float value = (float)( ((GLfloat *)buffer_pos)[i] );
-            for (size_t j=0; j<3; j++) {
-                (buffer_pos)[3*i+j] = (unsigned char)( floor(value*255.0) );
-                value = 255.0*value - floor(value*255.0);
+
+        // downsampling, without bi-linear interpolation
+        if( m_logger_callback != NULL ) { // (This goes to /tmp/job-id.stderr)
+            m_logger_callback( m_logger_data, 0, package.c_str(), "Current canvas and depth buffer size: %d %d and %d %d", width, height, depth_width, depth_height );
+        }
+
+        if ( (depth_width!=width) || (depth_height!=height) ) {
+            float *tmp_buffer = new float[depth_width*depth_height];
+            const float * const m_buf = (float *)buffer_pos;
+
+            for (size_t i=0; i<depth_height; i++) {
+                size_t ii = size_t( floor( (i*height)/double(depth_height) + 0.5 ) );
+                for (size_t j=0; j<depth_width; j++) {
[…]

### 2014-11-30 src/trell/IPCJobController.cpp

@@ -82,11 +82,13 @@ IPCJobController::cleanup()
 
 bool
 IPCJobController::onGetSnapshot( char*               buffer,
-                               TrellPixelFormat    pixel_format,
-                               const size_t        width,
-                               const size_t        height,
-                               const std::string&  session,
-                               const std::string&  key )
+                                 TrellPixelFormat    pixel_format,
+                                 const size_t        width,
+                                 const size_t        height,
+                                 const size_t        depth_width,
+                                 const size_t        depth_height,
+                                 const std::string&  session,
+                                 const std::string&  key )
 {
     return false;
 }
@@ -223,6 +225,10 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
         session = std::string( q->session_id );
         key     = std::string( q->key );
 
+        // These are coming from the url-parameters depth_w and depth_h...
+        int depth_width = q->depth_h;
+        int depth_height = q->depth_h;
+
         std::string key_list_string = std::string( q->viewer_key_list );
         std::vector<std::string> key_list;
         boost::split( key_list, key_list_string, boost::is_any_of(",") );
@@ -241,7 +247,10 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
                 // We could have the size of each canvas associated with the keys we have, but this information is currently
                 // not available. This will only work for equally sized canvases then, and no problem will be checked for or detected if it
                 // is not the case!
-                data_size = 4*((3*w*h+3)/4) * 2 + sizeof(float)*16*2; // Two long word aligned images + 2 matrices
+//                data_size = 4*((3*w*h+3)/4) * 2 + sizeof(float)*16*2; // Two long word aligned images + 2 matrices
+                data_size = 4*((3*w*h+3)/4);                        // One long word aligned image for rgb
+                data_size += 4*((3*depth_width*depth_height+3)/4);  // + one for depth
+                data_size += sizeof(float)*16*2;                    // + 2 matrices
                 // ... times the number of keys:
                 data_size *= key_list.size();
                 buf_size_required = 4*((3*w*h+3)/4) + 4*w*h + sizeof(float)*16*2; // One long word aligned image + one depth buffer (with floats) + 2 matrices
[…]

### 2014-11-30 js/gui/Canvas.js
Reduced depth resolution also works for trell. Some cleanup needed.

@@ -203,8 +203,9 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                 // console.log("/model/updateParsed: response[" + this._key + "].view = " + response_obj[this._key].view);
                                 // console.log("/model/updateParsed: response[" + this._key + "].proj = " + response_obj[this._key].proj);
                                 var depthwidth = response_obj[this._key].depthwidth;
-                                console.log("Receieved depth width: " + depthwidth + ", current model's depth width: " + this._modelLib.getElementValue("ap_depthWidth"));
-                                if ( (depthwidth==this._modelLib.getElementValue("ap_depthWidth")) || (depthwidth==0) ) {
+                                // console.log("Receieved depth width: " + depthwidth + ", current model's depth width: " + this._modelLib.getElementValue("ap_depthWidth"));
+                                // if ( (depthwidth==this._modelLib.getElementValue("ap_depthWidth")) || (depthwidth==0) )
+                                {
                                     this._setImageFromText( response_obj[this._key].rgb, response_obj[this._key].depth, response_obj[this._key].view, response_obj[this._key].proj );
                                     var snaptype = response_obj[this._key].snaptype;
                                     if (response_obj[this._key].snaptype == "jpg") {
@@ -214,8 +215,8 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                     this._snapshotTimings.update( snaptype, (t0 - response_obj[this._key].timestamp) );
                                     this._snapshotTimings.print();
                                     this._autoSelectSnapshotType(this._snapshotTimings);
-                                } else {
-                                    console.log("mismatch in depth size");
+//                                } else {
+//                                    console.log("mismatch in depth size");
                                 }
                             })
                         });
@@ -435,8 +436,9 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
 //                console.log("/model/updateSendPartialComplete: response[" + this._key + "].view = " + response_obj[this._key].view);
 //                console.log("/model/updateSendPartialComplete: response[" + this._key + "].proj = " + response_obj[this._key].proj);
                 var depthwidth = response_obj[this._key].depthwidth;
-                console.log("Receieved depth width: " + depthwidth + ", current model's depth width: " + this._modelLib.getElementValue("ap_depthWidth"));
[…]

### 2014-11-30 src/mod_trell/pass_reply_png.c

@@ -395,10 +395,10 @@ static int trell_pass_reply_png_bundle( void*          data,
 
 
                 }
-                const float * const MV = (const float * const)( encoder_state->buffer + i*canvas_size + 2*padded_img_size );
+                const float * const MV = (const float * const)( encoder_state->buffer + i*canvas_size + padded_img_size + padded_depth_size );
                 BB_APPEND_STRING( encoder_state->r->pool, bb, "\", view: \"%g %g %g %g %g %g %g %g %g %g %g %g %g %g %g %g\"",
                                   MV[0], MV[1], MV[2], MV[3], MV[4], MV[5], MV[6], MV[7], MV[8], MV[9], MV[10], MV[11], MV[12], MV[13], MV[14], MV[15] );
-                const float * const PM = (const float * const)( encoder_state->buffer + i*canvas_size + 2*padded_img_size + sizeof(float)*16 );
+                const float * const PM = (const float * const)( encoder_state->buffer + i*canvas_size + padded_img_size + padded_depth_size + sizeof(float)*16 );
                 BB_APPEND_STRING( encoder_state->r->pool, bb, ", proj: \"%g %g %g %g %g %g %g %g %g %g %g %g %g %g %g %g\"",
                                   PM[0], PM[1], PM[2], PM[3], PM[4], PM[5], PM[6], PM[7], PM[8], PM[9], PM[10], PM[11], PM[12], PM[13], PM[14], PM[15] );
             }

### 2014-11-30 src/qtcontroller/ServerThread.cpp

@@ -341,7 +341,8 @@ void ServerThread::getSnapshotTxt( QTextStream &os, const QString &request,
             }
             os << "\"";
         }
-        os << ",\n\"revision\": " << revision << ",\n\"timestamp\": " << timestamp << ",\n\"snaptype\": " << "\"" << snaptype.c_str() << "\" }";
+        os << ",\n\"revision\": " << revision << ",\n\"timestamp\": " << timestamp << ",\n\"snaptype\": " << "\"" << snaptype.c_str() << "\"";
+        os << ",\n\"depthwidth\": " << depth_w << " }";
         if ( i < vk_list.size() - 1 ) {
             os << ", ";
         }

### 2014-12-01 include/tinia/qtcontroller/moc/OpenGLServerGrabber.hpp
Should be ready for testing on FRView now

@@ -57,9 +57,10 @@ public:
                unsigned int height,
                const std::string &key,
                const unsigned depth_w = 0,                  // The default value 0 means that the size of the canvas (i.e., width x height) will be used for the depth buffer also.
-               const unsigned depth_h = 0,                  // NB! Downscaling may still be perfored, but then by the caller of this routine. This happens when QImage.scaled() is
+               const unsigned depth_h = 0,                  // NB! Downscaling may still be performed, but then by the caller of this routine. This happens when QImage.scaled() is
                                                             // used for downscaling, this happens in ~SnapshotAsTextFetcher(), see ServerThread.cpp.
-               const bool bi_linear_filtering = false );    // Only for our own downscaling, QImage.scaled() will never do filtering.
+               const bool bi_linear_filtering = false,      // Only for our own downscaling, QImage.scaled() will never do filtering.
+               const bool depth16 = false );
 
 private:
     void setupOpenGL();

### 2014-12-01 include/tinia/trell/IPCGLJobController.hpp

@@ -64,6 +64,7 @@ protected:
                    const size_t        height,
                    const size_t        depth_width,
                    const size_t        depth_height,
+                   const bool          depth16,
                    const std::string&  session,
                    const std::string&  key );
 

### 2014-12-01 include/tinia/trell/IPCJobController.hpp

@@ -124,6 +124,7 @@ protected:
                    const size_t        height,
                    const size_t        depth_width,
                    const size_t        depth_height,
+                   const bool          depth16,
                    const std::string&  session,
                    const std::string&  key );
     virtual

### 2014-12-01 js/gui/Canvas.js

@@ -202,10 +202,15 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                 var response_obj = eval( '(' + response + ')' );
                                 // console.log("/model/updateParsed: response[" + this._key + "].view = " + response_obj[this._key].view);
                                 // console.log("/model/updateParsed: response[" + this._key + "].proj = " + response_obj[this._key].proj);
-                                var depthwidth = response_obj[this._key].depthwidth;
-                                // console.log("Receieved depth width: " + depthwidth + ", current model's depth width: " + this._modelLib.getElementValue("ap_depthWidth"));
-                                // if ( (depthwidth==this._modelLib.getElementValue("ap_depthWidth")) || (depthwidth==0) )
+                                var depthwidth  = response_obj[this._key].depthwidth;
+                                var depthheight = response_obj[this._key].depthheight;
+                                // console.log("Received depth size: " + depthwidth + " " + depthheight + ", current model's depth size: " +
+                                //             this._modelLib.getElementValue("ap_depthWidth") + " " + this._modelLib.getElementValue("ap_depthHeight"));
+                                if ( ( (depthwidth ==this._modelLib.getElementValue("ap_depthWidth" )) || (depthwidth ==0) ) &&
+                                     ( (depthheight==this._modelLib.getElementValue("ap_depthHeight")) || (depthheight==0) ) )
                                 {
+                                    // Not completely sure, but it may be a good idea to not update with a received bundle, if the depth size does not match what the shader is told...
+                                    // Currently, the shader gets the macros DEPTH_* from these exposed model elements.
                                     this._setImageFromText( response_obj[this._key].rgb, response_obj[this._key].depth, response_obj[this._key].view, response_obj[this._key].proj );
                                     var snaptype = response_obj[this._key].snaptype;
                                     if (response_obj[this._key].snaptype == "jpg") {
@@ -215,8 +220,8 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                     this._snapshotTimings.update( snaptype, (t0 - response_obj[this._key].timestamp) );
                                     this._snapshotTimings.print();
                                     this._autoSelectSnapshotType(this._snapshotTimings);
-//                                } else {
-//                                    console.log("mismatch in depth size");
+                                } else {
[…]

### 2014-12-01 js/gui/ProxyRenderer.js

@@ -124,8 +124,8 @@ dojo.declare("gui.ProxyRenderer", null, {
         this._initProxyCoverage(2, glContext);
 
         // --- For debugging, start
-        this._frameOutputInterval         = 1000;
-        this._frameMeasureInterval        = 100;
+        this._frameOutputInterval         = 50;
+        this._frameMeasureInterval        = 50;
         this._pausePerFrameInMilliseconds = 0; // (100 is useful for GPU fans that we don't want to spin up too much... :-) )
         this._debugSplatCol               = 0;
         this._decayMode                   = 0;
@@ -684,6 +684,7 @@ dojo.declare("gui.ProxyRenderer", null, {
             this.gl.uniform1i( this.gl.getUniformLocation(this._splatProgram, "rgbImage"), 1 );
 
             var t0 = performance.now();
+
             for (var i=0; i<this._proxyModelCoverage.bufferRingSize; i++) {
                 if (this._proxyModelCoverage.proxyModelRing[i].state==2) {
                     this._setUniform1i(this._splatProgram, "splatSetIndex", i);
@@ -745,10 +746,11 @@ dojo.declare("gui.ProxyRenderer", null, {
                 }
             }
 
-            // Timing-stuff
-            //            if (this._debugging) {
+            // Timing-stuff ------------------------------------------------------
             if (false) {
-                //            this.gl.flush();
+                // A couple of problems here... 1) Strange things happen whenever the string put to ap_consoleLog changes in length!!! (Tinia-bug?!?!)
+                // 2) This is only timing rendering, and perhaps not even that... Loading/sending/decoding etc. of snapshots is not properly accounted for.
+
                 if ( this._frameMeasureCounter < this._frameMeasureInterval ) {
                     var t1 = performance.now();
                     this._frameTime += t1 - t0;
@@ -773,6 +775,7 @@ dojo.declare("gui.ProxyRenderer", null, {
                 }
                 this.exposedModel.updateElement("ap_cntr", this.exposedModel.getElementValue("ap_cntr") + 1 );
             }
+            // Timing-stuff ------------------------------------------------------
 
             if ( this._useBlending ) {
                 this.gl.colorMask(false, false, false, true);

### 2014-12-01 src/mod_trell/pass_reply_png.c

@@ -350,20 +350,13 @@ static int trell_pass_reply_png_bundle( void*          data,
                 {
                     p = png; // Reusing the old buffer, should be ok when we use the "transient" buckets that copy data.
 
-
-
-
-
                     // Is this safe, i.e., just poking into this structure?
                     tinia_msg_image_t* msg = (tinia_msg_image_t*)buffer;
-                    ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, encoder_state->r, "trell_pass_reply_png_bundle: Setter ny depth size: %d %d",
-                                   msg->depth_width, msg->depth_height );
+//                    ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, encoder_state->r,
+//                                   "trell_pass_reply_png_bundle: ********** Setting new temporary encoding size for possibly reduced depth size: %d %d",
+//                                   msg->depth_width, msg->depth_height );
                     encoder_state->width  = msg->depth_width;   // Now changing to size of depth image
                     encoder_state->height = msg->depth_height;
-
-
-
-
 
                     int rv = trell_png_encode( data, i*canvas_size + padded_img_size , &p );
                     if ( p-png > total_bound ) {
@@ -380,19 +373,12 @@ static int trell_pass_reply_png_bundle( void*          data,
                     }
                     APR_BRIGADE_INSERT_TAIL( bb, apr_bucket_transient_create( base64, base64_size, bb->bucket_alloc ) );
 
-
-
-
                     // Setting size back again for rgb image
-
-
-                    // Is this safe, i.e., just poking into this structure?
-                    ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, encoder_state->r, "trell_pass_reply_png_bundle: ********** Setter ny encoder-size: %d %d",
-                                   msg->width, msg->height );
-                    encoder_state->width  = msg->width;   // Now changing to size of rgb image
+//                    ap_log_rerror( APLOG_MARK, APLOG_ERR, 0, encoder_state->r,
+//                                   "trell_pass_reply_png_bundle: ********** Resetting encoding size back to rgb image size: %d %d",
+//                                   msg->width, msg->height );
+                    encoder_state->width  = msg->width;   // Now changing back to size of rgb image
                     encoder_state->height = msg->height;
-
-
 
                 }
[…]

### 2014-12-01 src/qtcontroller/OpenGLServerGrabber.cpp

@@ -89,7 +89,8 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
                                 const std::string& key,
                                 const unsigned depth_w, /* = 0 */
                                 const unsigned depth_h, /* = 0 */
-                                const bool bi_linear_filtering ) /* = false */
+                                const bool bi_linear_filtering, /* = false */
+                                const bool depth16 ) /* = false */
 {
     if( !m_openglIsReady ) {
         setupOpenGL();
@@ -214,12 +215,23 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
             }
         }
 
-        // Depth encoded as 24 bit fixed point values.
-        for (size_t i=0; i<depth_w*depth_h; i++) {
-            float value = tmp_buffer[i];
-            for (size_t j=0; j<3; j++) {
-                ((unsigned char *)m_buffer)[3*i+j] = (unsigned char)( floor(value*255.0) );
+        if (depth16) {
+            // Depth encoded as 24 bit fixed point values, least significant bits set to 0
+            for (size_t i=0; i<depth_w*depth_h; i++) {
+                float value = tmp_buffer[i];
+                ((unsigned char *)m_buffer)[3*i+0] = (unsigned char)( floor(value*255.0) );
                 value = 255.0*value - floor(value*255.0);
+                ((unsigned char *)m_buffer)[3*i+1] = (unsigned char)( floor(value*255.0) );
+                ((unsigned char *)m_buffer)[3*i+2] = 0;
+            }
+        } else {
+            // Depth encoded as 24 bit fixed point values.
+            for (size_t i=0; i<depth_w*depth_h; i++) {
+                float value = tmp_buffer[i];
+                for (size_t j=0; j<3; j++) {
+                    ((unsigned char *)m_buffer)[3*i+j] = (unsigned char)( floor(value*255.0) );
+                    value = 255.0*value - floor(value*255.0);
+                }
             }
         }
 
@@ -231,12 +243,23 @@ OpenGLServerGrabber::grabDepth( jobcontroller::OpenGLJob *job,
         // The downsampling will be done after the float->rgb encoding, by QImage.scaled(), so this is a bit dangerous.
         // However, the QImage.scaled() should just downsample without filtering, so it should work. (See ServerThread.cpp)
 
-        // Depth encoded as 24 bit fixed point values.
-        for (size_t i=0; i<width*height; i++) {
[…]

### 2014-12-01 src/qtcontroller/ServerThread.cpp

@@ -130,7 +130,7 @@ public:
                 bool tmp;
                 m_job->getExposedModel()->getElementValue( "ap_simulate_downsampling", tmp );
                 if (!tmp) {
-                    std::cout << "scaling to " << m_depth_w << " x " << m_depth_h << std::endl;
+                    // std::cout << "Scaling (QImage) to " << m_depth_w << " x " << m_depth_h << std::endl;
                     img = img.scaled(m_depth_w, m_depth_h); // Should ignore aspect ratio, and do no (bi-)linear filtering, according to the man pages.
                 } else {
                     std::cout << "Not downsampling with QImage.scaled(), we're simulating downsampling in shaders" << std::endl;
@@ -182,10 +182,12 @@ public:
             m_job->getExposedModel()->getElementValue( "ap_use_qt_img_scaling", use_qt_scaling );
             bool bi_linear_filtering;
             m_job->getExposedModel()->getElementValue( "ap_bi_linear_filtering", bi_linear_filtering );
+            bool depth16;
+            m_job->getExposedModel()->getElementValue( "ap_16_bit_depth", depth16 );
             if (use_qt_scaling) {
-                m_gl_grabber->grabDepth( m_job, m_width, m_height, m_key );
+                m_gl_grabber->grabDepth( m_job, m_width, m_height, m_key, 0, 0, false, depth16 );
             } else {
-                m_gl_grabber->grabDepth( m_job, m_width, m_height, m_key, m_depth_w, m_depth_h, bi_linear_filtering );
+                m_gl_grabber->grabDepth( m_job, m_width, m_height, m_key, m_depth_w, m_depth_h, bi_linear_filtering, depth16 );
             }
             
         }
@@ -342,7 +344,7 @@ void ServerThread::getSnapshotTxt( QTextStream &os, const QString &request,
             os << "\"";
         }
         os << ",\n\"revision\": " << revision << ",\n\"timestamp\": " << timestamp << ",\n\"snaptype\": " << "\"" << snaptype.c_str() << "\"";
-        os << ",\n\"depthwidth\": " << depth_w << " }";
+        os << ",\n\"depthwidth\": " << depth_w << ",\"depthheight\": " << depth_h << " }";
         if ( i < vk_list.size() - 1 ) {
             os << ", ";
         }

### 2014-12-01 src/trell/IPCGLJobController.cpp

@@ -391,6 +391,7 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
                                    const size_t        height,
                                    const size_t        depth_width,
                                    const size_t        depth_height,
+                                   const bool          depth16,
                                    const std::string&  session,
                                    const std::string&  key )
 {
@@ -507,15 +508,14 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
         // (This should be taken care of further up the stack. The caller of this method is IPCJobController::handle().)
         glReadPixels( 0, 0, width, height, GL_DEPTH_COMPONENT, GL_FLOAT, buffer_pos );
 
-        // downsampling, without bi-linear interpolation
         if( m_logger_callback != NULL ) { // (This goes to /tmp/job-id.stderr)
             m_logger_callback( m_logger_data, 0, package.c_str(), "Current canvas and depth buffer size: %d %d and %d %d", width, height, depth_width, depth_height );
         }
 
         if ( (depth_width!=width) || (depth_height!=height) ) {
+            // New path, downsampling, without bi-linear interpolation
             float *tmp_buffer = new float[depth_width*depth_height];
             const float * const m_buf = (float *)buffer_pos;
-
             for (size_t i=0; i<depth_height; i++) {
                 size_t ii = size_t( floor( (i*height)/double(depth_height) + 0.5 ) );
                 for (size_t j=0; j<depth_width; j++) {
@@ -523,27 +523,45 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
                     tmp_buffer[ i*depth_width + j ] = m_buf[ ii*width + jj ];
                 }
             }
-
-            // Depth encoded as 24 bit fixed point values.
-            for (size_t i=0; i<depth_width*depth_height; i++) {
-                float value = (float)( tmp_buffer[i] );
-                for (size_t j=0; j<3; j++) {
-                    (buffer_pos)[3*i+j] = (unsigned char)( floor(value*255.0) );
+            if (depth16) {
+                // Depth encoded as 24 bit fixed point values, least significant bits set to 0
+                for (size_t i=0; i<depth_width*depth_height; i++) {
+                    float value = tmp_buffer[i];
+                    (buffer_pos)[3*i+0] = (unsigned char)( floor(value*255.0) );
                     value = 255.0*value - floor(value*255.0);
+                    (buffer_pos)[3*i+1] = (unsigned char)( floor(value*255.0) );
[…]

### 2014-12-01 src/trell/IPCJobController.cpp

@@ -87,6 +87,7 @@ IPCJobController::onGetSnapshot( char*               buffer,
                                  const size_t        height,
                                  const size_t        depth_width,
                                  const size_t        depth_height,
+                                 const bool          depth16,
                                  const std::string&  session,
                                  const std::string&  key )
 {
@@ -228,6 +229,7 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
         // These are coming from the url-parameters depth_w and depth_h...
         int depth_width = q->depth_h;
         int depth_height = q->depth_h;
+        // bool depth16 = q->depth16; @@@ should probably be gotten this way, too... for now, fetching from exposed model below...
 
         std::string key_list_string = std::string( q->viewer_key_list );
         std::vector<std::string> key_list;
@@ -271,11 +273,14 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
             return sizeof(tinia_msg_t);
         }
 
+        bool depth16;
+        m_model->getElementValue( "ap_16_bit_depth", depth16 );
+
         // Looping through all keys and grabbing GL-content
         char *buf = (char*)msg + sizeof(tinia_msg_image_t);
         for (size_t i=0; i<key_list.size(); i++) {
             key = key_list[i]; // Overriding the key gotten from the 'msg' parameter!
-            if ( onGetSnapshot(buf, format, w, h, depth_width, depth_height, session, key) ) { // onGetSnapshot() in IPCGLJobController grabs pixels for a given key
+            if ( onGetSnapshot(buf, format, w, h, depth_width, depth_height, depth16, session, key) ) { // onGetSnapshot() in IPCGLJobController grabs pixels for a given key
 #ifdef DEBUG
                 m_logger_callback( m_logger_data, 2, package.c_str(),
                                    "Queried for snapshot, ok. format = %d", format );

### 2014-12-01 src/utils/ProxyDebugGUI.cpp

@@ -122,7 +122,7 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addElement<bool>( "ap_mid_texel_sampling", false );              // false best
         model->addAnnotation("ap_mid_texel_sampling", "Sample mid-texel");
         model->addElement<bool>( "ap_use_qt_img_scaling", true );
-        model->addAnnotation("ap_use_qt_img_scaling", "Use QImage::scaled()");
+        model->addAnnotation("ap_use_qt_img_scaling", "Use QImage::scaled() (Qtc)");
         model->addElement<bool>( "ap_set_depth_size_32", false );
         model->addAnnotation("ap_set_depth_size_32", "32");
         model->addElement<bool>( "ap_set_depth_size_64", false );
@@ -140,9 +140,11 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addElement<bool>( "ap_mid_splat_sampling", false );
         model->addAnnotation("ap_mid_splat_sampling", "Sample mid-splat");
         model->addElement<bool>( "ap_simulate_downsampling", false );
-        model->addAnnotation("ap_simulate_downsampling", "Simulate downsampling");
+        model->addAnnotation("ap_simulate_downsampling", "Simulate downsampling (Qtc?)");
         model->addElement<bool>( "ap_bi_linear_filtering", false );
-        model->addAnnotation("ap_bi_linear_filtering", "Bi-linear filtering");
+        model->addAnnotation("ap_bi_linear_filtering", "Bi-linear filtering (Qtc)");
+        model->addElement<bool>( "ap_16_bit_depth", false );
+        model->addAnnotation("ap_16_bit_depth", "16 bit depth");
     }
 }
 
@@ -257,6 +259,7 @@ tinia::model::gui::Grid *ProxyDebugGUI::getGrid()
         mainGrid->setChild(row, 0, new tinia::model::gui::CheckBox("ap_use_qt_img_scaling"));
         mainGrid->setChild(row, 1, new tinia::model::gui::CheckBox("ap_mid_splat_sampling"));
         mainGrid->setChild(row, 2, new tinia::model::gui::CheckBox("ap_simulate_downsampling"));
+        mainGrid->setChild(row, 3, new tinia::model::gui::CheckBox("ap_16_bit_depth"));
         row++;
     }
 

### 2014-12-01 examples/cuberenderer/sockets_main.cpp
And now it works, need to have QApplication at the top...

@@ -24,16 +24,17 @@
 
 int main(int argc, char** argv)
 {
+    QCoreApplication blah(argc, argv );
+
     using namespace tinia::example;
     tinia::trell::FRVGLJobController *trellController = new tinia::trell::FRVGLJobController();
 
     CubeJob *testJob = new CubeJob();
     trellController->setJob(testJob);
-
     trellController->run(argc, argv);
-    tinia::trell::FRVQtController* server  = new tinia::trell::FRVQtController( trellController );
-
-    QCoreApplication blah(argc, argv );
+    
+    tinia::trell::FRVQtController* server  = new tinia::trell::FRVQtController( trellController, 9999, nullptr );
+   
     QObject::connect( server, &tinia::trell::FRVQtController::closed, &blah, &QCoreApplication::quit );
     auto retVal = blah.exec();
 

### 2014-12-01 include/tinia/trell/FRVQtController.hpp

@@ -15,14 +15,14 @@ class FRVQtController : public QObject
     Q_OBJECT
 public:
     //FRVQtController();
-    explicit FRVQtController( tinia::trell::FRVGLJobController* glJob, QObject* parent = Q_NULLPTR );
+    explicit FRVQtController( tinia::trell::FRVGLJobController* glJob, int port = 8097, QObject* parent = Q_NULLPTR );
     ~FRVQtController();
 
 Q_SIGNALS:
     void closed();
     
 private Q_SLOTS:
-     void onNewConnection();
+    void onNewConnection();
     void processTextMessage(QString message);
     void processBinaryMessage(QByteArray message);
     void socketDisconnected();

### 2014-12-01 src/trell/FRVQtController.cpp

@@ -5,11 +5,12 @@
 #include <QObject>
 
 
-tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, QObject* parent /*= Q_NULLPTR */ )
+tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, int port, QObject* parent )
     :
     QObject(parent),
     m_glJob( glJob ),
-    m_port(8081)
+    m_port(port),
+    m_clients()
 {
     m_pWebSocketServer = new QWebSocketServer(QStringLiteral("Echo Server"),
         QWebSocketServer::NonSecureMode, this);
@@ -42,12 +43,14 @@ tinia::trell::FRVQtController::~FRVQtController()
 
 void tinia::trell::FRVQtController::onNewConnection()
 {
+    qDebug() << "got a new connection attempt";
     QWebSocket *pSocket = m_pWebSocketServer->nextPendingConnection();
 
     connect(pSocket, &QWebSocket::textMessageReceived, this, &FRVQtController::processTextMessage);
     connect(pSocket, &QWebSocket::binaryMessageReceived, this, &FRVQtController::processBinaryMessage);
     connect(pSocket, &QWebSocket::disconnected, this, &FRVQtController::socketDisconnected);
-    m_glJob->render();
+    if(m_glJob)
+        m_glJob->render();
 
     m_clients << pSocket;
 
@@ -55,12 +58,16 @@ void tinia::trell::FRVQtController::onNewConnection()
 
 void tinia::trell::FRVQtController::processTextMessage(QString message)
 {
-     m_glJob->render();
+    qDebug() << "got a new text message";
+    if(m_glJob)
+        m_glJob->render();
 }
 
 void tinia::trell::FRVQtController::processBinaryMessage(QByteArray message)
 {
-    m_glJob->render();
+    qDebug() << "got a new binary message";
+    if(m_glJob)
+        m_glJob->render();
 }
 
 void tinia::trell::FRVQtController::socketDisconnected()

### 2014-12-02 static/demoes.js
Additional static page with nicely list of a few demoes. Initial sketch

@@ -1,0 +1,47 @@
+function addDemoJob( application )
+{
+    window.console.log("Starting addDemoJob with application " + application);
+    var req = new XMLHttpRequest();
+    req.open( "POST", "../master/rpc.xml?xml&foo=bar" );
+    req.setRequestHeader( "Content-Type", "text/xml" );
+    
+    req.onreadystatechange = function() {
+        if( this.readyState == 4 && this.status == 200 ) {
+            update( this );
+            window.console.log( "rpcMaster: Got: " + this.responseText );
+        }
+    };
+
+    var date = new Date();
+    var job_id = application + "_" + date.getTime();
+    
+    window.console.log( "rpcMaster: Sending: <here should there be xml...>"  );
+    req.send( '<?xml version="1.0"?>' +
+              '<query xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance" xmlns="http://cloudviz.sintef.no/trell/1.0">'+
+              '<addJob>' +
+              '<job>' + job_id + '</job>' +
+              '<application>' + application + '</application>' +   
+              '<renderingDeviceId>:0.0</renderingDeviceId>' +
+              '</addJob>' +
+              '</query>' );
+    
+    // Open job:
+    job_url = "../job/" + job_id + "/sessionid/index.html";
+    window.open(job_url, "_blank");
+}
+
+
+function demo_startClock()
+{
+    addDemoJob("clock_cloud_job");
+}
+
+function demo_startAP()
+{
+    addDemoJob("jpc_cloud_job");
+}
+
+function demo_startTwoCanvas()
+{
+    addDemoJob("two_cloud_job");
+}

### 2014-12-02 include/tinia/jobcontroller/OpenGLJob.hpp
Updated to work with FRView. Need to add GPB and image encoding

@@ -60,7 +60,13 @@ public:
            unsigned int fbo,
            const size_t        width,
            const size_t        height )=0;
-
+    
+    virtual
+    bool renderFrame( unsigned int        fbo,
+                      const size_t        width,
+                      const size_t        height,
+                      float*              modelView,
+                      float*              projection){ return true;};
 
 
     /** Let the controller access the render list for a particular session and key.

### 2014-12-02 src/trell/FRVGLJobController.cpp

@@ -70,7 +70,7 @@ FRVGLJobController::init()
     // --- OpenGL context created, init glew etc. ------------------------------
     glewInit();
 
-    GLDebugMessages::setupGLDebugMessages();
+    //GLDebugMessages::setupGLDebugMessages();
 
     
     GLint major, minor;
@@ -324,7 +324,7 @@ int FRVGLJobController::run( int argc, char** argv )
     //call glutMainloop.
     m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
     glutSwapBuffers();
-    render();
+
 
     return 1;
 }
@@ -334,12 +334,13 @@ void FRVGLJobController::render()
     static float rotation = 0.001;
     float prev = 0;
     tinia::example::CubeJob* cj = (tinia::example::CubeJob*) m_openGLJob;
-    for (int i = 0; i < 150; i++ )
+    for (int i = 0; i < 1; i++ )
     {
-        prev = cj->rotate( rotation * i );
+        prev = cj->rotate( rotation * i + 0.01 );
 
         m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
         glutSwapBuffers();
+
     }   
 }
 

### 2014-12-02 src/trell/FRVQtController.cpp

@@ -59,6 +59,13 @@ void tinia::trell::FRVQtController::onNewConnection()
 void tinia::trell::FRVQtController::processTextMessage(QString message)
 {
     qDebug() << "got a new text message";
+
+    QWebSocket *pClient = qobject_cast<QWebSocket *>(sender());
+
+    if ( pClient ) {
+        pClient->sendTextMessage("frv " + message);
+    }
+
     if(m_glJob)
         m_glJob->render();
 }

### 2014-12-02 tex/short_2015/body.tex
more stuff

@@
 \end{abstract}
+
+%-------------------------------------------------------------------------
+\section{Notes}
+
+\begin{itemize}
+
+\item We restrict ourselves to discuss ``splat sets'' in this paper. (Nevne der det
+tas frem andre alternativer)
+
+\item
+Husk aa faa med det viktige i at vi faar oekt serverkapasiteten til aa
+serve flere klienter ogsaa!
+
+\item
+``server-image'' i steden for ``server-rendered image'' for aa spare plass?
+
+\item received stavekontroll
+
+\item Hvor passer det aa faa inn 1) bilde og 2) tekst om FRView?
+
+\item Maa ta et lite soek paa prior art innen omraadet
+
+\end{itemize}
 
@@
 
-We compute $\sv_{dx}$ and $\sv_{dy}$ by sampling proxy model positions and
+% After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
+% After persp division: ndc [-1, 1]^3
+% After viewport transf: window coordinates
+% (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
+
+We compute $\sv_{dx}$ and $\sv_{dy}$ by evaluating proxy model positions and
 performing the perspective division and transformation into window coordinates
-in the vertex shader.
-Let
+in the vertex shader. Let proxy model positions corresponding to three corners
+of a quadrilateral splat in clip coordinates be
+given by
 \[
-  \text{pos} =
+  \pv =
   U { \small \begin{pmatrix} x \\ y \\ d \\ 1 \end{pmatrix} },
-  \text{pos}_{\Delta x} =
+  \pv_{x+\Delta x} =
   U { \small \begin{pmatrix} x+\Delta x \\ y \\ d_{\Delta x} \\ 1 \end{pmatrix} }
   \text{and}\, \, \, 
-  \text{pos}_{\Delta y} =
+  \pv_{y+\Delta y} =
   U { \small \begin{pmatrix} x \\ y+\Delta y \\ d_{\Delta y} \\ 1 \end{pmatrix} },
@@
 2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
-y+\Delta y) - 1$ are depths sampled from the received depth buffer and transformed
-to $[-1, 1]$.
+y+\Delta y) - 1$ are depths sampled from the received depth buffer and
+transformed to $[-1, 1]$. By combining this with a $w$-component set to one, we
+have in effect done a perspective division, so that the right-hand side vertices
+are in clip space, suitable for multiplication with $U$.
 
@@
 
-Then, for the 2D vectors spanning the splats in client window coordinates, we
-get
+\textbf{ Litt om valgene vi kan gjoere for Delta}
+
+This again leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
+coordinates, and after perspective division and viewport transformation, we get
+corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
+y}$, which we can subtract to obtain 2D vectors spanning the splat,
 \[
   \sv_{\Delta x} =
-    \frac{\text{viewport width}}{2\Delta x} \left(
-        \frac{\text{pos}_{\Delta x}.xy}{\text{pos}_{\Delta x}.w} -
-        \frac{\text{pos}.xy}{\text{pos}.w}
+  \sv_{x+\Delta x} - \sv =
+    \frac{\text{width}}{2\Delta x} \left(
+        \frac{\pv_{x+\Delta x}.xy}{\pv_{x+\Delta x}.w} -
+        \frac{\pv.xy}{\pv.w}
     \right),
@@
   \sv_{\Delta y} =
-    \frac{\text{viewport width}}{2\Delta y} \left(
-        \frac{\text{pos}_{\Delta y}.xy}{\text{pos}_{\Delta y}.w} -
-        \frac{\text{pos}.xy}{\text{pos}.w}
+  \sv_{y+\Delta y} - \sv =
+    \frac{\text{height}}{2\Delta y} \left(
+        \frac{\pv_{y+\Delta y}.xy}{\pv_{y+\Delta y}.w} -
+        \frac{\pv.xy}{\pv.w}
     \right),
@@
 where we have used the ``shader notation'' to extract 2D vectors and scalars
-from the 4D vectors.
+from the 4D vectors, and $\text{width}$ and $\text{height}$ are the viewport
+width and height, respectively.
 
@@
 
+The problem with this, is that the geometry is not planar, and our simple texture
+coordinate transformation will produce the wrong result. Some ways to remedy
+this, is to choose smaller (and more) splats, detect the problem and discard
+such splats, or introduce more complex texture transformations. Since the proxy
+geometry in principle is to be short-lived, and only a temporary geometry to be
+rendered before the server-rendered image can be receieved, we have chosen the
+first alternative. Discarding such splats may offer a good alternative,
+especially when combining several proxy geometries, where the union of retained
+splats increases the likelihood of all geometry parts being covered. Note also
+that implementing a more sophisticated texture coordinate transformation may
+actually amount to performing the same work as for more and smaller splats. Or,
+in other words, using more and smaller splats may be regarded as a better
+texture transform implementation.
+
 %-------------------------------------------------------------------------
@@
 
-%-------------------------------------------------------------------------
+For larger splats, covering many fragments, it makes sense to also compute and
+use depth fragments. This will improve the usage of more proxy models
+simultaneously, since the risk of a splat with a not very representative depth
+value covers another and better splat is reduced. The ``intra-splat'' depth
+values can easily be fetched from the depth image, just as the texture is looked
+up for color. One problem is that many thin clients may not support this, in
+WebGL, the ``frag depth'' feature is an extension, which may not be supported
+everywhere.
 
@@
 
-In the case of splat sets: Replacement algorithms
+In this paper we concern ourselves with proxy models defined as sets of
+splats. Since each splat set is the product of one instance of a server-rendered
+image, depth buffer and viewing parameters, it makes sense to retain more than
+one such model on the client. By doing this, the client may combine them, in
+order to provide renderings for viewing parameters very different from the last
+one received. One can imagine a plethora of splat set replacement algorithms, we
+have tested two approaches which both works well; both retain a constant number
+of proxy models. The first algorithm is to replace the one with a viewing
+direction differing the most from the newly received model. The second simply
+replaces the oldest one in store. The third replaces a proxy model $j$ if the
+replacement results in the following objective function being reduced,
+\[
+  \text{coverage}_j = 
+  \sum_{i=1, i\neq j}^n 
+    \angle(\textbf{camdir}_i, \textbf{new camdir} )^2,
+\]
+where $\textbf{camdir}$ is the direction in which the camera was looking for the
+generation of that particular proxy model.
+
+There are some issues to consider with respect to panning and zooming that we
+will not discuss here due to the limited space available.  All of these model
+replacement algorithms work decently well.
 
@@
 
+Since the main purpose of the automatically generated proxy model is to
+facilitate a client-rendered image while waiting for one from the server for
+increased interactivity, it is important that the process of generating and
+sending proxy model data from the server itself is not hold up more than
+necessary. There are mainly three sources of delay (\textbf{passer aa nevne
+tidligere?}) for server-rendered images to the client; high
+latency, low bandwidth, and slow server-rendering itself.
+
+In the first case, it seems prudent to have a better proxy model on the client,
+that can be used for longer time and for interactivity resulting in viewing
+parameters deviating more from the last received server-rendered image. In the
+two other cases, it is important for the proxy model generation/transmission to
+be quicker, both in order to get the proxy model to the client and keep from
+delaying the server-image more than necessary.
+
+To take care of this, we have adopted an adaptive specification of
+proxy model data from the server. This involves a more light-weight image (lossy
+JPG-compression with adaptive quality control) while the user is interacting
+with the client, as well as a reduced resolution of the depth buffer sent from
+the server. See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
+high-quality proxy models.
+
+\begin{figure}[htb]
+  \centering
+  \subfigure[Low quality (tenker her jpg med q=0 og daarlig dybdebuffer, faa
+  splats kanskje]{
+    \input{fig3a.tex}
+  }
+  \subfigure[High quality (png med maks av alt, mange splats)]{
+    \input{fig3b.tex}
+  }
+  \caption{\label{fig:adapativeProxyModels}
+           Proxy models adaptively adjusted to available bandwidth and server
+           capacity.}
+\end{figure}
+
 %-------------------------------------------------------------------------
 \section{Results and discussion}
+
+We have tested the algorithms discussed in the Tinia framework, which is a
+programming framework for setting up and managing a client/server based
+interactive OpenGL-based visualization application. The client is
+Javascript/WebGL-based and runs in one of several possible browsers, \eg Chrome
+or Firefox. This client communicates over tcp/ip, using either http or
+websockets, with a server running as an Apache module.
+
+The automatic proxy geometry implementation is fully invisible to the
+application, meaning that all existing applications immediately will have the
+feature available. We have tested the algorithm on several smaller test cases,
+but also on a larger oild reservoir viewer.
+
+This viewer utilizes several GLSL shaders to render reservoir cells and
+boundaries, tubular wells etc., and operates in general on a geometric model
+that is not trivial to reduce in complexity. The model is typically also very
+large, meaning that rendering on a thin client is prohibitive. With the
+automatic proxy geometry, we can obtain interactive frame rates with a bad
+connection from a very lightweight client to the server. For a comparison of a
+server-rendered image and a client-rendered proxy model, see
+Figure~\ref{fig:FRView}.
+
+\begin{figure}[htb]
+  \centering
+  \subfigure[Server-rendered]{
+    \input{fig3a.tex}
+  }
+  \subfigure[Client-rendered proxy model]{
+    \input{fig3b.tex}
+  }
+  \caption{\label{fig:FRView}
+           The oil reservoir viewer FRView}
+\end{figure}
 
@@
 
-%-------------------------------------------------------------------------
-\subsection{References}
+The single most attractive feature of such an automatic proxy geometry as
+described, is exactly the automated generation. It means that problems with
+downscaling, sending and rendering existing geometries is bypassed
+altogether. The Tinia framework with which we have demonstrated this, is a
+framework for setting up and managing a client/server based interactive OpenGL
+based visualization application. The automatic proxy geometry implementation is
+fully invisible to the application, meaning that all existing applications
+immediately will have the feature available.
 
-List all bibliographical references in 9-point Times, single-spaced, at the
-end of your paper in alphabetical order. When referenced in the text, enclose
-the citation index in square brackets, for example~\cite{Lous90}. Where
-appropriate, include the name(s) of editors of referenced books.
-
-For your references please use the following algorithm:
+There are several directions in which we would like to follow up and improve
+this concept,
 \begin{itemize}
-\item \textbf{one} author: first 3 chars plus year -- 
-      e.g.\ \cite{Lous90}
-\item \textbf{two}, \textbf{three} or \textbf{four} authors: first char
-      of each family name plus year --  e.g.\ \cite{Fellner-Helmberg93} 
-      or \cite{Kobbelt97-USHDR} or \cite{Lafortune97-NARF}
-\item \textbf{more than 4} authors: first char of family name from 
-      first 3 authors followed by a '*' followed by the year -- 
-      e.g.\ \cite{Buhmann:1998:DCQ} or \cite{FolDamFeiHug.etal93} 
+\item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
+all relevant viewing parameters on the client, with the help of a minimal set of
+proxy models.
+\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
+envision a spatially partioned grid, for instance containing a distance field,
+with the geometry as an iso-surface.
 \end{itemize}
-
-For BibTeX users a style file \ \texttt{eg-alpha.bst} \ is available which
-uses the above algorithm.
-
-%-------------------------------------------------------------------------
-\subsection{Conclusions}
-
-Please direct any questions to the production editor in charge of
-these proceedings.
 

### 2014-12-02 examples/cuberenderer/CubeJob.cpp
Started work on gpb imageRequest and imageResponse. Not very pretty hacks to start the grabbing and encoding of pixels.

@@ -153,6 +153,79 @@ bool CubeJob::renderFrame(const std::string &session, const std::string &key, un
     return true;
 }
 
+bool CubeJob::renderFrame( unsigned int fbo, const size_t width, const size_t height, float* modelView, float* projection)
+{
+    glEnable(GL_DEPTH_TEST);
+
+    glClearColor(0, 0, 0, 1);
+    glClear(GL_COLOR_BUFFER_BIT|GL_DEPTH_BUFFER_BIT);
+
+    glViewport(0, 0, width, height);
+    // tinia::model::Viewer viewer;
+    //m_model->getElementValue( key, viewer);
+    glLoadIdentity();
+    glMatrixMode(GL_PROJECTION);
+    glLoadIdentity();
+    //glLoadMatrixf( viewer.projectionMatrix.data() );
+    glMatrixMode(GL_MODELVIEW);
+    //glLoadMatrixf( viewer.modelviewMatrix.data() );
+    glLoadIdentity();
+    glRotatef( m_rotation * 57.2957795, 0.001, 0.0, 1.0 );
+
+    glBegin(GL_POLYGON);
+    glColor3f(   1.0,  0.0, 0.0 );
+    glVertex3f(  0.5, -0.5, -0.5 );
+    glVertex3f(  0.5,  0.5, -0.5 );
+    glVertex3f( -0.5,  0.5, -0.5 );
+    glVertex3f( -0.5, -0.5, -0.5 );
+    glEnd();
+
+    glBegin(GL_POLYGON);
+    glColor3f(   0.0,  1.0, 0.0 );
+    glVertex3f(  0.5, -0.5, 0.5 );
+    glVertex3f(  0.5,  0.5, 0.5 );
+    glVertex3f( -0.5,  0.5, 0.5 );
+    glVertex3f( -0.5, -0.5, 0.5 );
+    glEnd();
+
+    glBegin(GL_POLYGON);
+    glColor3f(  .0,  0.0,  1.0 );
+    glVertex3f( 0.5, -0.5, -0.5 );
+    glVertex3f( 0.5,  0.5, -0.5 );
+    glVertex3f( 0.5,  0.5,  0.5 );
+    glVertex3f( 0.5, -0.5,  0.5 );
+    glEnd();
+
+    glBegin(GL_POLYGON);
+    glColor3f(   1.0,  1.0,  0.0 );
[…]

### 2014-12-02 examples/cuberenderer/CubeJob.hpp

@@ -33,6 +33,8 @@ public:
     bool renderFrame(const std::string &session, const std::string &key,
                      unsigned int fbo, const size_t width, const size_t height);
 
+    bool renderFrame(unsigned int fbo, const size_t width, const size_t height, float* modelView, float* projection);
+
     /** Database for renderlist.
       *
       * Desktop jobs doesn't usually provide render lists (as they are normally

### 2014-12-02 include/tinia/trell/FRVGLJobController.hpp

@@ -53,7 +53,7 @@ public:
     init( );
 
 
-    void render();
+    char* render( float* modelView = nullptr, float* projection = nullptr );
 
 
 protected:

### 2014-12-02 include/tinia/trell/FRVProtoBuffers.pb.hpp

@@ -1,0 +1,542 @@
+// Generated by the protocol buffer compiler.  DO NOT EDIT!
+// source: FRVProtoBuffers.proto
+
+#ifndef PROTOBUF_FRVProtoBuffers_2eproto__INCLUDED
+#define PROTOBUF_FRVProtoBuffers_2eproto__INCLUDED
+
+#include <string>
+
+#include <google/protobuf/stubs/common.h>
+
+#if GOOGLE_PROTOBUF_VERSION < 2006000
+#error This file was generated by a newer version of protoc which is
+#error incompatible with your Protocol Buffer headers.  Please update
+#error your headers.
+#endif
+#if 2006000 < GOOGLE_PROTOBUF_MIN_PROTOC_VERSION
+#error This file was generated by an older version of protoc which is
+#error incompatible with your Protocol Buffer headers.  Please
+#error regenerate this file with a newer version of protoc.
+#endif
+
+#include <google/protobuf/generated_message_util.h>
+#include <google/protobuf/message.h>
+#include <google/protobuf/repeated_field.h>
+#include <google/protobuf/extension_set.h>
+#include <google/protobuf/unknown_field_set.h>
+// @@protoc_insertion_point(includes)
+
+namespace frv {
+
+// Internal implementation detail -- do not call these.
+void  protobuf_AddDesc_FRVProtoBuffers_2eproto();
+void protobuf_AssignDesc_FRVProtoBuffers_2eproto();
+void protobuf_ShutdownFile_FRVProtoBuffers_2eproto();
+
+class imageRequest;
+class imageResponse;
+
+// ===================================================================
+
+class imageRequest : public ::google::protobuf::Message {
+ public:
+  imageRequest();
+  virtual ~imageRequest();
+
+  imageRequest(const imageRequest& from);
+
+  inline imageRequest& operator=(const imageRequest& from) {
+    CopyFrom(from);
+    return *this;
+  }
+
+  inline const ::google::protobuf::UnknownFieldSet& unknown_fields() const {
+    return _unknown_fields_;
+  }
+
+  inline ::google::protobuf::UnknownFieldSet* mutable_unknown_fields() {
+    return &_unknown_fields_;
+  }
+
+  static const ::google::protobuf::Descriptor* descriptor();
+  static const imageRequest& default_instance();
+
+  void Swap(imageRequest* other);
+
+  // implements Message ----------------------------------------------
+
+  imageRequest* New() const;
+  void CopyFrom(const ::google::protobuf::Message& from);
[…]

### 2014-12-02 src/trell/FRVGLJobController.cpp

@@ -66,11 +66,12 @@ FRVGLJobController::init()
     glutInitWindowPosition (1000, 100);
     glutInitContextFlags( GLUT_DEBUG );
     glutCreateWindow ("FRViewJobDebugWindow");    
-    
+    m_width = 1000;
+    m_height = 1000; 
     // --- OpenGL context created, init glew etc. ------------------------------
     glewInit();
 
-    //GLDebugMessages::setupGLDebugMessages();
+    GLDebugMessages::setupGLDebugMessages();
 
     
     GLint major, minor;
@@ -329,8 +330,16 @@ int FRVGLJobController::run( int argc, char** argv )
     return 1;
 }
 
-void FRVGLJobController::render()
+char* FRVGLJobController::render(float* modelView, float* projection )
 {    
+    int rgb = 3;
+    char* buffer;
+    if( modelView != nullptr || projection != nullptr){
+         buffer = (char*)malloc( m_width * m_height * rgb * sizeof(float) );
+    }
+
+    glBindFramebuffer( GL_FRAMEBUFFER, m_fbo );
+    glViewport( 0, 0, m_width, m_height );
     static float rotation = 0.001;
     float prev = 0;
     tinia::example::CubeJob* cj = (tinia::example::CubeJob*) m_openGLJob;
@@ -338,10 +347,14 @@ void FRVGLJobController::render()
     {
         prev = cj->rotate( rotation * i + 0.01 );
 
-        m_openGLJob->renderFrame( "blah", "viewer", 0, 512, 512 );
+        m_openGLJob->renderFrame( 0,  m_width, m_height, modelView, projection );
         glutSwapBuffers();
 
     }   
+   // glPixelStorei( GL_PACK_ALIGNMENT, 1 );
+
+    //glReadPixels( 0, 0, m_width, m_height, GL_RGB, GL_UNSIGNED_BYTE, buffer );
+    return buffer;
 }
 
 

### 2014-12-02 src/trell/FRVProtoBuffers.pb.cpp

@@ -1,0 +1,802 @@
+// Generated by the protocol buffer compiler.  DO NOT EDIT!
+// source: FRVProtoBuffers.proto
+
+#define INTERNAL_SUPPRESS_PROTOBUF_FIELD_DEPRECATION
+#include "tinia/trell/FRVProtoBuffers.pb.hpp"
+
+#include <algorithm>
+
+#include <google/protobuf/stubs/common.h>
+#include <google/protobuf/stubs/once.h>
+#include <google/protobuf/io/coded_stream.h>
+#include <google/protobuf/wire_format_lite_inl.h>
+#include <google/protobuf/descriptor.h>
+#include <google/protobuf/generated_message_reflection.h>
+#include <google/protobuf/reflection_ops.h>
+#include <google/protobuf/wire_format.h>
+// @@protoc_insertion_point(includes)
+
+namespace frv {
+
+namespace {
+
+const ::google::protobuf::Descriptor* imageRequest_descriptor_ = NULL;
+const ::google::protobuf::internal::GeneratedMessageReflection*
+  imageRequest_reflection_ = NULL;
+const ::google::protobuf::Descriptor* imageResponse_descriptor_ = NULL;
+const ::google::protobuf::internal::GeneratedMessageReflection*
+  imageResponse_reflection_ = NULL;
+
+}  // namespace
+
+
+void protobuf_AssignDesc_FRVProtoBuffers_2eproto() {
+  protobuf_AddDesc_FRVProtoBuffers_2eproto();
+  const ::google::protobuf::FileDescriptor* file =
+    ::google::protobuf::DescriptorPool::generated_pool()->FindFileByName(
+      "FRVProtoBuffers.proto");
+  GOOGLE_CHECK(file != NULL);
+  imageRequest_descriptor_ = file->message_type(0);
+  static const int imageRequest_offsets_[3] = {
+    GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageRequest, modelview_),
+    GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageRequest, projection_),
+    GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageRequest, zscale_),
+  };
+  imageRequest_reflection_ =
+    new ::google::protobuf::internal::GeneratedMessageReflection(
+      imageRequest_descriptor_,
+      imageRequest::default_instance_,
+      imageRequest_offsets_,
[…]

### 2014-12-02 src/trell/FRVQtController.cpp

@@ -4,6 +4,7 @@
 #include <QtWebSockets/QWebSocketServer>
 #include <QObject>
 
+#include "tinia/trell/FRVProtoBuffers.pb.h"
 
 tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, int port, QObject* parent )
     :
@@ -72,9 +73,26 @@ void tinia::trell::FRVQtController::processTextMessage(QString message)
 
 void tinia::trell::FRVQtController::processBinaryMessage(QByteArray message)
 {
+    QWebSocket *pClient = qobject_cast<QWebSocket *>(sender());
     qDebug() << "got a new binary message";
-    if(m_glJob)
-        m_glJob->render();
+    if( !pClient || &m_glJob){
+        qDebug() << "Major problem, we are not ready for messages yet. Please try again later!";
+        return;
+    }
+    
+   
+
+    frv::imageRequest ir;
+    frv::imageResponse response;
+    ir.ParseFromString( (std::string)message );
+    if( ir.has_zscale() ){
+        qDebug() << "it has z-scale, something got through and that was: " << ir.zscale();
+    }
+
+    char* results =  m_glJob->render();
+    response.set_allocated_image_bytes( (std::string*)results );
+    pClient->sendBinaryMessage( response.SerializeAsString().c_str() );
+    
 }
 
 void tinia::trell::FRVQtController::socketDisconnected()

### 2014-12-03 examples/cuberenderer/CubeJob.cpp
Now renders and transmits image over GPB as png. Requires qt plugin to support JPEG. Only tested with nullptr for modelView and projection in example4.

@@ -99,7 +99,7 @@ bool CubeJob::renderFrame(const std::string &session, const std::string &key, un
     glMatrixMode(GL_MODELVIEW);
     //glLoadMatrixf( viewer.modelviewMatrix.data() );
     glLoadIdentity();
-    glRotatef( m_rotation * 57.2957795, 0.001, 0.0, 1.0 );
+    glRotatef( m_rotation * 57.2957795, 0.1, 0.0, 1.0 );
 
     glBegin(GL_POLYGON);
     glColor3f(   1.0,  0.0, 0.0 );

### 2014-12-03 include/tinia/trell/FRVGLJobController.hpp

@@ -52,8 +52,10 @@ public:
     bool
     init( );
 
+    int getWidth() const { return m_width; }
+    int getHeight() const { return m_height; }
 
-    char* render( float* modelView = nullptr, float* projection = nullptr );
+    unsigned char* render( float* modelView = nullptr, float* projection = nullptr );
 
 
 protected:

### 2014-12-03 include/tinia/trell/FRVProtoBuffers.pb.hpp

@@ -210,17 +210,17 @@ class imageResponse : public ::google::protobuf::Message {
   inline ::std::string* release_image_bytes();
   inline void set_allocated_image_bytes(::std::string* image_bytes);
 
-  // required string imamge_type = 2;
-  inline bool has_imamge_type() const;
-  inline void clear_imamge_type();
-  static const int kImamgeTypeFieldNumber = 2;
-  inline const ::std::string& imamge_type() const;
-  inline void set_imamge_type(const ::std::string& value);
-  inline void set_imamge_type(const char* value);
-  inline void set_imamge_type(const char* value, size_t size);
-  inline ::std::string* mutable_imamge_type();
-  inline ::std::string* release_imamge_type();
-  inline void set_allocated_imamge_type(::std::string* imamge_type);
+  // required string image_type = 2;
+  inline bool has_image_type() const;
+  inline void clear_image_type();
+  static const int kImageTypeFieldNumber = 2;
+  inline const ::std::string& image_type() const;
+  inline void set_image_type(const ::std::string& value);
+  inline void set_image_type(const char* value);
+  inline void set_image_type(const char* value, size_t size);
+  inline ::std::string* mutable_image_type();
+  inline ::std::string* release_image_type();
+  inline void set_allocated_image_type(::std::string* image_type);
 
   // optional float zScale = 3;
   inline bool has_zscale() const;
@@ -233,8 +233,8 @@ class imageResponse : public ::google::protobuf::Message {
  private:
   inline void set_has_image_bytes();
   inline void clear_has_image_bytes();
-  inline void set_has_imamge_type();
-  inline void clear_has_imamge_type();
+  inline void set_has_image_type();
+  inline void clear_has_image_type();
   inline void set_has_zscale();
   inline void clear_has_zscale();
 
@@ -243,7 +243,7 @@ class imageResponse : public ::google::protobuf::Message {
   ::google::protobuf::uint32 _has_bits_[1];
   mutable int _cached_size_;
   ::std::string* image_bytes_;
-  ::std::string* imamge_type_;
+  ::std::string* image_type_;
   float zscale_;
   friend void  protobuf_AddDesc_FRVProtoBuffers_2eproto();
   friend void protobuf_AssignDesc_FRVProtoBuffers_2eproto();
[…]

### 2014-12-03 src/trell/FRVGLJobController.cpp

@@ -330,13 +330,13 @@ int FRVGLJobController::run( int argc, char** argv )
     return 1;
 }
 
-char* FRVGLJobController::render(float* modelView, float* projection )
+unsigned char* FRVGLJobController::render(float* modelView, float* projection )
 {    
     int rgb = 3;
-    char* buffer;
-    if( modelView != nullptr || projection != nullptr){
-         buffer = (char*)malloc( m_width * m_height * rgb * sizeof(float) );
-    }
+    unsigned char* buffer;
+    //if( modelView != nullptr || projection != nullptr){
+         buffer = (unsigned char*)malloc( m_width * m_height * rgb * sizeof(float) );
+   // }
 
     glBindFramebuffer( GL_FRAMEBUFFER, m_fbo );
     glViewport( 0, 0, m_width, m_height );
@@ -348,12 +348,14 @@ char* FRVGLJobController::render(float* modelView, float* projection )
         prev = cj->rotate( rotation * i + 0.01 );
 
         m_openGLJob->renderFrame( 0,  m_width, m_height, modelView, projection );
-        glutSwapBuffers();
+        //glutSwapBuffers();
 
     }   
    // glPixelStorei( GL_PACK_ALIGNMENT, 1 );
-
-    //glReadPixels( 0, 0, m_width, m_height, GL_RGB, GL_UNSIGNED_BYTE, buffer );
+    
+    
+    glReadPixels( 0, 0, m_width, m_height, GL_RGB, GL_UNSIGNED_BYTE, buffer );
+    glutSwapBuffers();
     return buffer;
 }
 

### 2014-12-03 src/trell/FRVProtoBuffers.pb.cpp

@@ -56,7 +56,7 @@ void protobuf_AssignDesc_FRVProtoBuffers_2eproto() {
   imageResponse_descriptor_ = file->message_type(1);
   static const int imageResponse_offsets_[3] = {
     GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageResponse, image_bytes_),
-    GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageResponse, imamge_type_),
+    GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageResponse, image_type_),
     GOOGLE_PROTOBUF_GENERATED_MESSAGE_FIELD_OFFSET(imageResponse, zscale_),
   };
   imageResponse_reflection_ =
@@ -106,9 +106,9 @@ void protobuf_AddDesc_FRVProtoBuffers_2eproto() {
   ::google::protobuf::DescriptorPool::InternalAddGeneratedFile(
     "\n\025FRVProtoBuffers.proto\022\003frv\"M\n\014imageReq"
     "uest\022\025\n\tmodelView\030\001 \003(\002B\002\020\001\022\026\n\nprojectio"
-    "n\030\002 \003(\002B\002\020\001\022\016\n\006zScale\030\003 \001(\002\"I\n\rimageResp"
-    "onse\022\023\n\013image_bytes\030\001 \002(\014\022\023\n\013imamge_type"
-    "\030\002 \002(\t\022\016\n\006zScale\030\003 \001(\002", 182);
+    "n\030\002 \003(\002B\002\020\001\022\016\n\006zScale\030\003 \001(\002\"H\n\rimageResp"
+    "onse\022\023\n\013image_bytes\030\001 \002(\014\022\022\n\nimage_type\030"
+    "\002 \002(\t\022\016\n\006zScale\030\003 \001(\002", 181);
   ::google::protobuf::MessageFactory::InternalRegisterGeneratedFile(
     "FRVProtoBuffers.proto", &protobuf_RegisterTypes);
   imageRequest::default_instance_ = new imageRequest();
@@ -469,7 +469,7 @@ void imageRequest::Swap(imageRequest* other) {
 
 #ifndef _MSC_VER
 const int imageResponse::kImageBytesFieldNumber;
-const int imageResponse::kImamgeTypeFieldNumber;
+const int imageResponse::kImageTypeFieldNumber;
[…]

### 2014-12-03 src/trell/FRVQtController.cpp

@@ -3,8 +3,10 @@
 #include <QtWebSockets/QWebSocket>
 #include <QtWebSockets/QWebSocketServer>
 #include <QObject>
-
-#include "tinia/trell/FRVProtoBuffers.pb.h"
+#include <QImage>
+#include <QBuffer>
+#include <QIODevice>
+#include "tinia/trell/FRVProtoBuffers.pb.hpp"
 
 tinia::trell::FRVQtController::FRVQtController( FRVGLJobController* glJob, int port, QObject* parent )
     :
@@ -50,8 +52,6 @@ void tinia::trell::FRVQtController::onNewConnection()
     connect(pSocket, &QWebSocket::textMessageReceived, this, &FRVQtController::processTextMessage);
     connect(pSocket, &QWebSocket::binaryMessageReceived, this, &FRVQtController::processBinaryMessage);
     connect(pSocket, &QWebSocket::disconnected, this, &FRVQtController::socketDisconnected);
-    if(m_glJob)
-        m_glJob->render();
 
     m_clients << pSocket;
 
@@ -63,12 +63,30 @@ void tinia::trell::FRVQtController::processTextMessage(QString message)
 
     QWebSocket *pClient = qobject_cast<QWebSocket *>(sender());
 
-    if ( pClient ) {
-        pClient->sendTextMessage("frv " + message);
-    }
+    //if ( pClient ) {
+    //    pClient->sendTextMessage("frv " + message);
+    //}
+    //
+    unsigned char* result  = m_glJob->render();
 
-    if(m_glJob)
-        m_glJob->render();
+    //convert to JPEG
+    std::string imageType = "PNG";
+    QImage qi( result, m_glJob->getWidth(), m_glJob->getHeight(), QImage::Format_RGB888 );
+    qi = qi.mirrored();
+    QByteArray qba;
+    QBuffer buffer(&qba);
+    buffer.open(QIODevice::WriteOnly);
+    qi.save( &buffer, imageType.c_str());
+    buffer.close();
+    std::string payload( qba.constData(), qba.size());
+    //payload = new std::string
+    frv::imageResponse *response = new frv::imageResponse();
+    response->set_allocated_image_type(&imageType);
+    response->set_allocated_image_bytes( &payload );
+    QByteArray responseBytes (response->SerializeAsString().c_str(), response->ByteSize() );
+    pClient->sendBinaryMessage( responseBytes );
+    free(result);
+ 
 }
 
[…]

### 2014-12-03 js/gui/Canvas.js
More GUI elements for debug-control. 512 to 1024. Jpg works again. Background can be set. 'hold png' got a toggle.

@@ -51,8 +51,8 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
         this._boundingboxKey = params.boundingboxKey;
         this._resetViewKey = params.resetViewKey;
         this._renderListURL = params.renderListURL;
-        this._width = 512;
-        this._height = 512;
+        this._width = 1024;
+        this._height = 1024;
         this._modelLib = params.modelLib;
         // The modification of fields of 'params' in this constructor is probably not necessary, because the call (there seems to be
         // only one) to the constructor uses a very short-lived automatic variable that is not used again before going out of scope.
@@ -206,8 +206,8 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                 var depthheight = response_obj[this._key].depthheight;
                                 // console.log("Received depth size: " + depthwidth + " " + depthheight + ", current model's depth size: " +
                                 //             this._modelLib.getElementValue("ap_depthWidth") + " " + this._modelLib.getElementValue("ap_depthHeight"));
-                                if ( ( (depthwidth ==this._modelLib.getElementValue("ap_depthWidth" )) || (depthwidth ==0) ) &&
-                                     ( (depthheight==this._modelLib.getElementValue("ap_depthHeight")) || (depthheight==0) ) )
+                                if ( ( (depthwidth ==this._modelLib.getElementValue("ap_depthWidth" )) || (depthwidth ==0) || (depthwidth ===undefined) ) &&
+                                     ( (depthheight==this._modelLib.getElementValue("ap_depthHeight")) || (depthheight==0) || (depthheight===undefined) ) )
                                 {
                                     // Not completely sure, but it may be a good idea to not update with a received bundle, if the depth size does not match what the shader is told...
                                     // Currently, the shader gets the macros DEPTH_* from these exposed model elements.
@@ -221,7 +221,9 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                                     this._snapshotTimings.print();
                                     this._autoSelectSnapshotType(this._snapshotTimings);
                                 } else {
-                                    console.log("Depth size of received bundle does not match what the shader has been told to expect. Ignoring this bundle.");
+                                    console.log("Depth size of received bundle does not match what the shader has been told to expect. Ignoring this bundle. (1)");
+                                    console.log("depthwidth=" + depthwidth + ", depthheight=" + depthheight +
[…]

### 2014-12-03 js/gui/ProxyModel.js

@@ -57,7 +57,7 @@ dojo.declare("gui.ProxyModel", null, {
             this._gl.texParameteri(this._gl.TEXTURE_2D, this._gl.TEXTURE_MAG_FILTER, this._gl.NEAREST);
             this._gl.texParameteri(this._gl.TEXTURE_2D, this._gl.TEXTURE_MIN_FILTER, this._gl.NEAREST);
             this._gl.bindTexture(this._gl.TEXTURE_2D, null);
-            console.log("Depth image (" + depthBufferAsText.length + " bytes) loaded in " + (Date.now()-depth_t0) + " ms. (" + Math.floor((depthBufferAsText.length/(Date.now()-depth_t0))) + " bytes/ms)");
+            // console.log("Depth image (" + depthBufferAsText.length + " bytes) loaded in " + (Date.now()-depth_t0) + " ms. (" + Math.floor((depthBufferAsText.length/(Date.now()-depth_t0))) + " bytes/ms)");
             imagesLoaded = imagesLoaded + 1;
             if (imagesLoaded==2) {
                 this.projection         = projMatAsText.split(/ /);

### 2014-12-03 js/gui/ProxyRenderer.js

@@ -51,7 +51,7 @@ dojo.declare("gui.ProxyRenderer", null, {
                                                                         (180.0/1) / 180.0*3.1415926535,     // Is this a sensible value? 180/#models degrees
                                                                         1.1);                               // "Zoom threshold"
             break;
-        case "3) ReplaceOldestWhenDifferent-5":
+        case "3) ReplOldestWhnDiff-5":
         case 3:
             this._proxyModelCoverage = new gui.ProxyModelCoverageReplaceOldestWhenDifferent(glContext,
                                                                         5,                                  // Number of proxy models to keep
@@ -175,6 +175,29 @@ dojo.declare("gui.ProxyRenderer", null, {
         // 140911: Allowing these listeners even without debugging mode being enabled, so that applications may set their own values.
         //         Not adding a test for construction-time setting of all of these, just a subset.
         if (true) { // if (this._debugging) {
+            //-------------------------------------------------------
+            this.exposedModel.addLocalListener("ap_whitebg", dojo.hitch(this, function(event) {
+                if (this.exposedModel.getElementValue("ap_whitebg")) {
+                    this._backgroundCol = vec3.createFrom(1.0, 1.0, 1.0);
+                } else {
+                    this.exposedModel.updateElement("ap_whitebg", false);
+                }
+            }));
+            this.exposedModel.addLocalListener("ap_blackbg", dojo.hitch(this, function(event) {
+                if (this.exposedModel.getElementValue("ap_blackbg")) {
+                    this._backgroundCol = vec3.createFrom(0.0, 0.0, 0.0);
+                } else {
+                    this.exposedModel.updateElement("ap_blackbg", false);
+                }
+            }));
+            this.exposedModel.addLocalListener("ap_greenbg", dojo.hitch(this, function(event) {
+                if (this.exposedModel.getElementValue("ap_greenbg")) {
+                    this._backgroundCol = vec3.createFrom(0.0, 0.2, 0.0);
+                } else {
+                    this.exposedModel.updateElement("ap_greenbg", false);
+                }
+            }));
+
             //-------------------------------------------------------
             this.exposedModel.addLocalListener("ap_set_depth_size_32", dojo.hitch(this, function(event) {
                 if(this.exposedModel.getElementValue("ap_set_depth_size_32")) {
[…]

### 2014-12-03 src/utils/ProxyDebugGUI.cpp

@@ -30,7 +30,7 @@ namespace {
     const char *allowed_auto_proxy_algos[] = { "0) AngleCoverage-5",
                                                "1) AngleCoverage-2",
                                                "2) OnlyMostRecent",
-                                               "3) ReplaceOldestWhenDifferent-5",
+                                               "3) ReplOldestWhnDiff-5",
                                                "4) ReplaceOldest-5",
                                                NULL };
 
@@ -54,7 +54,7 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
     if ( m_w_ap && m_w_jpg && m_w_as ) {
         model->addElement<bool>( "ap_autoSelect", false );            // Selects whatever proxy method works fastest.
         model->addElement<bool>( "ap_autoSelectSampleAll", false );
-        model->addAnnotation("ap_autoSelectSampleAll", "Smple all snapsh. types");
+        model->addAnnotation("ap_autoSelectSampleAll", "Sample all");
         model->addElement( "ap_autoSelectIndicator", "---" );
         model->addConstrainedElement<int>("ap_autoSelectTargetTime", 100, 0, 200);
         model->addAnnotation("ap_autoSelectTargetTime", "Target time:");
@@ -84,11 +84,11 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
             algos++;
         }
         model->addElementWithRestriction<std::string>( "ap_autoProxyAlgo", allowed_auto_proxy_algos[2], &allowed_auto_proxy_algos[0], &allowed_auto_proxy_algos[0]+algos );
-        model->addAnnotation("ap_autoProxyAlgo", "Proxy model replacement algo");
+        model->addAnnotation("ap_autoProxyAlgo", "Proxy replacement algo");
         model->addElement<bool>( "ap_debugSplatCol", false );
-        model->addAnnotation("ap_debugSplatCol", "Index coloring (r, g, b, y, c, m)");
+        model->addAnnotation("ap_debugSplatCol", "Index coloring"); //  (r, g, b, y, c, m)");
         model->addElement<bool>( "ap_decayMode", false );
-        model->addAnnotation("ap_decayMode", "Splats decaying from center");
[…]

### 2014-12-03 js/gui/ProxyModel.js
Commented out annoying dbg msg

@@ -71,7 +71,7 @@ dojo.declare("gui.ProxyModel", null, {
             }
         });
         depth_t0 = Date.now();
-        console.log("Starting depth image loading");
+        // console.log("Starting depth image loading");
         depthImage.src = "data:image/png;base64," + depthBufferAsText;
 
         var rgbImage = new Image();

### 2014-12-03 examples/cuberenderer/CubeJob.cpp
Changed rotation, not gone over to offscreen rendering yet.

@@ -170,7 +170,7 @@ bool CubeJob::renderFrame( unsigned int fbo, const size_t width, const size_t he
     glMatrixMode(GL_MODELVIEW);
     //glLoadMatrixf( viewer.modelviewMatrix.data() );
     glLoadIdentity();
-    glRotatef( m_rotation * 57.2957795, 0.001, 0.0, 1.0 );
+    glRotatef( m_rotation * 57.2957795, 0.25, 1.0, 0.2 );
 
     glBegin(GL_POLYGON);
     glColor3f(   1.0,  0.0, 0.0 );
@@ -181,7 +181,7 @@ bool CubeJob::renderFrame( unsigned int fbo, const size_t width, const size_t he
     glEnd();
 
     glBegin(GL_POLYGON);
-    glColor3f(   0.0,  1.0, 0.0 );
+    glColor3f(   1.0,  1.0, 1.0 );
     glVertex3f(  0.5, -0.5, 0.5 );
     glVertex3f(  0.5,  0.5, 0.5 );
     glVertex3f( -0.5,  0.5, 0.5 );

### 2014-12-03 src/trell/FRVGLJobController.cpp

@@ -65,7 +65,8 @@ FRVGLJobController::init()
     glutInitWindowSize (1000, 1000);
     glutInitWindowPosition (1000, 100);
     glutInitContextFlags( GLUT_DEBUG );
-    glutCreateWindow ("FRViewJobDebugWindow");    
+    glutCreateWindow ("FRViewJobDebugWindow");   
+    
     m_width = 1000;
     m_height = 1000; 
     // --- OpenGL context created, init glew etc. ------------------------------
@@ -88,9 +89,28 @@ FRVGLJobController::init()
 
 
     //set up the FBO that we will be rendering to
-    m_fbo = 0; //direct to screen right now please.
-
-
+    //glGenFramebuffers( 1, &m_fbo );
+    //glBindFramebuffer( GL_FRAMEBUFFER, m_fbo );
+    //GLuint fboTex;
+    //glGenTextures( 1, &fboTex );
+    //glBindTexture( GL_TEXTURE_2D, fboTex );
+    //glTexImage2D( GL_TEXTURE_2D, 0, GL_RGB, m_width, m_height, 0, GL_RGB, GL_FLOAT, nullptr );
+    //glTexParameteri( GL_TEXTURE_2D, GL_TEXTURE_MAG_FILTER, GL_NEAREST );
+    //glTexParameteri( GL_TEXTURE_2D, GL_TEXTURE_MIN_FILTER, GL_NEAREST );
+    //glTexParameteri( GL_TEXTURE_2D, GL_TEXTURE_MAX_LEVEL, 0 );
+    //glTexParameteri( GL_TEXTURE_2D, GL_TEXTURE_BASE_LEVEL, 0 );
+    //glBindTexture( GL_TEXTURE_2D, 0 );
+    //
+    //glFramebufferTexture2D( GL_FRAMEBUFFER, GL_COLOR_ATTACHMENT0, GL_TEXTURE_2D, fboTex, 0 );
+    ////glBind
+    //if(checkFramebufferCompleteness()){
+    //    glDebugMessageInsert( GL_DEBUG_SOURCE_APPLICATION, GL_DEBUG_TYPE_OTHER, 114, GL_DEBUG_SEVERITY_NOTIFICATION, -1, "Framebuffer complete, continuing with offscreen rendering.");
+    //}else{
[…]

### 2014-12-03 js/gui/Canvas.js
jpg-mode also honor the hold-png setting now

@@ -705,10 +705,13 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                    ( (this._modelLib.hasKey("ap_autoSelect")) && (this._modelLib.getElementValue("ap_autoSelect")) )      ) {
                 // console.log("Mouse up:   In JPG mode *or* autoSelect mode");
                 // console.log("Mouse up:   Setting PNG mode");
-                this._snapshotURL = this._snapshotStrings.png;
-                this._urlHandler.setURL(this._snapshotURL);
-                this._urlHandler.updateParams({snaptype: this._snapShotStringToType( this._urlHandler.getURL() )});
-                this._requestImageIfNotBusy();
+                // 141203: Wrapping this, for figure-generation
+                if ( ! ( (this._modelLib.hasKey("ap_hold_up_png")) && (this._modelLib.getElementValue("ap_hold_up_png")) ) ) {
+                    this._snapshotURL = this._snapshotStrings.png;
+                    this._urlHandler.setURL(this._snapshotURL);
+                    this._urlHandler.updateParams({snaptype: this._snapShotStringToType( this._urlHandler.getURL() )});
+                    this._requestImageIfNotBusy();
+                }
             }
         }
         this._showCorrect();

### 2014-12-03 tex/short_2015/body.tex
new stuff

@@
   \centering
-%  \includegraphics[width=.8\linewidth]{test.jpg}
-%  \vspace{-5cm}
   \input{fig2.tex}
@@
            The rendered image (of \eg, a cube) as a height map seen from above,
-           for a certain transformation.}
+           for a certain transformation. \textbf{Droppe denne?}}
 \end{figure}
@@
   \centering
-  \framebox{\includegraphics[width=.8\linewidth]{test.jpg}}
+  \subfigure[Server-rendered]{
+    \includegraphics[width=.3\linewidth]{splat1.jpg}
+  }
+  \subfigure[$P_c M_c = P_s M_s$]{
+    \includegraphics[width=.3\linewidth]{splat1.jpg}
+  }
+  \subfigure[$P_c M_c \neq P_s M_s$]{
+    \includegraphics[width=.3\linewidth]{splat2.jpg}
+  }
   \caption{\label{fig:LargeSplatsOnCorners}
-           The rendered image (of \eg, a cube) as a height map seen from above,
-           for a certain transformation.}
+           Splats for which depths are sampled on a planar region look nice on
+  exactly this region, but contorted elsewhere, notice the corner of the cube.}
 \end{figure}

### 2014-12-04 examples/jpg_proxy_test/JPCJob.cpp
More figure-generation: dumping of images, white bg, bugfix in jpg-retain-png

@@ -132,7 +132,8 @@ bool JPCJob::renderFrame(const std::string &session, const std::string &key, uns
 
     glEnable(GL_DEPTH_TEST);
 
-    glClearColor(0, 0, 0, 1);
+    // glClearColor(0, 0, 0, 1);
+    glClearColor(1.0, 1.0, 1.0, 1);
     glClear(GL_COLOR_BUFFER_BIT|GL_DEPTH_BUFFER_BIT);
 
     glViewport(0, 0, width, height);

### 2014-12-04 include/tinia/trell/IPCGLJobController.hpp

@@ -65,6 +65,7 @@ protected:
                    const size_t        depth_width,
                    const size_t        depth_height,
                    const bool          depth16,
+                   const bool          dump_images,
                    const std::string&  session,
                    const std::string&  key );
 

### 2014-12-04 include/tinia/trell/IPCJobController.hpp

@@ -125,6 +125,7 @@ protected:
                    const size_t        depth_width,
                    const size_t        depth_height,
                    const bool          depth16,
+                   const bool          dump_images,
                    const std::string&  session,
                    const std::string&  key );
     virtual

### 2014-12-04 js/gui/Canvas.js

@@ -381,6 +381,30 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
             }
         }) );
 
+//        this._modelLib.addLocalListener( "ap_dump_button", dojo.hitch(this, function(event) {
+//            if ( this._modelLib.getElementValue("ap_dump_button") ) {
+//                console.log("pressed");
+//                this._modelLib.updateElement("ap_dump_button", false);
+
+//                var img = this._canvas.toDataURL("image/png");
+
+//                console.log(img);
+
+//                fs = dojo.require('fs');
+//                sys = dojo.require('sys');
+
+//                // strip off the data: url prefix to get just the base64-encoded bytes
+//                var data = img.replace(/^data:image\/\w+;base64,/, "");
+//                dojo.require('buffer');
+//                var buf = new Buffer(data, 'base64');
+//                writeFile('image.png', buf);
+
+
+//            } else {
+//                console.log("released");
+//            }
+//        }) );
+
 
 
         // Some timing issues:
@@ -706,7 +730,7 @@ dojo.declare("gui.Canvas", [dijit._Widget], {
                 // console.log("Mouse up:   In JPG mode *or* autoSelect mode");
                 // console.log("Mouse up:   Setting PNG mode");
                 // 141203: Wrapping this, for figure-generation
-                if ( ! ( (this._modelLib.hasKey("ap_hold_up_png")) && (this._modelLib.getElementValue("ap_hold_up_png")) ) ) {
+                if ( (this._modelLib.hasKey("ap_hold_up_png")) && (!this._modelLib.getElementValue("ap_hold_up_png")) ) {
                     this._snapshotURL = this._snapshotStrings.png;
                     this._urlHandler.setURL(this._snapshotURL);
                     this._urlHandler.updateParams({snaptype: this._snapShotStringToType( this._urlHandler.getURL() )});

### 2014-12-04 js/gui/ProxyRenderer.js

@@ -168,6 +168,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
         this._shaderSourceLoaded = false;
         this._loadShaders();
+        var reloadShaders = false; // Will be set to true if some magical exposedModel variable dictates the modification of a macro used by one or more shaders.
 
         // Setting up listeners for known configurable parameters. These are mainly for debugging and testing. (Meaning
         // that modification of defaults are for testing.) The application proxyCube sets up a GUI for manipulating
@@ -327,7 +328,7 @@ dojo.declare("gui.ProxyRenderer", null, {
             if ( this.exposedModel.hasKey("ap_useFragExt") ) {
                 this._useFragDepthExt = this.exposedModel.getElementValue("ap_useFragExt") ? 1 : 0;
                 //console.log("xxxxxxxxxxxxxxxx compiling shaders because initial ap_useFragExt exists...");
-                this._loadShaders(); // Must use this and not _compileShaders directly, since we cannot be sure that source has been loaded otherwise.
+                reloadShaders = true;
             }
             this.exposedModel.addLocalListener( "ap_useFragExt", dojo.hitch(this, function(event) {
                 this._useFragDepthExt = this.exposedModel.getElementValue("ap_useFragExt") ? 1 : 0;
@@ -336,7 +337,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
             //-------------------------------------------------------
             if ( this.exposedModel.hasKey("ap_depthWidth") ) {
-                this._loadShaders();
+                reloadShaders = true;
             }
             this.exposedModel.addLocalListener( "ap_depthWidth", dojo.hitch(this, function(event) {
                 this._loadShaders();
@@ -344,7 +345,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
             //-------------------------------------------------------
             if ( this.exposedModel.hasKey("ap_depthHeight") ) {
-                this._loadShaders();
+                reloadShaders = true;
             }
             this.exposedModel.addLocalListener( "ap_depthHeight", dojo.hitch(this, function(event) {
                 this._loadShaders();
@@ -352,7 +353,7 @@ dojo.declare("gui.ProxyRenderer", null, {
 
             //-------------------------------------------------------
             if ( this.exposedModel.hasKey("ap_simulate_downsampling") ) {
-                this._loadShaders();
+                reloadShaders = true;
             }
[…]

### 2014-12-04 src/trell/IPCGLJobController.cpp

@@ -392,6 +392,7 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
                                    const size_t        depth_width,
                                    const size_t        depth_height,
                                    const bool          depth16,
+                                   const bool          dump_images,
                                    const std::string&  session,
                                    const std::string&  key )
 {
@@ -564,26 +565,26 @@ IPCGLJobController::onGetSnapshot( char*               buffer,
                     }
                 }
             }
-#if 0
-            static int cntr=0;
-            {
-                char fname[1000];
-                sprintf(fname, "/tmp/rgb_%05d.ppm", cntr);
-                FILE *fp = fopen(fname, "w");
-                fprintf(fp, "P6\n%lu\n%lu\n255\n", width, height);
-                fwrite(buffer_pos - 4*((width*height*3 + 3)/4), 1, 3*width*height, fp);
-                fclose(fp);
+            if (dump_images) {
+                static int cntr=0;
+                {
+                    char fname[1000];
+                    sprintf(fname, "/tmp/trell_rgb_%05d.ppm", cntr);
+                    FILE *fp = fopen(fname, "w");
+                    fprintf(fp, "P6\n%lu\n%lu\n255\n", width, height);
+                    fwrite(buffer_pos - 4*((width*height*3 + 3)/4), 1, 3*width*height, fp);
+                    fclose(fp);
+                }
+                {
+                    char fname[1000];
+                    sprintf(fname, "/tmp/trell_depth_%05d.ppm", cntr);
+                    FILE *fp = fopen(fname, "w");
+                    fprintf(fp, "P6\n%lu\n%lu\n255\n", width, height);
+                    fwrite(buffer_pos, 1, 3*width*height, fp);
+                    fclose(fp);
+                }
+                cntr++;
             }
-            {
-                char fname[1000];
-                sprintf(fname, "/tmp/depth_%05d.ppm", cntr);
-                FILE *fp = fopen(fname, "w");
-                fprintf(fp, "P6\n%lu\n%lu\n255\n", width, height);
-                fwrite(buffer_pos, 1, 3*width*height, fp);
-                fclose(fp);
-            }
-            cntr++;
-#endif
         }
        break;
     }

### 2014-12-04 src/trell/IPCJobController.cpp

@@ -88,6 +88,7 @@ IPCJobController::onGetSnapshot( char*               buffer,
                                  const size_t        depth_width,
                                  const size_t        depth_height,
                                  const bool          depth16,
+                                 const bool          dump_images,
                                  const std::string&  session,
                                  const std::string&  key )
 {
@@ -230,6 +231,7 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
         int depth_width = q->depth_h;
         int depth_height = q->depth_h;
         // bool depth16 = q->depth16; @@@ should probably be gotten this way, too... for now, fetching from exposed model below...
+        // bool dump_images = q->dump_images; @@@ should probably be gotten this way, too... for now, fetching from exposed model below...
 
         std::string key_list_string = std::string( q->viewer_key_list );
         std::vector<std::string> key_list;
@@ -275,12 +277,14 @@ IPCJobController::handle( tinia_msg_t* msg, size_t msg_size, size_t buf_size )
 
         bool depth16;
         m_model->getElementValue( "ap_16_bit_depth", depth16 );
+        bool dump_images;
+        m_model->getElementValue( "ap_dump", dump_images );
 
         // Looping through all keys and grabbing GL-content
         char *buf = (char*)msg + sizeof(tinia_msg_image_t);
         for (size_t i=0; i<key_list.size(); i++) {
             key = key_list[i]; // Overriding the key gotten from the 'msg' parameter!
-            if ( onGetSnapshot(buf, format, w, h, depth_width, depth_height, depth16, session, key) ) { // onGetSnapshot() in IPCGLJobController grabs pixels for a given key
+            if ( onGetSnapshot(buf, format, w, h, depth_width, depth_height, depth16, dump_images, session, key) ) { // onGetSnapshot() in IPCGLJobController grabs pixels for a given key
 #ifdef DEBUG
                 m_logger_callback( m_logger_data, 2, package.c_str(),
                                    "Queried for snapshot, ok. format = %d", format );

### 2014-12-04 src/utils/ProxyDebugGUI.cpp

@@ -154,6 +154,10 @@ ProxyDebugGUI::ProxyDebugGUI( boost::shared_ptr<model::ExposedModel> model,
         model->addAnnotation("ap_greenbg", "Green bg");
         model->addElement<bool>( "ap_blackbg", false );
         model->addAnnotation("ap_blackbg", "Black bg");
+        model->addElement<bool>( "ap_dump", false );
+        model->addAnnotation("ap_dump", "Dump rgb (trell)");
+        model->addElement<bool>( "ap_dump_button", false );
+        model->addAnnotation("ap_dump_button", "Dump (js)");
     }
 }
 
@@ -276,6 +280,9 @@ tinia::model::gui::Grid *ProxyDebugGUI::getGrid()
         mainGrid->setChild(row, 0, new tinia::model::gui::Button("ap_whitebg"));
         mainGrid->setChild(row, 1, new tinia::model::gui::Button("ap_greenbg"));
         mainGrid->setChild(row, 2, new tinia::model::gui::Button("ap_blackbg"));
+        row++;
+        mainGrid->setChild(row, 0, new tinia::model::gui::CheckBox("ap_dump"));
+        mainGrid->setChild(row, 1, new tinia::model::gui::Button("ap_dump_button"));
         row++;
     }
 

### 2014-12-04 tex/short_2015/body.tex
mainly figures

@@
 \item Maa ta et lite soek paa prior art innen omraadet
+
+\item problem med shading/shadows ++ i forb. med ikke deferred shading/normal map
+
+\item About our test case: corners, smooth surfaces, texturing for which artifacts are easy to
+spot. What it does lack is self-occluding parts. The FRView rservoir model that
+we have tested on does contain some of this.
+
+\item Burde vi vise chemco?
 
@@
   \subfigure[$P_c M_c = P_s M_s$]{
-    \includegraphics[width=.3\linewidth]{splat1.jpg}
+    \includegraphics[width=.3\linewidth]{splat2.jpg}
   }
   \subfigure[$P_c M_c \neq P_s M_s$]{
-    \includegraphics[width=.3\linewidth]{splat2.jpg}
+    \includegraphics[width=.3\linewidth]{splat3.jpg}
   }
@@
   \centering
-  \subfigure[Low quality (tenker her jpg med q=0 og daarlig dybdebuffer, faa
-  splats kanskje]{
-    \input{fig3a.tex}
+  \subfigure[Full png]{
+    \includegraphics[width=.30\linewidth]{auto1.jpg}
   }
-  \subfigure[High quality (png med maks av alt, mange splats)]{
-    \input{fig3b.tex}
+  \subfigure[Coarse jpg]{
+    \includegraphics[width=.30\linewidth]{auto2.jpg}
+  }
+  \subfigure[Auto-proxy w/$260^2$ splats and 1/256 rgb to depth fragment ratio]{
+    \includegraphics[width=.30\linewidth]{auto3.jpg}
   }
@@
            Proxy models adaptively adjusted to available bandwidth and server
-           capacity.}
+           capacity. The server-rendered JPG image adaptively selects a quality
+  setting based on timing of image sending. Shown is the worst
+  case, \texttt{quality}=0, which is the last step before switching to ``auto
+  proxy''. The auto-proxy shown displays artifacts caused by the magenta surface
+  to being almost parallel to the viewing direction at the moment of proxy model
+  recording, the number of splats is set low ($260^2$) on purpose, and the depth
+  buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$.}
 \end{figure}
@@
   \subfigure[Server-rendered]{
-    \input{fig3a.tex}
+    \includegraphics[width=.47\linewidth]{frview1.jpg}
   }
   \subfigure[Client-rendered proxy model]{
-    \input{fig3b.tex}
+    \includegraphics[width=.47\linewidth]{frview2.jpg}
   }
   \caption{\label{fig:FRView}
-           The oil reservoir viewer FRView}
+           The oil reservoir viewer FRView. Not that we have chosen parameters
+  (small number of splats ($128^2$), forcing display of splats that would normally be
+  discarded) to emphasize artifacts, \eg, the ragged edges, a few holes, and the
+  oddly placed (and rendered) axis indicator.}
 \end{figure}

### 2014-12-05 tex/short_2015/body.tex
new revision

@@
 \begin{abstract}
-   We describe an algorithm, and implementation, of an automatic proxy geometry
-   system for a server-client remote rendering setup. The client may be a thin
+   We describe an algorithm and implementation of an automatic proxy geometry
+   system for a server/client remote rendering setup. The client may be a thin
    client in the form of a hand-held pad, for our implementation we only assume
-   the availability of WebGL. For our rendering server, we make use of a
-   GL-capable web server. In situations where the rate of received images
-   deteriorate due to low bandwith, high latency or long rendering times, the
-   client displays an automatically generated proxy geometry. This proxy
+   the availability of WebGL and Javascript. For our rendering server, we make
+   use of a GL-capable web server. In situations where the rate of received
+   images deteriorate due to low bandwith, high latency or long rendering times,
+   the client displays an automatically generated proxy geometry. This proxy
    geometry is computed from depth buffers bundled with rendered images from the
-   server.
+   server, and requires no modification of the application or its data itself.
    % Husk: Skal vaere en blank linje nedenfor
@@
 %-------------------------------------------------------------------------
-\section{Notes}
+\section{Introduction}
 
-\begin{itemize}
-
-\item We restrict ourselves to discuss ``splat sets'' in this paper. (Nevne der det
-tas frem andre alternativer)
-
-\item
-Husk aa faa med det viktige i at vi faar oekt serverkapasiteten til aa
-serve flere klienter ogsaa!
-
-\item
-``server-image'' i steden for ``server-rendered image'' for aa spare plass?
-
-\item received stavekontroll
-
-\item Hvor passer det aa faa inn 1) bilde og 2) tekst om FRView?
-
-\item Maa ta et lite soek paa prior art innen omraadet
-
-\item problem med shading/shadows ++ i forb. med ikke deferred shading/normal map
-
-\item About our test case: corners, smooth surfaces, texturing for which artifacts are easy to
-spot. What it does lack is self-occluding parts. The FRView rservoir model that
-we have tested on does contain some of this.
-
-\item Burde vi vise chemco?
-
-\end{itemize}
-
-%-------------------------------------------------------------------------
-\section{Introduction}
+\begin{figure}[htb]
+  \centering
+  \subfigure[Server-rendered (OpenGL)]{
+    \includegraphics[width=.47\linewidth]{frview1.jpg}
+  }
+  \subfigure[Client-rendered proxy model (Google Chrome)]{
+    \includegraphics[width=.47\linewidth]{frview2.jpg}
+  }
+  \caption{\label{fig:FRView} The oil reservoir viewer FRView, showing a
+  server-rendered image and a proxy geometry generated from this and the
+  accompanying depth buffer. Note that we have chosen parameters for the proxy
+  generation (small number of splats ($128^2$), forcing display of splats that
+  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
+  the ragged edges, a few holes, and the oddly placed (and rendered) axis
+  indicator.}
+\end{figure}
 
@@
 only become more interactive. Alas, hand in hand with this development we also
-find ever increasing requirements on computation accuracy, power efficiency,
-data sizes and so on.
+find ever increasing requirements on computational accuracy, power efficiency,
+data sizes, scaling, etc.
 
-One result of this is that server-side rendering is as relevant as ever
-before. And, with a client-server model for rendering, the problems with
-transferring images from the server to the model is of course also still
-present.
+One effect of this, also enhanced by new cloud-based approaches, wireless
+connection, and mobile devices with display capabilities outpacing their
+computation and storage capabilities, is that interactivity still is a difficult
+issue. In this work, we consider a server/client model for rendering, but this
+is just a way of separating the interaction from the rendering; the discussion
+also applies to a more integrated setup.  In order to obtain the frame rates on
+the client suitable for interactive visualizations, demands are put on
+bandwidth, latency and rendering speed. If one or more of these demands are not
+met, according to a suitable measure, the frame rate will deteriorate together
+with the user experience.
 
-In order to obtain the frame rates on the client suitable for interactive
-visualizations, demands are put on bandwidth, latency and rendering speed. If
-one or more of these demands are not met, according to a suitable measure, the
-frame rate will deteriorate together with the user experince.
-
-One way to deal with such situations is for the client to have a so-called proxy
-model, which is defined as a temporary model to be shown and manipulated locally
-on the client while waiting for the appropriate image from the server.
-
+One way to deal with such situations is for the client to have a {\em
+proxy model}, defined as a temporary model to be shown and manipulated
+locally on the client while waiting for the appropriate image from the server.
 A particularly unwelcomed chore is that of building proxy geometries for various
@@
 box wire frame, but this will degrade the user experience unacceptably for most
-applications. One alternative is to hand-build such simpler proxy
-geometries. They have to be simple both for the building cost, and also for the
-actual rendering on the possibly very thin client. Another alternative is to
-automate this process, a large triangulation may for instance be represented by
-a much coarser approximating triangulation that can be built automatically. One
-disadvantage of this approach is that this is not easy to do for many kinds of
-data.
+applications. One alternative is to hand craft such simpler proxies. They have
+to be simple both for the building cost, and also for the actual rendering on
+the possibly very thin client. Another alternative is to automate this process,
+a large triangulation may for instance be represented by a much coarser
+approximating triangulation that can be built automatically. One disadvantage of
+this approach is that this is not easy to do for many kinds of data, like our
+main case in which we have an oil reservoir viewer (FRView) that renders large
+{\em corner points grids}, together with {\em faults}, oil wells, and more. See
+Figure~\ref{fig:FRView} above for an example of both a fully completed
+rendering, and an automatically generated proxy geometry rendered in Google
+Chrome.
 
-Our suggested solution is to pass along depth information from the server along
-with ordinary rendered images. From this depth image, a rudimentary 3D model can
-be built, and using the ordinary image as a texture, this model can be
-transformed and rendered on the client while waiting for the next update from
-the server. If the server does not take too long to respond, and/or the client
-does not change the position or orientation of the model too much, it is to be
-hoped that this display of a proxy model integrates seemlessly with the slower
-stream of server-rendered frames.
+Our suggested solution is to pass depth information from the server along with
+ordinary rendered images. From this depth image, a rudimentary 3D model can be
+built, and using the ordinary image as a texture, this model can be transformed
+and rendered on the client while waiting for the next update from the server. If
+the server does not take too long to respond, and/or the client does not change
+the position or orientation of the model too much, this display of a proxy model
+integrates seemlessly with the slower stream of server-rendered frames.
 
-[Kanskje ha inn en foreløpig figur-skisse her, med dybdefelt vist som en
-  høydeflate...?]
+We also mention that even if bandwidth and latency is not a problem, it may be
+desirable to let a server of limited capacity serve many simultaneous users,
+hence decreasing the effective server time available for each one. In this case,
+suitable scaling properties may still be achieved without increasing server
+capacity. \textbf{referanse til ArrowHead her?}
 
-[Trenger referanser til state of the art o.l. her]
+\subsection{Previous work}
+
+Our approach has some similarities to {\em image based rendering} (IBR)
+techniques (\textbf{referanse til noe av dette?}), with the difference that an
+important IBR problem would be the reconstruction of a depth map from images,
+while we have access to the full depth map from a rendering pass.
+Another way to use depth maps similar to what we do, is for ``immersive
+streaming'',
+see \eg, \texttt{http://www.cc...},
+% http://www.cc.gatech.edu/conferences/3DPVT08/Program/Papers/paper203.pdf
+where focus is on depth map compression, an issue that we also must confront.
+
+We are not aware of anybody else using depth maps exactly for our purpose, most
+research seems to be focused on the retrieval of the depth map from images,
+while we use an existing depth map to distill and render temporary geometry.
 
@@
 
-The first main idea is to automatically construct a proxy geometry on the client
-side from depth data provided by the server. For a new frame to be displayed on
-the client, the server renders an image in a frame buffer, and it is assumed
-that it in this process also renders a corresponding depth buffer. This depth
-buffer may be regarded as a height map over a viewing plane relative to the
-observing client. Hence, this depth map does not contain information about scene
-elements not in view for the set of view parameters at the time of server side
+The main idea is to automatically construct a proxy geometry on the client side
+from depth data provided by the server. For a new image to be displayed on the
+client, the server renders an image in a frame buffer, and as part of this
+process it also renders a corresponding depth buffer. (This means that our
+technique is not immediately ready for \eg, ray-tracing.) This depth buffer may
+be regarded as a height map over a viewing plane relative to the observing
+client. Hence, this depth map does not contain information about scene elements
+not in view for the set of view parameters at the time of server side
 rendering. Our approach is based on the assumption that small transformations of
@@
 while waiting for the next update from the server. See
-Figure~\ref{fig:2Dheightmap} and Figure~\ref{fig:2DheightmapRoated} for an
+Figure~\ref{fig:2Dheightmap} and Figure~\ref{fig:2DheightmapRotated} below for an
 illustration of this heightmap computed on the server, and a recovered 3D model
@@
   \input{fig2.tex}
-  \caption{\label{fig:2Dheightmap}
-           The rendered image (of \eg, a cube) as a height map seen from above,
-           for a certain transformation. \textbf{Droppe denne?}}
+
+  \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
+           height map seen from above, for a certain transformation. Also shown
+           are the corresponding RGB-fragments (red and green) and depth
+           fragments (floating point scalars shown as grayscale pixels.)}
 \end{figure}
+
+In Figure~\ref{fig:2DheightmapRotated} below, a sequence of three server-rendered
+images (solid lines, red and blue) is shown, together with intermediate
+client-rendered proxy models with different features. These features will be
+discussed later, in Section~\ref{sec:client}.
 
@@
 information in the server side model. For instance, the solid surface shown in
-Figure~\ref{fig:2Dheightmap} cannot be deduced, hence we show the dots in
-Figure~\ref{fig:2DheightmapRotated}. Still, there are several things we can do
-to improve the client-side computed approximation of the server-side image, and
-these will now be described.
+Figure~\ref{fig:2Dheightmap} (\ie, the {\em topology}) cannot be deduced, hence
+we show the dots in Figure~\ref{fig:2DheightmapRotated}~(b). Still, there are
+several things we can do to improve the client's approximation of
+the server-side image, and these will now be described.
 
@@
 with a connected depth buffer, it has a negligible cost for the server to bundle
-the resulting depth map with the rendered image upon return of the render
-result. In practice, this is easily handled by a few extra lines of computer
-code where the image is extracted from its framebuffer and prepared for sending
-to the client.
+the resulting depth map with the rendered image upon return of the result. In
+practice, this is easily handled by a few extra lines of computer code where the
+image is extracted from its framebuffer and prepared for sending to the
+client.
 
-The cost must of course be considered if the operation of packaging images for
-sending to the client is significant, which may be the case if, \eg, images must
-be compressed or converted to some special format. An example of the latter
-might be {\texttt base64}-encoding for converting the image to a purely ASCII
-format. Together with the rendered image (which we will call ``rgb image'' to
-avoid confusion with the ``depth image'') and the depth image, we will also send
-the current view transformations, back to the client.
+Since the depth buffer adds somewhat to the data being sent, it is important to
+keep it to a minimum. We have found that reducing the spatial resolution of the
+depth buffer (for instance by a factor of $1/16$) only to a negligible extent
+degrades the proxy model being reconstructed by the client. We encode each depth
+value in the range $[0, 1]$ as a 16 bit fixed point number, this also works
+quite well. These measures cause the bundling of the depth buffer with the rgb
+image to impose just a small data overhead.
+
+Further compression methods (\textbf{passende med en ref her?}) may be employed
+to bring this data size down even more, but then the cost of the compression
+itself must also be added to the cost of rendering on the server.
+%
+%The cost must of course be considered if the operation of packaging images for
+%sending to the client is significant, which may be the case if, \eg, images must
+%be compressed or converted to some special format. An example of the latter
+%might be {\texttt base64}-encoding for converting the image to a purely ASCII
+%format.
+%
+Together with the rendered rgb image and the depth image we will also send the
+current view transformations back to the client.
 
@@
 \subsection{The client}
+\label{sec:client}
 
-When the client receieves an rgb image, a depth image and view transformations,
-it builds a ``proxy model'' from this. This model can then be rendered directly,
-if there is no available image from the server for a new transformation provided
-by the user, or it can be combined with other proxy models the client already
-has. The latter option will be discussed in
-Section~\ref{sec:proxyModelReplacement} below.
+When the client receieves an rgb and depth image, together with view
+transformations, it builds a ``proxy model'' from this. This model can then be
+transformed and rendered directly if there is no available image from the
+server, or it can be combined with other proxy models the client already has in
+store. The latter option will be discussed in
+Section~\ref{sec:proxyModelReplacement}.
 
 %-------------------------------------------------------------------------
-%\subsubsection{Proxy model building / Splatting}
+\subsubsection{Proxy model building, splats}
 
-As indicated by Figure~\ref{fig:2DheightmapRotated}~(a), the received
+As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
 height map does not really allow us more than concluding where a set of points
@@
 set of 3D points with the color sampled from the corresponding location in the
-server-rendered image, as illustrated in
+server-rendered image, this is illustrated in
 Figure~\ref{fig:2DheightmapRotated}~(b).
 
-We may set up the server to send a depth buffer of different resolution than the
-actual image, and the client can render at most one 3D point for each depth
+It makes sense for the client only to render at most one 3D point for each depth
 fragment available. This may be too much for a thin client, and one may want to
 operate with a smaller number of 3D points, and instead render each of them with
-a larger number of pixels on the client side. This opens up for problems and
-possibilities connected to {\em splatting}, \eg, what size and shape of
-geometries to render for each 3D point, how to color them, how to compute depth
-fragments and so on.
+a larger number of pixels on the client side, this is called a {\em
+splat}. A set of such splats for a given depth image, we will call a {\em proxy
+model}. There are some issues to discuss in relation to this, \eg, what size and
+shape of geometries to render for each 3D point, how to color them, how to
+compute depth fragments and so on.
 
-The most simple solution is to render each 3D point as a fixed geometry, for
-instance a 2D sphere or rectangle, with the correspondig color from the image,
-see Figure~\ref{fig:2DheightmapRotated}~(c). We will briefly describe some
-improvements to this. We also note that other possibilities exist apart
-from this ``splat-based'' proxy model, \eg, building and maintaining a 3D
-occupancy mesh, a computing a distance field from which iso-surfaces can be
-extracted, and so on.
-
-In the following, we will refer to a set of reconstructed 3D points from a depth
-map as a {\em proxy model}, and the reconstructed points as {\em splats}.
-
-%-------------------------------------------------------------------------
-\subsubsection{Splat sizing}
-
-Note that as soon as a given proxy model is transformed, each splat should
-ideally be rendered into a number of client pixels according to the new splats
-3D position. A splat transformed into view closer to the observer, should result
-in more client fragments, and vice versa.
-
-...
+The simplest solution is to render each 3D point as a fixed geometry, for
+instance a 2D disk or rectangle, with the correspondig color from the image, see
+Figure~\ref{fig:2DheightmapRotated}~(c), where we have added the transformation
+local to the client, which the server has not been able to provide an image for.
+We will briefly describe some improvements to this. We also note that other
+possibilities exist apart from this ``splat-based'' proxy model, \eg, building
+and maintaining a 3D occupancy mesh, computing a distance field from which
+iso-surfaces can be extracted, etc.
 
@@
 
-From the previous section, we note that the splats may occupy more client
-fragments than just one, even if the should be one to one correspondence between
-server depth fragments and client pixels. This will of course also be the case
-if the client's capabilities dictate a realtively small number of splats.
-
-In any case, a splat may result in many client fragments, which necessitates an
+Our splats may result in many client window fragments, necessitating an
 ``intra-splat'' fragment coloring or texturing. This is not entirely straight
 forward, since the client really do not know how a finitely-sized splat should
-map to the server's model, having only a discrete set of 3D points on that
+map to the server's geometry, having only a discrete set of 3D points on that
 model.
-
-\vspace{10cm}
 
@@
   T =
-  \begin{pmatrix} 1/\text{splats}_x & 0 \\ 0 & 1/\text{splats}_y \end{pmatrix}
+  \begin{pmatrix}
+    \frac{1}{\text{splats}_x} & 0 \\
+    0 & \frac{1}{\text{splats}_y}
+  \end{pmatrix}
   \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
-  \begin{pmatrix} \text{vpwidth}/\text{splats}_x & 0 \\ 0 & \text{vpheight}/\text{splats}_y \end{pmatrix}
+  \begin{pmatrix} 
+    \frac{\text{vpwidth}}{\text{splats}_x} & 0 \\
+     0 & \frac{\text{vpheight}}{\text{splats}_y}
+  \end{pmatrix}
   \cdot \text{overlap},
 \]
-where $(\ev_1 \, \, \, \ev_2)$ maps the ``client's splat region'' (in $[0,
-1]^2$) to the corresponding texture area, $(\sv_{dx} \, \, \, \sv_{dy})^{-1}$
-maps the ``client's screen space splat area'' to $[0, 1]^2$, and $\text{splatSize}_x
-= \text{vp}_\text{width}/\text{splats}_x$ and $\text{splatSize}_y
-= \text{vp}_\text{height}/\text{splats}_y$ together with the ``overlap'' just
-provides a scaling factor so that ``overlap=1'' produces non-overlapping but
-covering splats when $M_c=M_s$, \ie, the client does no transformation of the
-proxy model.
-
+\[
+   =
+  A
+  \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
+  B
+  \cdot \text{overlap},
+\]
+where $A$ maps the ``client's splat region'' (in $[0, 1]^2$) to the
+corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
+``client's screen space splat area'' to $[0, 1]^2$, and $B$ together with the
+``overlap'' just provides a scaling factor so that ``overlap=1'' produces
+non-overlapping but covering splats when $M_c=M_s$, \ie, the client does no
+transformation of the proxy model.
+%
 % After mult by proj matrix: clip coo, range of (x, y, z) = [-w, w]^3
@@
 % (Range [vp[0] vp[1]] x [[vp[2] vp[3]] x [0, 1] ?)
-
-We compute $\sv_{dx}$ and $\sv_{dy}$ by evaluating proxy model positions and
+%
+We compute $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions and
 performing the perspective division and transformation into window coordinates
@@
 
-We use $\Delta = \Delta x = \Delta y = ...$, so that ...
+We use
+\begin{equation}
+  \Delta x = \text{splats}_x / \text{DepthImg-width}
+  \label{eq:deltaChoice}
+\end{equation}
+\[
+  \Delta y = \text{splats}_y / \text{DepthImg-height}.
+\]
+\textbf{fiks de likningene over}
+Note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
+chosen something larger. The values are used for looking up depth image samples,
+so we should not choose something too small, our choice is exactly to look for
+neighbouring depth samples. Since this calculation of $\sv_x$ and $\sv_y$ is
+really a difference approximating a gradient on the geometry, we should not make
+$\Delta x$ and $\Delta y$ to large either. Experiments show us
+that~(\ref{eq:deltaChoice}) is really a good choice.
 
-\textbf{ Litt om valgene vi kan gjoere for Delta}
-
-This again leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
+This then leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
 coordinates, and after perspective division and viewport transformation, we get
@@
 
-Then, ...
+Then, on the client we render a \texttt{glPoint} for each splat, with texture
+coordinates $(s, t)^\prime$ and \texttt{glPointSize} set to a larger value that will
+produce many fragments. Each fragment will then look up the server-rendered
+image at position
+\[
+  \begin{pmatrix}
+    s \\ t
+  \end{pmatrix} +
+  T 
+  \begin{pmatrix}
+    u \\ v
+  \end{pmatrix}
+\]
+where $(u, v)^\prime = \texttt{gl\_PointCoord}-0.5$ are ``intra-splat texture
+coordinates''. \textbf{Boer kanskje sjekke at det ikke er noen faktor 2 som har
+falt ut noe sted paa veien...}
 
-such that the client fragment $\pv_{i, j} + \delta_{i, j}$ gets texel
-\[
-  I_{p, q},
-\]
-with $(p, q) = \ldots$.
-
-When the assumption does not hold, \eg, if the splats are very large, or they
-originates from a non-planar part of the server's model, this may look rather
-odd, see Figure~\ref{fig:LargeSplatsOnCorners}.
+When the assumption that the geometry is locally planar does not hold, \eg, if
+the splats are very large, or they originate from a rather curved or non-smooth
+part of the geometry, this may look rather odd, see
+Figure~\ref{fig:LargeSplatsOnCorners}.
 
@@
            Splats for which depths are sampled on a planar region look nice on
-  exactly this region, but contorted elsewhere, notice the corner of the cube.}
+  exactly this region, but contorted elsewhere, notice the corner and edges of
+  the cube.}
 \end{figure}
 
-The problem with this, is that the geometry is not planar, and our simple texture
-coordinate transformation will produce the wrong result. Some ways to remedy
-this, is to choose smaller (and more) splats, detect the problem and discard
-such splats, or introduce more complex texture transformations. Since the proxy
-geometry in principle is to be short-lived, and only a temporary geometry to be
-rendered before the server-rendered image can be receieved, we have chosen the
-first alternative. Discarding such splats may offer a good alternative,
-especially when combining several proxy geometries, where the union of retained
-splats increases the likelihood of all geometry parts being covered. Note also
-that implementing a more sophisticated texture coordinate transformation may
-actually amount to performing the same work as for more and smaller splats. Or,
-in other words, using more and smaller splats may be regarded as a better
-texture transform implementation.
+The problem illustrated in Figure~\ref{fig:LargeSplatsOnCorners}, is that the
+geometry is far from planar, and our simple texture coordinate transformation
+will produce the wrong result for the corresponding part of involved
+splats. Some ways to remedy this is, \eg, to choose smaller (and more) splats,
+detect the problem and discard such splats, or introduce more complex texture
+transformations. Since the proxy geometry in principle is to be short-lived, and
+only a temporary geometry to be rendered before the server-rendered image can be
+receieved, we have chosen a combination of the first two alternatives. Note that
+implementing a more sophisticated texture coordinate transformation may actually
+amount to performing the same work as for more and smaller splats. Equivalently
+, using more and smaller splats may be regarded as a better texture transform
+implementation.
+
+%-------------------------------------------------------------------------
+\subsubsection{Splat sizing}
+
+Note that as soon as a given proxy model is transformed, each splat should
+ideally be rendered into a number of client pixels according to the new splats
+3D position. A splat transformed into view closer to the observer, should result
+in more client fragments, and vice versa. To achieve this, we simply use the
+vectors $\sv_{\Delta x}$ and $\sv_{\Delta y}$ from the previous section.
+In addition to this, we scale the splats up a bit, typically with a factor 2, so
+that they overlap. The benefit of this is that we can render rectangular
+window-aligned splats on the client, without getting large uncovered areas when
+interactively rotating and scaling the model quickly on the client.
 
@@
 For larger splats, covering many fragments, it makes sense to also compute and
-use depth fragments. This will improve the usage of more proxy models
-simultaneously, since the risk of a splat with a not very representative depth
-value covers another and better splat is reduced. The ``intra-splat'' depth
-values can easily be fetched from the depth image, just as the texture is looked
-up for color. One problem is that many thin clients may not support this, in
-WebGL, the ``frag depth'' feature is an extension, which may not be supported
-everywhere.
+use depth fragments in the client's fragment shader. This will improve the usage
+of more proxy models simultaneously, since the risk of a splat with a not very
+representative depth value covers another and better splat, is reduced. The
+``intra-splat'' depth values can easily be fetched from the depth image, just as
+the texture is looked up for color, with the same texture transformation. One
+problem is that many thin clients may not support this, in WebGL, the ``frag
+depth'' feature is an extension, which may not be supported everywhere.
 
@@
 
-In this paper we concern ourselves with proxy models defined as sets of
+In this paper we mainly concern ourselves with proxy models defined as sets of
 splats. Since each splat set is the product of one instance of a server-rendered
 image, depth buffer and viewing parameters, it makes sense to retain more than
-one such model on the client. By doing this, the client may combine them, in
+one such model on the client. By doing this, the client may combine them in
 order to provide renderings for viewing parameters very different from the last
-one received. One can imagine a plethora of splat set replacement algorithms, we
-have tested two approaches which both works well; both retain a constant number
-of proxy models. The first algorithm is to replace the one with a viewing
-direction differing the most from the newly received model. The second simply
-replaces the oldest one in store. The third replaces a proxy model $j$ if the
-replacement results in the following objective function being reduced,
+one received.
+
+As soon as the client has a set of such proxy models, it will also need an
+algorithm for replacing old models with new ones.  One can imagine a plethora of
+{\em splat set replacement algorithms}, we have tested three approaches that all
+retain a constant number of proxy models. The first algorithm is to replace the
+one with a viewing direction differing the most from the newly received
+model. The second simply replaces the oldest one in store. The third replaces a
+proxy model $j$ if the replacement results in the following objective function
+being reduced,
 \[
@@
 %-------------------------------------------------------------------------
-\subsection{Dynamic/heuristic method selection}
+\subsection{Dynamic/heuristic method selection, depth compression}
 
@@
   \centering
-  \subfigure[Full png]{
+  \subfigure[Full png image,\newline full depth buffer]{
     \includegraphics[width=.30\linewidth]{auto1.jpg}
   }
-  \subfigure[Coarse jpg]{
+  \subfigure[Coarse jpg,\newline quality $q=0$]{
     \includegraphics[width=.30\linewidth]{auto2.jpg}
   }
-  \subfigure[Auto-proxy w/$260^2$ splats and 1/256 rgb to depth fragment ratio]{
+  \subfigure[Auto-proxy, trun-\newline cated depth data]{
     \includegraphics[width=.30\linewidth]{auto3.jpg}
   }
-  \caption{\label{fig:adapativeProxyModels}
-           Proxy models adaptively adjusted to available bandwidth and server
-           capacity. The server-rendered JPG image adaptively selects a quality
-  setting based on timing of image sending. Shown is the worst
-  case, \texttt{quality}=0, which is the last step before switching to ``auto
-  proxy''. The auto-proxy shown displays artifacts caused by the magenta surface
-  to being almost parallel to the viewing direction at the moment of proxy model
-  recording, the number of splats is set low ($260^2$) on purpose, and the depth
+  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
+           available bandwidth and server capacity. The server-rendered JPG
+           image adaptively selects a quality setting based on recorded
+           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
+           the last step before switching to ``auto proxy'' (c). The auto-proxy
+           model shown displays artifacts caused both by the magenta surface
+           being almost parallel to the viewing direction at the moment of proxy
+           model recording, and by the non-smooth parts of the geometry
+           (corners). The number of splats is set low on purpose, and the depth
            buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$.}
@@
 programming framework for setting up and managing a client/server based
-interactive OpenGL-based visualization application. The client is
-Javascript/WebGL-based and runs in one of several possible browsers, \eg Chrome
-or Firefox. This client communicates over tcp/ip, using either http or
+interactive OpenGL-based visualization application, see~\cite{tinia}. As client
+we use Google Chrome, the code is written in standard Javascript/WebGL.
+The client communicatesover tcp/ip, using either http or
 websockets, with a server running as an Apache module.
@@
 feature available. We have tested the algorithm on several smaller test cases,
-but also on a larger oild reservoir viewer.
+but also on a larger oil reservoir viewer, FRView, developed in the CloudViz
+project, see~\cite{cloudviz}.
 
 This viewer utilizes several GLSL shaders to render reservoir cells and
-boundaries, tubular wells etc., and operates in general on a geometric model
-that is not trivial to reduce in complexity. The model is typically also very
-large, meaning that rendering on a thin client is prohibitive. With the
-automatic proxy geometry, we can obtain interactive frame rates with a bad
-connection from a very lightweight client to the server. For a comparison of a
-server-rendered image and a client-rendered proxy model, see
-Figure~\ref{fig:FRView}.
-
-\begin{figure}[htb]
-  \centering
-  \subfigure[Server-rendered]{
-    \includegraphics[width=.47\linewidth]{frview1.jpg}
-  }
-  \subfigure[Client-rendered proxy model]{
-    \includegraphics[width=.47\linewidth]{frview2.jpg}
-  }
-  \caption{\label{fig:FRView}
-           The oil reservoir viewer FRView. Not that we have chosen parameters
-  (small number of splats ($128^2$), forcing display of splats that would normally be
-  discarded) to emphasize artifacts, \eg, the ragged edges, a few holes, and the
-  oddly placed (and rendered) axis indicator.}
-\end{figure}
+boundaries ({\em corner point grids}), tubular wells etc., and operates in
+general on a geometric model that is not trivial to reduce in complexity. The
+model is typically also very large, meaning that rendering the full model on a
+thin client is prohibitive. With the automatic proxy geometry, we can obtain
+interactive frame rates with a limited connection from a very lightweight client
+to the server. For a comparison of a server-rendered image and a client-rendered
+proxy model that is slightly rotated on the client, see Figure~\ref{fig:FRView}.
 
@@
 %-------------------------------------------------------------------------
+\section{Midlertidige notater}
+
+\begin{itemize}
+
+\item We restrict ourselves to discuss ``splat sets'' in this paper. (Nevne der det
+tas frem andre alternativer)
+
+\item
+Husk aa faa med det viktige i at vi faar oekt serverkapasiteten til aa
+serve flere klienter ogsaa!
+
+\item
+``server-image'' i steden for ``server-rendered image'' for aa spare plass?
+
+\item received stavekontroll
+
+\item Hvor passer det aa faa inn 1) bilde og 2) tekst om FRView?
+
+\item Maa ta et lite soek paa prior art innen omraadet
+
+\item problem med shading/shadows ++ i forb. med ikke deferred shading/normal map
+
+\item About our test case: corners, smooth surfaces, texturing for which artifacts are easy to
+spot. What it does lack is self-occluding parts. The FRView rservoir model that
+we have tested on does contain some of this.
+
+\item Burde vi vise chemco?
+
+\item Sjekk konsistens ved bruk av kombinasjoner som ``server rendered'' og
+``server-rendered'' etc.
+
+\item Referanser? Prior work?
+
+\end{itemize}
+
+%-------------------------------------------------------------------------
 

### 2014-12-08 tex/short_2015/body.tex
more text

@@
 
-\title[Automatic Proxy Geometry]%
-      {Automatic Proxy Geometry}
+%\title[Automatic Proxy Geometry Generation]%
+%      {Automatic Proxy Geometry Generation}
+
+\title[Automatic Proxy Model Generation]%
+      {Hybrid Client/Server Rendering with Automatic Proxy Model Generation}
 
@@
         and J. Mikkelsen$^{1}$
-        and Husk Anonymisering Hvis Paakrevet!$^{1}$
+        and Husk Anonymisering Hvis Paakrevet! \textbf{JA DET ER DET}$^{1}$
 %        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
@@
 % For Computer Graphics Forum: Please use the abbreviation of your first name.
-         $^1$SINTEF ICT, applied mathematics, Norway
+%         $^1$SINTEF ICT, applied mathematics, Norway
 %        $^2$ Another Department to illustrate the use in papers from authors
@@
 
-% \teaser{
-%  \includegraphics[width=\linewidth]{eg_new}
-%  \centering
-%   \caption{New EG Logo}
-% \label{fig:teaser}
-% }
+\teaser{
+  \centering
+  \subfigure[Server-rendered (OpenGL)]{
+    \includegraphics[width=.2\linewidth]{frview1.jpg}
+  }
+  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts not subdued)]{
+    \includegraphics[width=.2\linewidth]{frview2.jpg}
+  }
+  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts subdued)]{
+    \includegraphics[width=.2\linewidth]{frview2.jpg}
+  }
+  \caption{\label{fig:FRView} The oil reservoir viewer FRView, developed in the
+  CloudViz project~\cite{cloudviz}, showing a server-rendered image and two
+  client-rendered images of automatically generated proxy models. For~(b),
+  parameters are chosen to highlight some artifacts (few and large splats
+  without intra-splat texturing, displaying of skewed splats) while the second
+  client-rendered image (c) has parameters chosen to make an optimal proxy
+  geometry (many more splats, intra-splat texturing, discarding of skewed
+  splats.)}
+}
 
@@
 \begin{abstract}
-   We describe an algorithm and implementation of an automatic proxy geometry
-   system for a server/client remote rendering setup. The client may be a thin
-   client in the form of a hand-held pad, for our implementation we only assume
-   the availability of WebGL and Javascript. For our rendering server, we make
-   use of a GL-capable web server. In situations where the rate of received
+   We describe an algorithm and implementation of an automatic proxy model
+   generation system for a client/server remote rendering setup. In this
+   context, a {\em proxy model}, is a light-weight replacement of a main model
+   that is much easier to manipulate and render. This proxy model is
+   automatically derived from the main model. The client may be a {\em thin
+   client}, \eg, a hand-held tablet. For our implementation we only assume the
+   availability of WebGL and Javascript on this client. On the server side we
+   make use of OpenGL and a web server. In situations where the rate of received
    images deteriorate due to low bandwith, high latency or long rendering times,
-   the client displays an automatically generated proxy geometry. This proxy
-   geometry is computed from depth buffers bundled with rendered images from the
-   server, and requires no modification of the application or its data itself.
+   the client displays the proxy model. The proxy geometry is computed from
+   depth buffers bundled with rendered images from the server, and the algorithm
+   is non-intrusive in that it requires no modification of the application or
+   its data itself.
    % Husk: Skal vaere en blank linje nedenfor
@@
 \begin{classification} % according to http://www.acm.org/class/1998/
-\CCScat{FIX THIS Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
+\CCScat{\textbf{FIX THIS}Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
 \end{classification}
@@
 \section{Introduction}
-
-\begin{figure}[htb]
-  \centering
-  \subfigure[Server-rendered (OpenGL)]{
-    \includegraphics[width=.47\linewidth]{frview1.jpg}
-  }
-  \subfigure[Client-rendered proxy model (Google Chrome)]{
-    \includegraphics[width=.47\linewidth]{frview2.jpg}
-  }
-  \caption{\label{fig:FRView} The oil reservoir viewer FRView, showing a
-  server-rendered image and a proxy geometry generated from this and the
-  accompanying depth buffer. Note that we have chosen parameters for the proxy
-  generation (small number of splats ($128^2$), forcing display of splats that
-  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
-  the ragged edges, a few holes, and the oddly placed (and rendered) axis
-  indicator.}
-\end{figure}
 
@@
 computation and storage capabilities, is that interactivity still is a difficult
-issue. In this work, we consider a server/client model for rendering, but this
+issue. In this work, we consider a client/server model for rendering, but this
 is just a way of separating the interaction from the rendering; the discussion
@@
 
-One way to deal with such situations is for the client to have a {\em
-proxy model}, defined as a temporary model to be shown and manipulated
-locally on the client while waiting for the appropriate image from the server.
-A particularly unwelcomed chore is that of building proxy geometries for various
-data sets. The typical ``hello world'' proxy geometry is the familiar bounding
-box wire frame, but this will degrade the user experience unacceptably for most
+%\begin{figure}[htb]
+%  \centering
+%  \subfigure[Server-rendered (OpenGL)]{
+%    \includegraphics[width=.47\linewidth]{frview1.jpg}
+%  }
+%  \subfigure[Client-rendered proxy model (Google Chrome)]{
+%    \includegraphics[width=.47\linewidth]{frview2.jpg}
+%  }
+%  \caption{\label{fig:FRView} The oil reservoir viewer FRView \textbf{FIKS HER ref}, showing a
+%  server-rendered image and a proxy geometry generated from this and the
+%  accompanying depth buffer. Note that we have chosen parameters for the proxy
+%  generation (small number of splats ($128^2$), forcing display of splats that
+%  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
+%  the ragged edges, a few holes, and the oddly placed (and rendered) axis
+%  indicator. \textbf{Flytte opp til teaser, tredje bilde: uten artifacter for aa
+%  vise at det kan bli bra}}
+%\end{figure}
+
+One way to deal with such situations is for the client to have a {\em proxy
+model}, defined as a temporary model to be shown and manipulated locally on the
+client while waiting for the appropriate image from the server.  A particularly
+unwelcomed chore is that of building proxy geometries for various data sets. The
+typical ``hello world'' proxy geometry is the familiar bounding box wire frame,
+but this will degrade the user experience unacceptably for most
 applications. One alternative is to hand craft such simpler proxies. They have
@@
 this approach is that this is not easy to do for many kinds of data, like our
-main case in which we have an oil reservoir viewer (FRView) that renders large
-{\em corner points grids}, together with {\em faults}, oil wells, and more. See
-Figure~\ref{fig:FRView} above for an example of both a fully completed
-rendering, and an automatically generated proxy geometry rendered in Google
-Chrome.
+main case in which we have an oil reservoir viewer (FRView,~\cite{cloudviz})
+that renders large {\em corner point grids}, together with faults, oil wells,
+and more. See Figure~\ref{fig:FRView} above for an example of both a fully
+completed rendering, and automatically generated proxy geometries rendered in
+Google Chrome.
 
@@
 suitable scaling properties may still be achieved without increasing server
-capacity. \textbf{referanse til ArrowHead her?}
+capacity. The results described in this paper will be commercialized in the new
+SINTEF project \mbox{ArrowHead}~\cite{arrowhead}.
 
 \subsection{Previous work}
+\label{sec:prevWork}
 
 Our approach has some similarities to {\em image based rendering} (IBR)
-techniques (\textbf{referanse til noe av dette?}), with the difference that an
-important IBR problem would be the reconstruction of a depth map from images,
-while we have access to the full depth map from a rendering pass.
-Another way to use depth maps similar to what we do, is for ``immersive
-streaming'',
-see \eg, \texttt{http://www.cc...},
-% http://www.cc.gatech.edu/conferences/3DPVT08/Program/Papers/paper203.pdf
-where focus is on depth map compression, an issue that we also must confront.
+techniques, with the difference that an important IBR problem would be the
+reconstruction of a depth map from images, while we have access to the full
+depth map from a rendering pass.  Another way to use depth maps similar to what
+we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
+depth map compression, an issue that we also consider.
+
+Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
+point clouds. We could also in principle do this for our 2.5D maps fetched from
+the server, but it is not clear that the benefit would outweight the
+cost. \marginpar{\textbf{REF?!}}
 
@@
 while we use an existing depth map to distill and render temporary geometry.
+
+Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
+(time-of-flight and other range image sensors) produce, but most authors
+considering these are typically building more complex geometries before
+visualization. % , see~\eg,~\cite{}
+\marginpar{\textbf{REF?!}}
 
@@
 The main idea is to automatically construct a proxy geometry on the client side
-from depth data provided by the server. For a new image to be displayed on the
-client, the server renders an image in a frame buffer, and as part of this
-process it also renders a corresponding depth buffer. (This means that our
-technique is not immediately ready for \eg, ray-tracing.) This depth buffer may
-be regarded as a height map over a viewing plane relative to the observing
-client. Hence, this depth map does not contain information about scene elements
-not in view for the set of view parameters at the time of server side
-rendering. Our approach is based on the assumption that small transformations of
-this height map still will give good approximations to the scene being
-rendered. These renderings are exactly what the client will compute and show
-while waiting for the next update from the server. See
-Figure~\ref{fig:2Dheightmap} and Figure~\ref{fig:2DheightmapRotated} below for an
-illustration of this heightmap computed on the server, and a recovered 3D model
-on the client.
+from depth data provided by the server. The server already produces a depth
+buffer as part of its rendering process.  This means that our technique is not
+immediately ready for \eg, ray-tracing.) This depth buffer may be regarded as a
+height map over a viewing plane relative to the observing client. Hence, this
+depth map does not contain information about scene elements not in view for the
+set of view parameters at the time of server side rendering. Our approach is
+based on the assumption that small transformations of this height map still will
+give good approximations to the scene being rendered. These renderings are
+exactly what the client will compute and show while waiting for the next update
+from the server. See Figure~\ref{fig:2Dheightmap} and
+Figure~\ref{fig:2DheightmapRotated} below for an illustration of this heightmap
+computed on the server, and a recovered 3D model on the client.
 
@@
   \input{fig2.tex}
-
   \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
@@
 
-Further compression methods (\textbf{passende med en ref her?}) may be employed
-to bring this data size down even more, but then the cost of the compression
-itself must also be added to the cost of rendering on the server.
-%
-%The cost must of course be considered if the operation of packaging images for
-%sending to the client is significant, which may be the case if, \eg, images must
-%be compressed or converted to some special format. An example of the latter
-%might be {\texttt base64}-encoding for converting the image to a purely ASCII
-%format.
-%
-Together with the rendered rgb image and the depth image we will also send the
-current view transformations back to the client.
+Further compression may bring this data size down even more, but the cost of
+compression/decompression must also be added to the cost of rendering on the
+client/server. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
+proposed that promises to be fast both for the compression and decompression
+stages. \marginpar{\textbf{REF?!}} Together with the rendered rgb image and the
+depth image we will also send the current view transformations back to the
+client.
 
@@
 
+\marginpar{\textbf{boer kunne krympes}}
 As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
@@
 operate with a smaller number of 3D points, and instead render each of them with
-a larger number of pixels on the client side, this is called a {\em
-splat}. A set of such splats for a given depth image, we will call a {\em proxy
+a larger number of pixels on the client side, this is called a {\em splat}. A
+set of such splats for a given depth image is what we have termed a {\em proxy
 model}. There are some issues to discuss in relation to this, \eg, what size and
 shape of geometries to render for each 3D point, how to color them, how to
-compute depth fragments and so on.
+compute depth fragments and so on.  By rendering splats, we get fewer ``false
+connections'' than if we connect 3D points and reconstruct topology, for
+instance in the shape of a triangulation. On the other hand, we risk getting
+more ``holes'' in our models.
 
@@
 case, a local 2D texture transformation will provide a good approximation to the
-intra-splat texturing to be performed on the client. For the splat centered in
-$\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be centered on the client's canvas at
-screen coordinate $\pv_{i, j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i,
-j}$, the texture coordinate transformation to be used is,
+intra-splat texturing to be performed on the client. Let $P_c$ and $P_s$ be
+projection matrices on the client and server, respectively, and $M_c$ and $M_s$
+corresponding view matrices. Let there be $n_x \times n_y$ splats laid out in a
+grid. For the splat centered in $\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be
+centered on the client's canvas at screen coordinate $\pv_{i, j} = P_c M_c
+M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
+transformation to be used is,
 \[
@@
   \begin{pmatrix}
-    \frac{1}{\text{splats}_x} & 0 \\
-    0 & \frac{1}{\text{splats}_y}
+    \frac{1}{n_x} & 0 \\
+    0 & \frac{1}{n_y}
   \end{pmatrix}
@@
   \begin{pmatrix} 
-    \frac{\text{vpwidth}}{\text{splats}_x} & 0 \\
-     0 & \frac{\text{vpheight}}{\text{splats}_y}
+    \frac{\text{vpwidth}}{n_x} & 0 \\
+     0 & \frac{\text{vpheight}}{n_y}
   \end{pmatrix}
@@
 2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
-y+\Delta y) - 1$ are depths sampled from the received depth buffer and
-transformed to $[-1, 1]$. By combining this with a $w$-component set to one, we
-have in effect done a perspective division, so that the right-hand side vertices
-are in clip space, suitable for multiplication with $U$.
+y+\Delta y) - 1$ are depths sampled from the received depth buffer
+$\text{depthTex}$ and transformed to $[-1, 1]$. By combining this with a
+$w$-component set to one, we have in effect done a perspective division, so that
+the right-hand side vertices are in clip space, suitable for multiplication with
+$U$.
 
@@
 \begin{equation}
-  \Delta x = \text{splats}_x / \text{DepthImg-width}
+  \begin{aligned}
+    \Delta x & = n_x / \text{width}(\text{depthTex}) \\
+    \Delta y & = n_y / \text{height}(\text{depthTex}),
+  \end{aligned}
   \label{eq:deltaChoice}
 \end{equation}
-\[
-  \Delta y = \text{splats}_y / \text{DepthImg-height}.
-\]
-\textbf{fiks de likningene over}
-Note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
+note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
 chosen something larger. The values are used for looking up depth image samples,
@@
 \[
-  \sv_{\Delta x} =
-  \sv_{x+\Delta x} - \sv =
-    \frac{\text{width}}{2\Delta x} \left(
-        \frac{\pv_{x+\Delta x}.xy}{\pv_{x+\Delta x}.w} -
+  \sv_{\delta} =
+  \sv_{x+\delta} - \sv =
+    \frac{L}{2\delta} \left(
+        \frac{\pv_{x+\delta}.xy}{\pv_{x+\delta}.w} -
         \frac{\pv.xy}{\pv.w}
@@
 \]
-and
-\[
-  \sv_{\Delta y} =
-  \sv_{y+\Delta y} - \sv =
-    \frac{\text{height}}{2\Delta y} \left(
-        \frac{\pv_{y+\Delta y}.xy}{\pv_{y+\Delta y}.w} -
-        \frac{\pv.xy}{\pv.w}
-    \right),
-\]
-where we have used the ``shader notation'' to extract 2D vectors and scalars
-from the 4D vectors, and $\text{width}$ and $\text{height}$ are the viewport
-width and height, respectively.
+where $\delta$ is either $\Delta x$ or $\Delta y$, $L$ is either
+$\text{width}$ or $\text{height}$, and we have used the ``shader
+notation'' to extract 2D vectors and scalars from the 4D vectors, and
+$\text{width}$ and $\text{height}$ are the viewport width and height,
+respectively.
 
@@
   \end{pmatrix}
+  =
+  \begin{pmatrix}
+    s \\ t
+  \end{pmatrix} +
+  T 
+  \begin{pmatrix}
+    \texttt{glPointCoord}-\frac{1}{2}
+  \end{pmatrix}
 \]
-where $(u, v)^\prime = \texttt{gl\_PointCoord}-0.5$ are ``intra-splat texture
-coordinates''. \textbf{Boer kanskje sjekke at det ikke er noen faktor 2 som har
-falt ut noe sted paa veien...}
+where $(u, v)$ are ``intra-splat texture coordinates''.
 
@@
 implementation.
+
+%-------------------------------------------------------------------------
+\subsubsection{Other splat considerations}
+
+\marginpar{\textbf{... samle avsnittene nedenfor og komprimere...}}
 
@@
 model. The second simply replaces the oldest one in store. The third replaces a
-proxy model $j$ if the replacement results in the following objective function
+proxy model $k$ if the replacement results in the following objective function
 being reduced,
 \[
-  \text{coverage}_j = 
-  \sum_{i=1, i\neq j}^n 
-    \angle(\textbf{camdir}_i, \textbf{new camdir} )^2,
+  \text{coverage} =
+  \sum_{i=1}^n 
+    \sum_{j\neq i}^n 
+      ( \angle(\textbf{camdir}_i, \textbf{camdir}_j )^2,
 \]
-where $\textbf{camdir}$ is the direction in which the camera was looking for the
-generation of that particular proxy model.
-
-There are some issues to consider with respect to panning and zooming that we
-will not discuss here due to the limited space available.  All of these model
-replacement algorithms work decently well.
+where $n$ is the number of proxy models available and the model to be replaced
+is the one that minimizes this after replacement.
+%\textbf{droppe fra her: }
+%\ie, if $\text{coverage}_k < \text{coverage}$, where
+%\[
+%  \text{coverage}_k =
+%  \min_j
+%  \sum_{i=1, i\neq j}^n 
+%    ( \angle(\textbf{camdir}_i, \textbf{new\ camdir}_j )^2.
+%\]
+%\textbf{til hit?}
+Here, $\textbf{camdir}$ is the direction in which the camera was looking for the
+generation of that particular proxy model.  There are some issues to consider
+with respect to panning and zooming that we will not discuss here due to the
+limited space available.
 
 %-------------------------------------------------------------------------
-\subsection{Dynamic/heuristic method selection, depth compression}
+\subsection{Dynamic/heuristic method selection, depth compression, auto-tuning}
+\label{sec:autoTuning}
 
@@
 sending proxy model data from the server itself is not hold up more than
-necessary. There are mainly three sources of delay (\textbf{passer aa nevne
-tidligere?}) for server-rendered images to the client; high
-latency, low bandwidth, and slow server-rendering itself.
+necessary. There are mainly three sources of delay for server-rendered images to
+the client; high latency, low bandwidth, and slow server-rendering itself.
 
@@
 with the client, as well as a reduced resolution of the depth buffer sent from
-the server. See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
-high-quality proxy models.
+the server. Another possible form of auto-tuning is to let the client adaptively
+choose the number of splats, stored number of proxy models, etc.
 
-\begin{figure}[htb]
-  \centering
-  \subfigure[Full png image,\newline full depth buffer]{
-    \includegraphics[width=.30\linewidth]{auto1.jpg}
-  }
-  \subfigure[Coarse jpg,\newline quality $q=0$]{
-    \includegraphics[width=.30\linewidth]{auto2.jpg}
-  }
-  \subfigure[Auto-proxy, trun-\newline cated depth data]{
-    \includegraphics[width=.30\linewidth]{auto3.jpg}
-  }
-  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
-           available bandwidth and server capacity. The server-rendered JPG
-           image adaptively selects a quality setting based on recorded
-           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
-           the last step before switching to ``auto proxy'' (c). The auto-proxy
-           model shown displays artifacts caused both by the magenta surface
-           being almost parallel to the viewing direction at the moment of proxy
-           model recording, and by the non-smooth parts of the geometry
-           (corners). The number of splats is set low on purpose, and the depth
-           buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$.}
-\end{figure}
+%See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
+%high-quality proxy models.
+%\begin{figure}[htb]
+%  \centering
+%  \subfigure[Full png image,\newline full depth buffer]{
+%    \includegraphics[width=.30\linewidth]{auto1.jpg}
+%  }
+%  \subfigure[Coarse jpg,\newline quality $q=0$]{
+%    \includegraphics[width=.30\linewidth]{auto2.jpg}
+%  }
+%  \subfigure[Auto-proxy, trun-\newline cated depth data]{
+%    \includegraphics[width=.30\linewidth]{auto3.jpg}
+%  }
+%  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
+%           available bandwidth and server capacity. The server-rendered JPG
+%           image adaptively selects a quality setting based on recorded
+%           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
+%           the last step before switching to ``auto proxy'' (c). The auto-proxy
+%           model shown displays artifacts caused both by the magenta surface
+%           being almost parallel to the viewing direction at the moment of proxy
+%           model recording, and by the non-smooth parts of the geometry
+%           (corners). The number of splats is set low on purpose, and the depth
+%           buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$. \textbf{KRYMPE HER}}
+%\end{figure}
 
@@
 
-The automatic proxy geometry implementation is fully invisible to the
+The automatic proxy geometry implementation in Tinia is fully invisible to the
 application, meaning that all existing applications immediately will have the
-feature available. We have tested the algorithm on several smaller test cases,
-but also on a larger oil reservoir viewer, FRView, developed in the CloudViz
+feature available. Without Tinia, the algorithm is minimally intrusive in that
+only the depth buffer will have to be added to the rendering output of the
+application. We have tested the algorithm on several smaller test cases, but
+also on a larger oil reservoir viewer, FRView, developed in the CloudViz
 project, see~\cite{cloudviz}.
@@
 to the server. For a comparison of a server-rendered image and a client-rendered
-proxy model that is slightly rotated on the client, see Figure~\ref{fig:FRView}.
+proxy model that is slightly rotated on the client, see again Figure~\ref{fig:FRView}.
 
@@
 There are several directions in which we would like to follow up and improve
-this concept,
+this concept, foremost of these are,
 \begin{itemize}
+
 \item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
@@
 proxy models.
-\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
-envision a spatially partioned grid, for instance containing a distance field,
-with the geometry as an iso-surface.
-\end{itemize}
 
-%-------------------------------------------------------------------------
-\section{Midlertidige notater}
+%\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
+%envision a spatially partioned grid, for instance containing a distance field,
+%with the geometry as an iso-surface.
 
-\begin{itemize}
+%\item \textbf{Client rendering with overlays} Items like, \eg, axis indicators,
+%legends etc. that do not use the same view matrix as the main model could be
+%rendered locally on the client, only. This will cancel some of the benefit
+%connected to the automatic proxy model generation, though. \textbf{Nettopp av
+%denne grunn kan kanskje dette vaere et punkt vi boer luke bort? Plassmangel
+%ogsaa...}
 
-\item We restrict ourselves to discuss ``splat sets'' in this paper. (Nevne der det
-tas frem andre alternativer)
+\item \textbf{Depth compression} We have currently done a unsophisticated
+truncation, but other approaches are possible,
+like discussed in~\cite{DBLP:journals/tvcg/Lindstrom14}.
 
-\item
-Husk aa faa med det viktige i at vi faar oekt serverkapasiteten til aa
-serve flere klienter ogsaa!
+%\item \textbf{Improved auto-tuning} The auto-tuning discussed in
+%Section~\ref{sec:autoTuning} should be extended to take not only server
+%parameters (rendering quality, compression etc.) into account, but also the
+%capabilities of various thin clients.
 
-\item
-``server-image'' i steden for ``server-rendered image'' for aa spare plass?
+\item \textbf{Deferred shading} In addition to sending an image and a depth map
+from the server, one could send also a normal map. (This can also be
+reconstructed from the depth map on the client.) This would allow more advanced
+shading on the client.
 
-\item received stavekontroll
-
-\item Hvor passer det aa faa inn 1) bilde og 2) tekst om FRView?
-
-\item Maa ta et lite soek paa prior art innen omraadet
-
-\item problem med shading/shadows ++ i forb. med ikke deferred shading/normal map
-
-\item About our test case: corners, smooth surfaces, texturing for which artifacts are easy to
-spot. What it does lack is self-occluding parts. The FRView rservoir model that
-we have tested on does contain some of this.
-
-\item Burde vi vise chemco?
-
-\item Sjekk konsistens ved bruk av kombinasjoner som ``server rendered'' og
-``server-rendered'' etc.
-
-\item Referanser? Prior work?
+%\item \textbf{Stitching of 2.5D point clouds} As mentioned in
+%Section~\ref{sec:prevWork}, we have considered but not experimented with
+%stitching of depth maps from the server.
 

### 2014-12-09 tex/short_2015/body.tex
Before serious shrinking

@@
   \subfigure[Server-rendered (OpenGL)]{
-    \includegraphics[width=.2\linewidth]{frview1.jpg}
+    \includegraphics[width=0.3\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
   }
-  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts not subdued)]{
-    \includegraphics[width=.2\linewidth]{frview2.jpg}
+  \subfigure[Client-rendered (WebGL)]{
+    \includegraphics[width=0.3\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
   }
-  \subfigure[Client-rendered proxy model (Google\newline Chrome, artifacts subdued)]{
-    \includegraphics[width=.2\linewidth]{frview2.jpg}
+  \subfigure[Client-rendered (WebGL)]{
+    \includegraphics[width=0.3\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
+    % The trim params: left lower right upper
   }
-  \caption{\label{fig:FRView} The oil reservoir viewer FRView, developed in the
-  CloudViz project~\cite{cloudviz}, showing a server-rendered image and two
-  client-rendered images of automatically generated proxy models. For~(b),
-  parameters are chosen to highlight some artifacts (few and large splats
-  without intra-splat texturing, displaying of skewed splats) while the second
-  client-rendered image (c) has parameters chosen to make an optimal proxy
-  geometry (many more splats, intra-splat texturing, discarding of skewed
-  splats.)}
+  \caption{\label{fig:FRView} The oil reservoir viewer FRView~\cite{cloudviz},
+  showing one server-rendered and two client-rendered images of automatically
+  generated and slightly rotated proxy models.  For~(b), parameters are chosen to produce the best
+  possible image, and in~(c) we want to highlight artifacts and implementational
+  details. See the text for further discussion of these.}
 }
@@
 \begin{abstract}
-   We describe an algorithm and implementation of an automatic proxy model
-   generation system for a client/server remote rendering setup. In this
-   context, a {\em proxy model}, is a light-weight replacement of a main model
-   that is much easier to manipulate and render. This proxy model is
-   automatically derived from the main model. The client may be a {\em thin
-   client}, \eg, a hand-held tablet. For our implementation we only assume the
-   availability of WebGL and Javascript on this client. On the server side we
-   make use of OpenGL and a web server. In situations where the rate of received
-   images deteriorate due to low bandwith, high latency or long rendering times,
-   the client displays the proxy model. The proxy geometry is computed from
-   depth buffers bundled with rendered images from the server, and the algorithm
-   is non-intrusive in that it requires no modification of the application or
-   its data itself.
+\color{red}
+   We describe an algorithm and implementation of automatic proxy model
+   generation for a client/server remote rendering setup. A {\em proxy model} is
+   a light-weight version of a main 3D model that is cheaper to render. We
+   automatically derive this from the main model. The client is typically a
+   tablet, for which we only assume availability of WebGL and Javascript. The
+   server makes use of OpenGL and a web server. When the rate of received
+   server-rendered images deteriorate, the client renders the proxy model, which
+   is computed from depth buffers bundled with rendered images from the
+   server. This algorithm requires little modification of the application
+   itself.
    % Husk: Skal vaere en blank linje nedenfor
@@
 \begin{classification} % according to http://www.acm.org/class/1998/
-\CCScat{\textbf{FIX THIS}Computer Graphics}{I.3.3}{Picture/Image Generation}{Line and curve generation}
+\CCScat{Computer Graphics}{I.3.5}{Computational Geometry and Object Modeling}{Geometric algorithms, languages, and systems}
+% http://www.acm.org/about/class/ccs98-html
 \end{classification}
@@
 with the user experience.
-
-%\begin{figure}[htb]
-%  \centering
-%  \subfigure[Server-rendered (OpenGL)]{
-%    \includegraphics[width=.47\linewidth]{frview1.jpg}
-%  }
-%  \subfigure[Client-rendered proxy model (Google Chrome)]{
-%    \includegraphics[width=.47\linewidth]{frview2.jpg}
-%  }
-%  \caption{\label{fig:FRView} The oil reservoir viewer FRView \textbf{FIKS HER ref}, showing a
-%  server-rendered image and a proxy geometry generated from this and the
-%  accompanying depth buffer. Note that we have chosen parameters for the proxy
-%  generation (small number of splats ($128^2$), forcing display of splats that
-%  would normally be discarded) to emphasize rather than reduce artifacts, \eg,
-%  the ragged edges, a few holes, and the oddly placed (and rendered) axis
-%  indicator. \textbf{Flytte opp til teaser, tredje bilde: uten artifacter for aa
-%  vise at det kan bli bra}}
-%\end{figure}
 
@@
 we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
-depth map compression, an issue that we also consider.
+depth map compression, an issue that we also consider. {\color{red}Another work
+focused on similar streaming and geometry compression,
+is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
+work, with respect to image-based rendering accelleration.}
 
@@
 the server, but it is not clear that the benefit would outweight the
-cost. \marginpar{\textbf{REF?!}}
+cost.
 
@@
 
-Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
+{\color{red}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
 (time-of-flight and other range image sensors) produce, but most authors
 considering these are typically building more complex geometries before
-visualization. % , see~\eg,~\cite{}
+visualization.} % , see~\eg,~\cite{}
 \marginpar{\textbf{REF?!}}
@@
 
-\marginpar{\textbf{... samle avsnittene nedenfor og komprimere...}}
+{\color{red} ... samle avsnittene nedenfor og komprimere...}
 
@@
 to the server. For a comparison of a server-rendered image and a client-rendered
-proxy model that is slightly rotated on the client, see again Figure~\ref{fig:FRView}.
+proxy model that is slightly rotated on the client, see again
+Figure~\ref{fig:FRView}. {\color{red}In Figure~\ref{fig:FRView}~(a), the full
+server-rendered image is shown, and in~(b) and~(c), a slightly (about 10
+degrees) rotated proxy model is rendered on the client. In~(b), parameters are
+chosen for best possible results, while in~(c), we want to highlight some
+effects. For the latter, we have chosen a small number of non-overlapping
+splats, which are also outlined. In this image, it is easy to spot areas not
+well covered by any splats, and also some areas where the texture coordinate
+transform $T$ does not produce optimal results.}
 

### 2014-12-09 tex/short_2015/body.tex
Now 4 pages

@@
 \begin{abstract}
-\color{red}
+\color{black}
    We describe an algorithm and implementation of automatic proxy model
@@
 
-With increasingly powerful rendering engines, in forms ranging from mobile and
-hand-held devices to traditional stationary computers, one might be led to
-believe that rendering frame rates could only increase and visualization could
-only become more interactive. Alas, hand in hand with this development we also
-find ever increasing requirements on computational accuracy, power efficiency,
-data sizes, scaling, etc.
+{\color{black} Hand in hand with increasingly powerful rendering engines comes
+ever increasing requirements on computational accuracy, power efficiency, data
+sizes, scaling, etc. Also enhanced by new cloud-based approaches etc., an effect
+is that interactivity still is a difficult issue. We consider a client/server
+model for 3D rendering, addressing latency, bandwidth and scaling problems in a
+novel way.
 
-One effect of this, also enhanced by new cloud-based approaches, wireless
-connection, and mobile devices with display capabilities outpacing their
-computation and storage capabilities, is that interactivity still is a difficult
-issue. In this work, we consider a client/server model for rendering, but this
-is just a way of separating the interaction from the rendering; the discussion
-also applies to a more integrated setup.  In order to obtain the frame rates on
-the client suitable for interactive visualizations, demands are put on
-bandwidth, latency and rendering speed. If one or more of these demands are not
-met, according to a suitable measure, the frame rate will deteriorate together
-with the user experience.
-
-One way to deal with such situations is for the client to have a {\em proxy
-model}, defined as a temporary model to be shown and manipulated locally on the
-client while waiting for the appropriate image from the server.  A particularly
-unwelcomed chore is that of building proxy geometries for various data sets. The
-typical ``hello world'' proxy geometry is the familiar bounding box wire frame,
-but this will degrade the user experience unacceptably for most
-applications. One alternative is to hand craft such simpler proxies. They have
-to be simple both for the building cost, and also for the actual rendering on
-the possibly very thin client. Another alternative is to automate this process,
-a large triangulation may for instance be represented by a much coarser
-approximating triangulation that can be built automatically. One disadvantage of
-this approach is that this is not easy to do for many kinds of data, like our
-main case in which we have an oil reservoir viewer (FRView,~\cite{cloudviz})
-that renders large {\em corner point grids}, together with faults, oil wells,
-and more. See Figure~\ref{fig:FRView} above for an example of both a fully
-completed rendering, and automatically generated proxy geometries rendered in
-Google Chrome.
-
-Our suggested solution is to pass depth information from the server along with
-ordinary rendered images. From this depth image, a rudimentary 3D model can be
-built, and using the ordinary image as a texture, this model can be transformed
-and rendered on the client while waiting for the next update from the server. If
-the server does not take too long to respond, and/or the client does not change
-the position or orientation of the model too much, this display of a proxy model
+We introduce a {\em proxy model}, defined as a temporary model to be shown and
+manipulated locally on the client while waiting for the appropriate image from
+the server. Producing such proxy models can be difficult for many kinds of 3D
+data, like in our main case in which we have an oil reservoir viewer
+(FRView,~\cite{cloudviz}) that renders large {\em corner point grids}, together
+with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
+example of both a fully completed rendering, and automatically generated proxy
+geometries rendered in Google Chrome.
+%
+Our solution is to pass depth information from the server along with ordinary
+rendered images. From this, a rudimentary 3D model is built, and with the RGB
+image as a texture, this model can be manipulated and rendered on the client
+while waiting for the next update from the server. If the client does not change
+the position or orientation of the model too much, this proxy model rendering
 integrates seemlessly with the slower stream of server-rendered frames.
-
-We also mention that even if bandwidth and latency is not a problem, it may be
-desirable to let a server of limited capacity serve many simultaneous users,
-hence decreasing the effective server time available for each one. In this case,
-suitable scaling properties may still be achieved without increasing server
-capacity. The results described in this paper will be commercialized in the new
-SINTEF project \mbox{ArrowHead}~\cite{arrowhead}.
+%
+Even if bandwidth and latency is not a problem, it may be desirable to let a
+server of limited capacity serve many simultaneous users, hence limiting the
+effective server time available for each one. Suitable scaling may still be
+achieved using our solution. This will be commercialized in the new SINTEF
+project \mbox{ArrowHead}~\cite{arrowhead}.}
 
@@
 we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
-depth map compression, an issue that we also consider. {\color{red}Another work
+depth map compression, an issue that we also consider. {\color{black}Another work
 focused on similar streaming and geometry compression,
@@
 work, with respect to image-based rendering accelleration.}
-
+%
 Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
@@
 cost.
-
+%
 We are not aware of anybody else using depth maps exactly for our purpose, most
@@
 while we use an existing depth map to distill and render temporary geometry.
-
-{\color{red}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
+%
+{\color{black}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
 (time-of-flight and other range image sensors) produce, but most authors
@@
 
-The main idea is to automatically construct a proxy geometry on the client side
-from depth data provided by the server. The server already produces a depth
-buffer as part of its rendering process.  This means that our technique is not
-immediately ready for \eg, ray-tracing.) This depth buffer may be regarded as a
-height map over a viewing plane relative to the observing client. Hence, this
-depth map does not contain information about scene elements not in view for the
-set of view parameters at the time of server side rendering. Our approach is
-based on the assumption that small transformations of this height map still will
-give good approximations to the scene being rendered. These renderings are
-exactly what the client will compute and show while waiting for the next update
-from the server. See Figure~\ref{fig:2Dheightmap} and
-Figure~\ref{fig:2DheightmapRotated} below for an illustration of this heightmap
-computed on the server, and a recovered 3D model on the client.
-
-\begin{figure}[htb]
-  \centering
-  \input{fig2.tex}
-  \caption{\label{fig:2Dheightmap} The rendered image (of \eg, a cube) as a
-           height map seen from above, for a certain transformation. Also shown
-           are the corresponding RGB-fragments (red and green) and depth
-           fragments (floating point scalars shown as grayscale pixels.)}
-\end{figure}
-
-In Figure~\ref{fig:2DheightmapRotated} below, a sequence of three server-rendered
-images (solid lines, red and blue) is shown, together with intermediate
-client-rendered proxy models with different features. These features will be
-discussed later, in Section~\ref{sec:client}.
+{\color{black} Since the depth buffer is a height map seen from the observer, it
+does not contain information about occluded scene elements. Our approach assumes
+that small transformations of this height map still will give good
+approximations of the scene. In Figure~\ref{fig:2DheightmapRotated} below, a
+sequence of three server-rendered images (thick lines) is shown, together with
+intermediate client-rendered proxy models with different features that will be
+discussed in Section~\ref{sec:client}.}
 
@@
   \caption{\label{fig:2DheightmapRotated}
-           The depth map and image sent from the server enables the client to
-           recover a 3D model approximating the one rendered on the server. This
-           client-side model is shown here with a small additional rotation, again
-           seen from above.}
+{\color{black} 
+Bird's view of 3D model (solid lines) and proxy model renderings.
+}
+%           The depth map and image sent from the server enables the client to
+%           recover a 3D model approximating the one rendered on the server. This
+%           client-side model is shown here with a small additional rotation, again
+%           seen from above.
+}
 \end{figure}
 
-Note that the depth map from the server does not allow the client to recover all
-information in the server side model. For instance, the solid surface shown in
-Figure~\ref{fig:2Dheightmap} (\ie, the {\em topology}) cannot be deduced, hence
-we show the dots in Figure~\ref{fig:2DheightmapRotated}~(b). Still, there are
-several things we can do to improve the client's approximation of
-the server-side image, and these will now be described.
+%Note that the depth map from the server does not allow the client to recover all
+%information in the server side model. For instance, the solid surface shown in
+%Figure~\ref{fig:2Dheightmap} (\ie, the {\em topology}) cannot be deduced, hence
+%we show the dots in Figure~\ref{fig:2DheightmapRotated}~(b). Still, there are
+%several things we can do to improve the client's approximation of
+%the server-side image, and these will now be described.
 
 %-------------------------------------------------------------------------
-\subsection{The server}
+\subsection{The server side}
 
-The server keeps a 3D model and accompanying viewing parameters, and will render
-an image for delivery to the client either on its own accord or because the
-client requests it. As long as the server renders a picture into a framebuffer
-with a connected depth buffer, it has a negligible cost for the server to bundle
-the resulting depth map with the rendered image upon return of the result. In
-practice, this is easily handled by a few extra lines of computer code where the
-image is extracted from its framebuffer and prepared for sending to the
-client.
-
-Since the depth buffer adds somewhat to the data being sent, it is important to
-keep it to a minimum. We have found that reducing the spatial resolution of the
-depth buffer (for instance by a factor of $1/16$) only to a negligible extent
-degrades the proxy model being reconstructed by the client. We encode each depth
-value in the range $[0, 1]$ as a 16 bit fixed point number, this also works
-quite well. These measures cause the bundling of the depth buffer with the rgb
-image to impose just a small data overhead.
-
-Further compression may bring this data size down even more, but the cost of
-compression/decompression must also be added to the cost of rendering on the
-client/server. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
+{\color{black}
+The server renders the 3D model into a framebuffer, and in the process generates
+a depth image that we also send to the client.
+Since this adds to the data being sent, it must be kept to a minimum. We have
+found that reducing the spatial resolution of the depth image (for instance by a
+factor of $1/16$) only degrades the proxy model imperceptibly.  We also encode
+each depth value in the range $[0, 1]$ as a 16 bit fixed point number, and the
+bundling of the depth image with the rgb image then imposes just a small data
+overhead.
+Further compression may bring this down even more, but the cost of
+compression/decompression must also be
+considered. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
 proposed that promises to be fast both for the compression and decompression
-stages. \marginpar{\textbf{REF?!}} Together with the rendered rgb image and the
-depth image we will also send the current view transformations back to the
-client.
+stages.
+}
 
 %-------------------------------------------------------------------------
-\subsection{The client}
+\subsection{The client side}
 \label{sec:client}
+
+{\color{black}
 
@@
 transformations, it builds a ``proxy model'' from this. This model can then be
-transformed and rendered directly if there is no available image from the
-server, or it can be combined with other proxy models the client already has in
-store. The latter option will be discussed in
-Section~\ref{sec:proxyModelReplacement}.
+transformed and rendered directly, or it can be combined with other proxy models
+the client already has in store, see Section~\ref{sec:proxyModelReplacement}.
 
-%-------------------------------------------------------------------------
-\subsubsection{Proxy model building, splats}
-
-\marginpar{\textbf{boer kunne krympes}}
-As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received
-height map does not really allow us more than concluding where a set of points
-belong to the 3D model, these points giving rise to exactly the depth map from
-the server. Given these points, the client may build a model approximating the
-actual scene being rendered on the server. Since the information from such a
-depth map can only contain the foremost point along any ray from the observer
-toward the scene, it is often said to be in 2.5D, as opposed to purely 2D or
-fully 3D.
-
-The most simple thing the client can do, is just to transform and render this
-set of 3D points with the color sampled from the corresponding location in the
-server-rendered image, this is illustrated in
+As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received height map
+does not allow us more than concluding where a set of points belong on the 3D
+model, \ie, we have little topologic information. Since the information from the
+depth map can only contain the foremost point along any ray from the observer,
+it is said to be in 2.5D, as opposed to 3D.  The simplest thing the client can
+do, is just to transform and render the set of 3D points with color sampled from
+the server-rendered image, as is illustrated in
 Figure~\ref{fig:2DheightmapRotated}~(b).
 
-It makes sense for the client only to render at most one 3D point for each depth
-fragment available. This may be too much for a thin client, and one may want to
-operate with a smaller number of 3D points, and instead render each of them with
-a larger number of pixels on the client side, this is called a {\em splat}. A
-set of such splats for a given depth image is what we have termed a {\em proxy
-model}. There are some issues to discuss in relation to this, \eg, what size and
-shape of geometries to render for each 3D point, how to color them, how to
-compute depth fragments and so on.  By rendering splats, we get fewer ``false
-connections'' than if we connect 3D points and reconstruct topology, for
-instance in the shape of a triangulation. On the other hand, we risk getting
-more ``holes'' in our models.
+Rendering a 3D point for each depth fragment available may tax a thin client,
+and it may be necessary to use a smaller number of primitives and instead render
+each with a larger number of pixels on the client side, a {\em splat}.  A set of
+such splats for a given depth image we call a {\em proxy model}. By rendering
+splats, we get fewer ``false connections'' than if we connect 3D points and
+reconstruct topology, but we risk getting more ``holes'' in our models.
 
-The simplest solution is to render each 3D point as a fixed geometry, for
-instance a 2D disk or rectangle, with the correspondig color from the image, see
-Figure~\ref{fig:2DheightmapRotated}~(c), where we have added the transformation
-local to the client, which the server has not been able to provide an image for.
-We will briefly describe some improvements to this. We also note that other
-possibilities exist apart from this ``splat-based'' proxy model, \eg, building
-and maintaining a 3D occupancy mesh, computing a distance field from which
-iso-surfaces can be extracted, etc.
+We can render each splat as a fixed geometry, \eg, a disk or rectangle, with the
+correspondig color from the image, see Figure~\ref{fig:2DheightmapRotated}~(c),
+where we have introduced a transformation local to the client.  We will briefly
+describe some improvements to this. Note that other possibilities include
+building and maintaining a 3D occupancy mesh, computing a distance field from
+which iso-surfaces can be extracted, etc.
+}
 
@@
 
-Our splats may result in many client window fragments, necessitating an
-``intra-splat'' fragment coloring or texturing. This is not entirely straight
-forward, since the client really do not know how a finitely-sized splat should
-map to the server's geometry, having only a discrete set of 3D points on that
-model.
-
+{\color{black}
+Each splat produces many client pixels, necessitating an ``intra-splat''
+fragment texturing for which we need a local texture coordinate transform.
 A first approximation is for the client to assume that the corresponding part of
@@
 projection matrices on the client and server, respectively, and $M_c$ and $M_s$
-corresponding view matrices. Let there be $n_x \times n_y$ splats laid out in a
-grid. For the splat centered in $\vv_{i, j} = (x_j, y_i, z_{i, j})$, to be
-centered on the client's canvas at screen coordinate $\pv_{i, j} = P_c M_c
-M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
+corresponding view matrices. For the splat centered in $\vv_{i, j} = (x_j, y_i,
+z_{i, j})$, to be centered on the client's canvas at screen coordinate $\pv_{i,
+j} = P_c M_c M_s^{-1} P_s^{-1}\vv_{i, j} = U\vv_{i, j}$, the texture coordinate
 transformation to be used is,
@@
   \begin{pmatrix} 
-    \frac{\text{vpwidth}}{n_x} & 0 \\
-     0 & \frac{\text{vpheight}}{n_y}
+    \frac{w}{n_x} & 0 \\
+     0 & \frac{h}{n_y}
   \end{pmatrix}
-  \cdot \text{overlap},
-\]
-\[
    =
@@
   \Big( \sv_x \, \, \, \sv_y \Big)^{-1}
-  B
-  \cdot \text{overlap},
+  B,
 \]
@@
 corresponding texture area, $(\sv_{x} \, \, \, \sv_{y})^{-1}$ maps the
-``client's screen space splat area'' to $[0, 1]^2$, and $B$ together with the
-``overlap'' just provides a scaling factor so that ``overlap=1'' produces
-non-overlapping but covering splats when $M_c=M_s$, \ie, the client does no
-transformation of the proxy model.
+``client's screen space splat area'' to $[0, 1]^2$, and $B$ provides a scaling
+factor to fill the {\em viewport} of size $w \times h$ with $n_x \times n_y$
+splats laid out in a grid, when $M_c=M_s$.
 %
@@
 %
-We compute $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions and
-performing the perspective division and transformation into window coordinates
-in the vertex shader. Let proxy model positions corresponding to three corners
-of a quadrilateral splat in clip coordinates be
-given by
+We obtain $\sv_{x}$ and $\sv_{y}$ by evaluating proxy model positions followed by 
+perspective division and transformation into window coordinates,
 \[
@@
 \]
-where $d = 2\text{depthTex}(x, y) - 1$, $d_{\Delta x} =
-2\text{depthTex}(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2\text{depthTex}(x,
+where $d = 2D(x, y) - 1$, $d_{\Delta x} =
+2D(x+\Delta x, y) - 1$, and $d_{\Delta y} = 2D(x,
 y+\Delta y) - 1$ are depths sampled from the received depth buffer
-$\text{depthTex}$ and transformed to $[-1, 1]$. By combining this with a
+$D$ and transformed to $[-1, 1]$. With a
 $w$-component set to one, we have in effect done a perspective division, so that
-the right-hand side vertices are in clip space, suitable for multiplication with
-$U$.
+we are in clip space and multiplication with $U$ is appropriate.
+%
+Note that $\Delta x$ and $\Delta y$ are not unique, we use
+\[
+  \Delta x = n_x / \text{width}(D) \text{\,\,\,\,and\,\,\,\,}
+  \Delta y = n_y / \text{height}(D),
+%  \label{eq:deltaChoice}
+\]
+but something larger may also be used. These are used for looking up depth image
+samples, and the calculation of $\sv_x$ and $\sv_y$ is really a gradient
+approximation, so we should not make them to large either.
 
-We use
-\begin{equation}
-  \begin{aligned}
-    \Delta x & = n_x / \text{width}(\text{depthTex}) \\
-    \Delta y & = n_y / \text{height}(\text{depthTex}),
-  \end{aligned}
-  \label{eq:deltaChoice}
-\end{equation}
-note that this choice of $\Delta x$ and $\Delta y$ is not unique, we might have
-chosen something larger. The values are used for looking up depth image samples,
-so we should not choose something too small, our choice is exactly to look for
-neighbouring depth samples. Since this calculation of $\sv_x$ and $\sv_y$ is
-really a difference approximating a gradient on the geometry, we should not make
-$\Delta x$ and $\Delta y$ to large either. Experiments show us
-that~(\ref{eq:deltaChoice}) is really a good choice.
-
-This then leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
-coordinates, and after perspective division and viewport transformation, we get
+This leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
+coordinates, and from this we get
 corresponding window coordinates $\sv$, $\sv_{x+\Delta x}$ and $\sv_{y+\Delta
-y}$, which we can subtract to obtain 2D vectors spanning the splat,
+y}$, from which we finally get splat spanning vectors
 \[
@@
 \]
-where $\delta$ is either $\Delta x$ or $\Delta y$, $L$ is either
-$\text{width}$ or $\text{height}$, and we have used the ``shader
-notation'' to extract 2D vectors and scalars from the 4D vectors, and
-$\text{width}$ and $\text{height}$ are the viewport width and height,
-respectively.
-
-Then, on the client we render a \texttt{glPoint} for each splat, with texture
-coordinates $(s, t)^\prime$ and \texttt{glPointSize} set to a larger value that will
-produce many fragments. Each fragment will then look up the server-rendered
-image at position
+with $\delta=\Delta x$ or $\delta=\Delta y$, $L$ is viewport size, either $w$ or
+$h$, and we have used the ``shader notation'' for vector components.  The client
+renders a large \texttt{glPoint} for each splat, with texture coordinates $(s,
+t)^\prime$, and each fragment then looks up the server-rendered image at
+position
 \[
@@
 \]
-where $(u, v)$ are ``intra-splat texture coordinates''.
-
-When the assumption that the geometry is locally planar does not hold, \eg, if
-the splats are very large, or they originate from a rather curved or non-smooth
-part of the geometry, this may look rather odd, see
-Figure~\ref{fig:LargeSplatsOnCorners}.
+where $(u, v)$ are ``intra-splat texture coordinates''.  When the assumption
+that the geometry is locally planar does not hold, \eg, if the splats are very
+large, or they originate from a curved or non-smooth part, this may look rather
+odd, see Figure~\ref{fig:LargeSplatsOnCorners}.
 
@@
 \end{figure}
+%\vspace{-4ex}
 
-The problem illustrated in Figure~\ref{fig:LargeSplatsOnCorners}, is that the
-geometry is far from planar, and our simple texture coordinate transformation
-will produce the wrong result for the corresponding part of involved
-splats. Some ways to remedy this is, \eg, to choose smaller (and more) splats,
-detect the problem and discard such splats, or introduce more complex texture
-transformations. Since the proxy geometry in principle is to be short-lived, and
-only a temporary geometry to be rendered before the server-rendered image can be
-receieved, we have chosen a combination of the first two alternatives. Note that
-implementing a more sophisticated texture coordinate transformation may actually
-amount to performing the same work as for more and smaller splats. Equivalently
-, using more and smaller splats may be regarded as a better texture transform
-implementation.
+When the geometry is not planar, $T$ produces the wrong result for parts of a
+splat.  Two ways to remedy this could be to choose either more and smaller
+splats, or introduce more complex texture transformations. Note that using a
+more sophisticated texture coordinate transform may amount to performing the
+same work as for more and simpler transforms. The latter may be regarded as
+exactly a better ``global'' texture transform implementation.
+}
 
@@
 \subsubsection{Other splat considerations}
-
-{\color{red} ... samle avsnittene nedenfor og komprimere...}
-
-%-------------------------------------------------------------------------
-\subsubsection{Splat sizing}
-
-Note that as soon as a given proxy model is transformed, each splat should
-ideally be rendered into a number of client pixels according to the new splats
-3D position. A splat transformed into view closer to the observer, should result
-in more client fragments, and vice versa. To achieve this, we simply use the
-vectors $\sv_{\Delta x}$ and $\sv_{\Delta y}$ from the previous section.
-In addition to this, we scale the splats up a bit, typically with a factor 2, so
-that they overlap. The benefit of this is that we can render rectangular
-window-aligned splats on the client, without getting large uncovered areas when
-interactively rotating and scaling the model quickly on the client.
-
-%-------------------------------------------------------------------------
-\subsubsection{Splat depth fragments}
-
-For larger splats, covering many fragments, it makes sense to also compute and
-use depth fragments in the client's fragment shader. This will improve the usage
-of more proxy models simultaneously, since the risk of a splat with a not very
-representative depth value covers another and better splat, is reduced. The
-``intra-splat'' depth values can easily be fetched from the depth image, just as
-the texture is looked up for color, with the same texture transformation. One
-problem is that many thin clients may not support this, in WebGL, the ``frag
-depth'' feature is an extension, which may not be supported everywhere.
-
-%------------------------------------------------------------------------
-\subsection{Splat set replacement algorithms}
 \label{sec:proxyModelReplacement}
 
-In this paper we mainly concern ourselves with proxy models defined as sets of
-splats. Since each splat set is the product of one instance of a server-rendered
-image, depth buffer and viewing parameters, it makes sense to retain more than
-one such model on the client. By doing this, the client may combine them in
-order to provide renderings for viewing parameters very different from the last
-one received.
+{\color{black}
 
-As soon as the client has a set of such proxy models, it will also need an
-algorithm for replacing old models with new ones.  One can imagine a plethora of
+\textbf{Splat sizing} 
+Each splat should be rendered into a number of client pixels according to the
+new splats 3D position. To achieve this, we use the vectors $\sv_{\Delta x}$ and
+$\sv_{\Delta y}$ from the previous section, and in addition, we scale the splats
+up a bit so that they overlap. Hence, we can render rectangular window-aligned
+splats with less risk of getting uncovered areas when interactively rotating and
+scaling the model on the client.
+
+\textbf{Splat depth fragments}
+For larger splats it makes sense to also compute and use depth fragments on the
+client. The ``intra-splat'' depth values can easily be fetched from the depth
+image, just as the texture is looked up for color. Not all clients support this
+at the time of writing, it is an extension in WebGL currently.
+
+\textbf{Splat set replacement algorithms}
+We mainly concern ourselves with proxy models defined as sets of splats, and it
+makes sense to keep a set of such models on the client, then we may combine them
+to cover larger ranges of client-side transformations. One can imagine a plethora of
 {\em splat set replacement algorithms}, we have tested three approaches that all
-retain a constant number of proxy models. The first algorithm is to replace the
+retain a constant number of proxy models. The first is to replace the
 one with a viewing direction differing the most from the newly received
@@
 being reduced,
-\[
+\begin{equation}
   \text{coverage} =
@@
       ( \angle(\textbf{camdir}_i, \textbf{camdir}_j )^2,
-\]
-where $n$ is the number of proxy models available and the model to be replaced
-is the one that minimizes this after replacement.
-%\textbf{droppe fra her: }
-%\ie, if $\text{coverage}_k < \text{coverage}$, where
-%\[
-%  \text{coverage}_k =
-%  \min_j
-%  \sum_{i=1, i\neq j}^n 
-%    ( \angle(\textbf{camdir}_i, \textbf{new\ camdir}_j )^2.
-%\]
-%\textbf{til hit?}
-Here, $\textbf{camdir}$ is the direction in which the camera was looking for the
-generation of that particular proxy model.  There are some issues to consider
-with respect to panning and zooming that we will not discuss here due to the
-limited space available.
+  \label{eq:coverage}
+\end{equation}
+where $n$ is the number of proxy models available, $\textbf{camdir}$ is the
+direction in which the camera was looking when a particular model was generated,
+and the model to be replaced is the one that minimizes~(\ref{eq:coverage}) after
+replacement.
+
+}
 
 %-------------------------------------------------------------------------
-\subsection{Dynamic/heuristic method selection, depth compression, auto-tuning}
+\subsection{Auto-tuning}
 \label{sec:autoTuning}
 
-Since the main purpose of the automatically generated proxy model is to
-facilitate a client-rendered image while waiting for one from the server for
-increased interactivity, it is important that the process of generating and
-sending proxy model data from the server itself is not hold up more than
-necessary. There are mainly three sources of delay for server-rendered images to
-the client; high latency, low bandwidth, and slow server-rendering itself.
+{\color{black}
 
-In the first case, it seems prudent to have a better proxy model on the client,
-that can be used for longer time and for interactivity resulting in viewing
-parameters deviating more from the last received server-rendered image. In the
-two other cases, it is important for the proxy model generation/transmission to
-be quicker, both in order to get the proxy model to the client and keep from
-delaying the server-image more than necessary.
+It is important that the process of generating and sending proxy model data from
+the server itself does not slow down transmission.  There are mainly three
+sources of delay for server-rendered images to the client; high latency, low
+bandwidth, and slow server-rendering.  In the first case, it seems prudent to
+have a better proxy model on the client, that can be used for longer time and
+for interactivity resulting in viewing parameters deviating more from the last
+received server-rendered image. In the two other cases, it is important for the
+proxy model generation/transmission to be quicker, both in order to get the
+proxy model to the client and keep from delaying the server-image more than
+necessary.
 
-To take care of this, we have adopted an adaptive specification of
-proxy model data from the server. This involves a more light-weight image (lossy
-JPG-compression with adaptive quality control) while the user is interacting
-with the client, as well as a reduced resolution of the depth buffer sent from
-the server. Another possible form of auto-tuning is to let the client adaptively
-choose the number of splats, stored number of proxy models, etc.
+We have adopted an adaptive specification of proxy model data from the server
+involving a more light-weight image (lossy JPG-compression with adaptive quality
+control) while interaction is ongoing.  Also, reduced resolution of the depth
+buffer sent from the server is used. Another possibility is to let the client
+dynamically set the number of splats, number of proxy models, etc.
 
-%See Figure~\ref{fig:adapativeProxyModels} for examples of low- and
-%high-quality proxy models.
-%\begin{figure}[htb]
-%  \centering
-%  \subfigure[Full png image,\newline full depth buffer]{
-%    \includegraphics[width=.30\linewidth]{auto1.jpg}
-%  }
-%  \subfigure[Coarse jpg,\newline quality $q=0$]{
-%    \includegraphics[width=.30\linewidth]{auto2.jpg}
-%  }
-%  \subfigure[Auto-proxy, trun-\newline cated depth data]{
-%    \includegraphics[width=.30\linewidth]{auto3.jpg}
-%  }
-%  \caption{\label{fig:adapativeProxyModels} Proxy models adaptively adjusted to
-%           available bandwidth and server capacity. The server-rendered JPG
-%           image adaptively selects a quality setting based on recorded
-%           timings. Shown is the worst case, \texttt{quality}=0, (b) which is
-%           the last step before switching to ``auto proxy'' (c). The auto-proxy
-%           model shown displays artifacts caused both by the magenta surface
-%           being almost parallel to the viewing direction at the moment of proxy
-%           model recording, and by the non-smooth parts of the geometry
-%           (corners). The number of splats is set low on purpose, and the depth
-%           buffer size is reduced by a factor of $16^{-2} \approx 0.4\%$. \textbf{KRYMPE HER}}
-%\end{figure}
+}
 
@@
 
+{\color{black}
+
 We have tested the algorithms discussed in the Tinia framework, which is a
-programming framework for setting up and managing a client/server based
-interactive OpenGL-based visualization application, see~\cite{tinia}. As client
-we use Google Chrome, the code is written in standard Javascript/WebGL.
-The client communicatesover tcp/ip, using either http or
-websockets, with a server running as an Apache module.
+programming framework for setting up and managing client/server based
+interactive visualization applications, see~\cite{tinia}. As client we use
+Google Chrome, code is written in standard Javascript/WebGL.  The implementation
+in Tinia is invisible to the application, so all existing Tinia-applications
+will have the feature available. The algorithm is minimally intrusive in that
+only the depth buffer will have to be added to the rendering output of the
+application. We have tested several smaller test cases, but also on a larger oil
+reservoir viewer, FRView~\cite{cloudviz}.
 
-The automatic proxy geometry implementation in Tinia is fully invisible to the
-application, meaning that all existing applications immediately will have the
-feature available. Without Tinia, the algorithm is minimally intrusive in that
-only the depth buffer will have to be added to the rendering output of the
-application. We have tested the algorithm on several smaller test cases, but
-also on a larger oil reservoir viewer, FRView, developed in the CloudViz
-project, see~\cite{cloudviz}.
-
-This viewer utilizes several GLSL shaders to render reservoir cells and
-boundaries ({\em corner point grids}), tubular wells etc., and operates in
-general on a geometric model that is not trivial to reduce in complexity. The
-model is typically also very large, meaning that rendering the full model on a
-thin client is prohibitive. With the automatic proxy geometry, we can obtain
-interactive frame rates with a limited connection from a very lightweight client
-to the server. For a comparison of a server-rendered image and a client-rendered
-proxy model that is slightly rotated on the client, see again
-Figure~\ref{fig:FRView}. {\color{red}In Figure~\ref{fig:FRView}~(a), the full
-server-rendered image is shown, and in~(b) and~(c), a slightly (about 10
-degrees) rotated proxy model is rendered on the client. In~(b), parameters are
-chosen for best possible results, while in~(c), we want to highlight some
-effects. For the latter, we have chosen a small number of non-overlapping
-splats, which are also outlined. In this image, it is easy to spot areas not
-well covered by any splats, and also some areas where the texture coordinate
-transform $T$ does not produce optimal results.}
+FRView utilizes several GLSL shaders to render reservoir cells, boundaries,
+tubular wells etc., and visualizes a 3D model that is not trivial to reduce in
+complexity. It is typically also very large, so rendering it on a thin client is
+prohibitive. With the automatic proxy model, we obtain interactive frame rates
+with limited connection from a lightweight client. For a comparison of a
+server-rendered image and a client-rendered proxy model that is slightly rotated
+on the client, see again Figure~\ref{fig:FRView}. In
+Figure~\ref{fig:FRView}~(a), the full server-rendered image is shown, and in~(b)
+and~(c), a slightly (about 10 degrees) rotated proxy model is rendered on the
+client. In~(b), parameters are chosen for best possible results, while in~(c),
+we want to highlight effects, so it uses a small number of non-overlapping large
+splats. One can easily spot areas not well covered, and areas where the texture
+coordinate transform $T$ does not produce optimal results.
 
@@
 
-The single most attractive feature of such an automatic proxy geometry as
-described, is exactly the automated generation. It means that problems with
-downscaling, sending and rendering existing geometries is bypassed
-altogether. The Tinia framework with which we have demonstrated this, is a
-framework for setting up and managing a client/server based interactive OpenGL
-based visualization application. The automatic proxy geometry implementation is
-fully invisible to the application, meaning that all existing applications
-immediately will have the feature available.
-
-There are several directions in which we would like to follow up and improve
-this concept, foremost of these are,
+The most attractive feature of the method described is the automated generation
+of the proxy model. Problems with compression and simplification of existing
+geometries is bypassed altogether. The automatic proxy model implementation is
+invisible to the application. There are several directions in which we would
+like to follow up and improve this concept, a couple of these are,
 \begin{itemize}
 
-\item \textbf{Proxy model replacement algorithms} The goal is to obtain a good view with
-all relevant viewing parameters on the client, with the help of a minimal set of
+\item \textbf{Proxy model replacement algorithms} Obtaining good results with a minimal set of
 proxy models.
 
-%\item \textbf{Grid-based proxy geometry} Instead of a set of ``splats'', one can
-%envision a spatially partioned grid, for instance containing a distance field,
-%with the geometry as an iso-surface.
+\item \textbf{Depth compression} Other approaches than truncation are possible,
+see\eg,~\cite{DBLP:journals/tvcg/Lindstrom14}.
 
-%\item \textbf{Client rendering with overlays} Items like, \eg, axis indicators,
-%legends etc. that do not use the same view matrix as the main model could be
-%rendered locally on the client, only. This will cancel some of the benefit
-%connected to the automatic proxy model generation, though. \textbf{Nettopp av
-%denne grunn kan kanskje dette vaere et punkt vi boer luke bort? Plassmangel
-%ogsaa...}
-
-\item \textbf{Depth compression} We have currently done a unsophisticated
-truncation, but other approaches are possible,
-like discussed in~\cite{DBLP:journals/tvcg/Lindstrom14}.
-
-%\item \textbf{Improved auto-tuning} The auto-tuning discussed in
-%Section~\ref{sec:autoTuning} should be extended to take not only server
-%parameters (rendering quality, compression etc.) into account, but also the
-%capabilities of various thin clients.
-
-\item \textbf{Deferred shading} In addition to sending an image and a depth map
-from the server, one could send also a normal map. (This can also be
-reconstructed from the depth map on the client.) This would allow more advanced
+\item \textbf{Deferred shading} Sending a normal map (can also be
+reconstructed from the depth map) would allow more advanced
 shading on the client.
 
-%\item \textbf{Stitching of 2.5D point clouds} As mentioned in
-%Section~\ref{sec:prevWork}, we have considered but not experimented with
-%stitching of depth maps from the server.
+\end{itemize}
 
-\end{itemize}
+}
 

### 2014-12-09 tex/short_2015/body.tex
final version

@@
 % instead of the author's name (and leave the affiliation blank) !!
-\author[J.\,O. Nygaard \& J. Mikkelsen]
-       {J.\,O. Nygaard$^{1}$
-        and J. Mikkelsen$^{1}$
-        and Husk Anonymisering Hvis Paakrevet! \textbf{JA DET ER DET}$^{1}$
-%        S. Spencer$^2$\thanks{Chairman Siggraph Publications Board}
-        \\
-% For Computer Graphics Forum: Please use the abbreviation of your first name.
-%         $^1$SINTEF ICT, applied mathematics, Norway
-%        $^2$ Another Department to illustrate the use in papers from authors
-%             with different affiliations
-       }
+\author[short1043]{short1043}
+%\author[J.\,O. Nygaard \& J. Mikkelsen]
+%       {J.\,O. Nygaard$^{1}$
+%        and J. Mikkelsen$^{1}$
+%        \\
+%         $^1$SINTEF ICT, Applied Mathematics, Norway
+%       }
 
@@
   \subfigure[Server-rendered (OpenGL)]{
-    \includegraphics[width=0.3\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
+    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
   }
   \subfigure[Client-rendered (WebGL)]{
-    \includegraphics[width=0.3\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
+    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
   }
   \subfigure[Client-rendered (WebGL)]{
-    \includegraphics[width=0.3\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
+    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
     % The trim params: left lower right upper
   }
-  \caption{\label{fig:FRView} The oil reservoir viewer FRView~\cite{cloudviz},
-  showing one server-rendered and two client-rendered images of automatically
-  generated and slightly rotated proxy models.  For~(b), parameters are chosen to produce the best
-  possible image, and in~(c) we want to highlight artifacts and implementational
-  details. See the text for further discussion of these.}
+  \caption{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
+  (name withheld for the double-blind review process) showing one
+  server-rendered and two client-rendered images of automatically generated and
+  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
+  best possible image, and in~(c) we want to highlight artifacts and
+  implementational details. See the text for further discussion.}
 }
@@
 \begin{abstract}
-\color{black}
    We describe an algorithm and implementation of automatic proxy model
    generation for a client/server remote rendering setup. A {\em proxy model} is
-   a light-weight version of a main 3D model that is cheaper to render. We
+   a light-weight version of a main 3D model that is cheaper to render and transfer. We
    automatically derive this from the main model. The client is typically a
@@
    itself.
+  % contribution sentence:
+  %
+  % The contribution of the work is the novel application of the depth buffer by-product of a rendering process to generate a simplified model of the rendered scene.
   % Husk: Skal vaere en blank linje nedenfor
@@
 
-{\color{black} Hand in hand with increasingly powerful rendering engines comes
-ever increasing requirements on computational accuracy, power efficiency, data
-sizes, scaling, etc. Also enhanced by new cloud-based approaches etc., an effect
-is that interactivity still is a difficult issue. We consider a client/server
-model for 3D rendering, addressing latency, bandwidth and scaling problems in a
-novel way.
+Hand in hand with increasingly powerful rendering engines comes ever increasing
+requirements on computational accuracy, power efficiency, data sizes, scaling
+properties, etc. This is also reinforced by new cloud-based approaches and
+wireless usage patterns. An effect of this is that interactivity still is a
+difficult issue. We consider a client/server model for 3D rendering, addressing
+latency, bandwidth and scaling problems in a novel way.
 
 We introduce a {\em proxy model}, defined as a temporary model to be shown and
-manipulated locally on the client while waiting for the appropriate image from
-the server. Producing such proxy models can be difficult for many kinds of 3D
-data, like in our main case in which we have an oil reservoir viewer
-(FRView,~\cite{cloudviz}) that renders large {\em corner point grids}, together
+manipulated locally on a client while waiting for the appropriate image from a
+connected server. Producing such proxy models can be difficult for many kinds of
+3D data, like in our main case in which we have an oil reservoir viewer
+(YYYY,~\cite{cloudviz}) that renders large {\em corner point grids}, together
 with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
-example of both a fully completed rendering, and automatically generated proxy
-geometries rendered in Google Chrome.
+example of both a full server-side rendering, and automatically generated proxy
+models rendered in Google Chrome.
 %
@@
 the position or orientation of the model too much, this proxy model rendering
-integrates seemlessly with the slower stream of server-rendered frames.
+integrates seamlessly with the slower stream of server-rendered frames.
 %
@@
 effective server time available for each one. Suitable scaling may still be
-achieved using our solution. This will be commercialized in the new SINTEF
-project \mbox{ArrowHead}~\cite{arrowhead}.}
+achieved using our solution. This is currently being commercialized in an
+ongoing XXXX (name withheld for the double-blind review process) project.
 
+%-------------------------------------------------------------------------
 \subsection{Previous work}
@@
 we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
-depth map compression, an issue that we also consider. {\color{black}Another work
+depth map compression, an issue that we also consider. Another work
 focused on similar streaming and geometry compression,
 is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
-work, with respect to image-based rendering accelleration.}
+work, with respect to image-based rendering acceleration.
 %
 Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
-point clouds. We could also in principle do this for our 2.5D maps fetched from
-the server, but it is not clear that the benefit would outweight the
-cost.
+point clouds. We could do this for our 2.5D maps fetched from the server, but it
+is unclear if the benefit would outweigh the cost.
 %
@@
 %
-{\color{black}Finally, we note that these 2.5D heightmaps are exactly what ``3D cameras''
+We observe that these 2.5D height maps are exactly what ``3D cameras''
 (time-of-flight and other range image sensors) produce, but most authors
 considering these are typically building more complex geometries before
-visualization.} % , see~\eg,~\cite{}
-\marginpar{\textbf{REF?!}}
+visualization, see~\cite{IMM2009-05801}.
 
 %-------------------------------------------------------------------------
-
 \section{The auto proxy algorithm}
 
-{\color{black} Since the depth buffer is a height map seen from the observer, it
+Since the depth buffer is a height map seen from the observer, it
 does not contain information about occluded scene elements. Our approach assumes
@@
 intermediate client-rendered proxy models with different features that will be
-discussed in Section~\ref{sec:client}.}
+discussed in Section~\ref{sec:client}.
 
@@
   \caption{\label{fig:2DheightmapRotated}
-{\color{black} 
 Bird's view of 3D model (solid lines) and proxy model renderings.
-}
 %           The depth map and image sent from the server enables the client to
@@
 
-{\color{black}
 The server renders the 3D model into a framebuffer, and in the process generates
@@
 each depth value in the range $[0, 1]$ as a 16 bit fixed point number, and the
-bundling of the depth image with the rgb image then imposes just a small data
+bundling of the depth image with the RGB image then imposes just a small data
 overhead.
@@
 compression/decompression must also be
-considered. In~\cite{DBLP:journals/tvcg/Lindstrom14} a suitable algorithm is
-proposed that promises to be fast both for the compression and decompression
-stages.
-}
+considered.
+One proposed solution is to be found in~\cite{DBLP:journals/tvcg/Lindstrom14},
+which promises to be fast both for the compression and decompression stages.
 
@@
 
-{\color{black}
-
-When the client receieves an rgb and depth image, together with view
-transformations, it builds a ``proxy model'' from this. This model can then be
+When the client receives an RGB and depth image, together with view
+transformations, it builds a proxy model from this. This model can then be
 transformed and rendered directly, or it can be combined with other proxy models
@@
 which iso-surfaces can be extracted, etc.
-}
 
@@
 
-{\color{black}
 Each splat produces many client pixels, necessitating an ``intra-splat''
@@
 samples, and the calculation of $\sv_x$ and $\sv_y$ is really a gradient
-approximation, so we should not make them to large either.
+approximation, so we should not make them too large either.
 
@@
            Splats for which depths are sampled on a planar region look nice on
-  exactly this region, but contorted elsewhere, notice the corner and edges of
-  the cube.}
+  this region, but possibly contorted elsewhere. Notice the corner and edges
+  here, where we have chosen parameters to display this effect.}
 \end{figure}
-%\vspace{-4ex}
 
@@
 exactly a better ``global'' texture transform implementation.
-}
 
@@
 
-{\color{black}
-
 \textbf{Splat sizing} 
 Each splat should be rendered into a number of client pixels according to the
-new splats 3D position. To achieve this, we use the vectors $\sv_{\Delta x}$ and
-$\sv_{\Delta y}$ from the previous section, and in addition, we scale the splats
-up a bit so that they overlap. Hence, we can render rectangular window-aligned
-splats with less risk of getting uncovered areas when interactively rotating and
-scaling the model on the client.
+new splats' 3D position. To achieve this, we use the vectors $\sv_{\Delta x}$
+and $\sv_{\Delta y}$ from the previous section, and in addition, we scale the
+splats up a bit so that they overlap. Hence, we can render rectangular
+window-aligned splats with less risk of getting uncovered areas when
+interactively rotating and scaling the model on the client. In most cases this
+removes the problem visualized in Figure~\ref{fig:LargeSplatsOnCorners}.
 
@@
 image, just as the texture is looked up for color. Not all clients support this
-at the time of writing, it is an extension in WebGL currently.
+WebGL extension.
 
@@
 
-}
-
 %-------------------------------------------------------------------------
@@
 \label{sec:autoTuning}
-
-{\color{black}
 
@@
 bandwidth, and slow server-rendering.  In the first case, it seems prudent to
-have a better proxy model on the client, that can be used for longer time and
-for interactivity resulting in viewing parameters deviating more from the last
-received server-rendered image. In the two other cases, it is important for the
-proxy model generation/transmission to be quicker, both in order to get the
-proxy model to the client and keep from delaying the server-image more than
-necessary.
+have a good proxy model on the client, which can be used for longer time and for
+a wider range of client-side transformations.  In the two other cases, it is
+important for the proxy model generation/transmission to be cheap/fast, both in order
+to get the proxy model to the client and keep from delaying the server-image
+more than necessary. These demands are not always compatible.
 
@@
 
-{\color{black}
+We have tested the algorithms discussed in the ZZZZ (name withheld for the
+double-blind review process) framework, which is a programming framework for
+setting up and managing client/server based interactive visualization
+applications, see~\cite{tinia}. As client we use Google Chrome, code is written
+in standard Javascript/WebGL.  The implementation in ZZZZ is invisible to the
+application, so all existing ZZZZ-applications will have the feature
+available. The algorithm is minimally intrusive in that only the depth buffer
+will have to be added to the rendering output of the application. We have tested
+several smaller test cases, but also on a larger oil reservoir viewer,
+YYYY~\cite{cloudviz}.
 
-We have tested the algorithms discussed in the Tinia framework, which is a
-programming framework for setting up and managing client/server based
-interactive visualization applications, see~\cite{tinia}. As client we use
-Google Chrome, code is written in standard Javascript/WebGL.  The implementation
-in Tinia is invisible to the application, so all existing Tinia-applications
-will have the feature available. The algorithm is minimally intrusive in that
-only the depth buffer will have to be added to the rendering output of the
-application. We have tested several smaller test cases, but also on a larger oil
-reservoir viewer, FRView~\cite{cloudviz}.
-
-FRView utilizes several GLSL shaders to render reservoir cells, boundaries,
+YYYY utilizes several GLSL shaders to render reservoir cells, boundaries,
 tubular wells etc., and visualizes a 3D model that is not trivial to reduce in
@@
 
-\item \textbf{Depth compression} Other approaches than truncation are possible,
-see\eg,~\cite{DBLP:journals/tvcg/Lindstrom14}.
+\item \textbf{Depth compression} We would like to ivestigate other approaches
+than just truncation, see \eg,~\cite{DBLP:journals/tvcg/Lindstrom14}.
 
-\item \textbf{Deferred shading} Sending a normal map (can also be
-reconstructed from the depth map) would allow more advanced
-shading on the client.
+\item \textbf{Deferred shading} With a normal map more advanced
+shading could be done on the client. Such a map could also be constructed from
+the depth map.
 
 \end{itemize}
-
-}
 

### 2014-12-10 src/qtcontroller/ServerThread.cpp
removed/commented out testing bits and pieces.

@@ -453,8 +453,7 @@ void ServerThread::errorCode(QTextStream &os, unsigned int code, const QString &
 QString ServerThread::getStaticContent(const QString &uri)
 {
 
-//    QString fullPath = ":javascript/" + uri;
-    QString fullPath = "/home/jnygaard/new_system/prosjekter/tinia_checkout_141127/tinia/js/" + uri;
+    QString fullPath = ":javascript/" + uri;
 
     QFile file(fullPath);
     if(file.open(QIODevice::ReadOnly)) {

### 2017-09-21 tex/full_2017/bibtemplate.tex
First commit of new paper-directory.

\documentclass[10pt]{IEEEtran}

\usepackage{times}
\usepackage[latin1]{inputenc}
\usepackage[hyphens]{url}
\usepackage[bookmarks={false},colorlinks={false},hyperref={false}]{hyperref}
\usepackage{nohyperref}

\begin{document}
\section{Introduction}
Examples for citations and references \cite{Firstman:2014:Approach} and \cite{Firstman:2013:Ongoing}
and \cite{Webpage:2014:Somewebpage}
and \cite{Bookman:2014:Somebook} and \cite{Chapman:2014:Starting}.
In a consecutive use these are referenced as
\cite{Firstman:2014:Approach,Firstman:2013:Ongoing,Webpage:2014:Somewebpage,Bookman:2014:Somebook,Chapman:2014:Starting}.

\bibliographystyle{IEEEtran}
\bibliography{bibtemplate_samples}

\end{document}

### 2017-09-21 tex/full_2017/body.tex
First commit of new paper-directory.

--- 2014-12-09 tex/short_2015/body.tex
@@
   %
-  % The contribution of the work is the novel application of the depth buffer by-product of a rendering process to generate a simplified model of the rendered scene.
+  % The contribution of the work is the novel application of the depth buffer
+  % by-product of a rendering process to generate a simplified model of the
+  % rendered scene.
   % Husk: Skal vaere en blank linje nedenfor
@@
 %
+\textbf{A very early work aiming at the same kind of ``inter-frame rendering'', is
+\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped, and
+subsequently combined, in order to avoid large unpainted areas caused by
+occlusion. This corresponds to our use of several proxy models, each from a pair
+of (rgb, depth)-images, the main difference being that they work on separate
+pixels, while we use larger, textured ``splats''. Our method avoids their
+slightly complicated meshing, discontinuity detection and final image
+composition stage.}
+%
 Common to many IBR-algorithms is also that of stitching together 3D or 2.5D
@@
 visualization, see~\cite{IMM2009-05801}.
+%
+\textbf{Another approach is that of~\cite{altenhofen16rixels}. They use websockets
+where we use the http protocol, a more significant difference is that we get
+away with transferring a lot less data due to our adaptive compression ratio
+depending on continuous bandwidth and latency measurements.
+%
+}
 

### 2017-09-21 tex/full_2017/paper2.tex
First commit of new paper-directory.

--- 2017-09-21 tex/full_2017/body.tex
@@
-% ---------------------------------------------------------------------
-% EG author guidelines plus sample file for EG publication using LaTeX2e input
-% D.Fellner, v1.17, Sep 23, 2010
+%% bare_conf.tex
+%% V1.3
+%% 2007/01/11
+%% by Michael Shell
+%% See:
+%% http://www.michaelshell.org/
+%% for current contact information.
+%%
+%% This is a skeleton file demonstrating the use of IEEEtran.cls
+%% (requires IEEEtran.cls version 1.7 or later) with an IEEE conference paper.
+%%
+%% Support sites:
+%% http://www.michaelshell.org/tex/ieeetran/
+%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
+%% and
+%% http://www.ieee.org/
 
-%\title[Automatic Proxy Geometry Generation]%
-%      {Automatic Proxy Geometry Generation}
+%%*************************************************************************
+%% Legal Notice:
+%% This code is offered as-is without any warranty either expressed or
+%% implied; without even the implied warranty of MERCHANTABILITY or
+%% FITNESS FOR A PARTICULAR PURPOSE! 
+%% User assumes all risk.
+%% In no event shall IEEE or any contributor to this code be liable for
+%% any damages or losses, including, but not limited to, incidental,
+%% consequential, or any other damages, resulting from the use or misuse
+%% of any information contained here.
+%%
+%% All comments are the opinions of their respective authors and are not
+%% necessarily endorsed by the IEEE.
+%%
+%% This work is distributed under the LaTeX Project Public License (LPPL)
+%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
+%% distributed and modified. A copy of the LPPL, version 1.3, is included
+%% in the base LaTeX documentation of all distributions of LaTeX released
+%% 2003/12/01 or later.
+%% Retain all contribution notices and credits.
+%% ** Modified files should be clearly indicated as such, including  **
+%% ** renaming them and changing author support contact information. **
+%%
+%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
+%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
+%%*************************************************************************
 
-\title[Automatic Proxy Model Generation]%
-      {Hybrid Client/Server Rendering with Automatic Proxy Model Generation}
+% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
+% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
+% *** with production work. IEEE's font choices can trigger bugs that do  ***
+% *** not appear when using other class files.                            ***
+% The testflow support page is at:
+% http://www.michaelshell.org/tex/testflow/
 
-% for anonymous conference submission please enter your SUBMISSION ID
-% instead of the author's name (and leave the affiliation blank) !!
-\author[short1043]{short1043}
+% Note that the a4paper option is mainly intended so that authors in
+% countries using A4 can easily print to A4 and see how their papers will
+% look in print - the typesetting of the document will not typically be
+% affected with changes in paper size (but the bottom and side margins will).
+% Use the testflow package mentioned above to verify correct handling of
+% both paper sizes by the user's LaTeX system.
+%
+% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
+% should be used if it is desired that the figures are to be displayed in
+% draft mode.
+%
+\documentclass[10pt,conference,compsocconf]{IEEEtran}
+%\documentclass[10pt]{IEEEtran}
+\usepackage{times}
+
+\usepackage{caption}
+\captionsetup{font=footnotesize,justification=centering,labelsep=period}
+
+% Add the compsoc option for Computer Society conferences.
+%
+% If IEEEtran.cls has not been installed into the LaTeX system files,
+% manually specify the path to it like:
+% \documentclass[conference]{../sty/IEEEtran}
+
+% Some very useful LaTeX packages include:
+% (uncomment the ones you want to load)
+
+% *** MISC UTILITY PACKAGES ***
+%
+%\usepackage{ifpdf}
+% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
+% compilation based on whether the output is pdf or dvi.
+% usage:
+% \ifpdf
+%   % pdf code
+% \else
+%   % dvi code
+% \fi
+% The latest version of ifpdf.sty can be obtained from:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
+% Also, note that IEEEtran.cls V1.7 and later provides a builtin
+% \ifCLASSINFOpdf conditional that works the same way.
+% When switching from latex to pdflatex and vice-versa, the compiler may
+% have to be run twice to clear warning/error messages.
+
+% *** CITATION PACKAGES ***
+%
+%\usepackage{cite}
+% cite.sty was written by Donald Arseneau
+% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
+% \cite{} output to follow that of IEEE. Loading the cite package will
+% result in citation numbers being automatically sorted and properly
+% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
+% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
+% \cite will automatically add leading space, if needed. Use cite.sty's
+% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
+% cite.sty is already installed on most LaTeX systems. Be sure and use
+% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
+% not currently provide for hyperlinked citations.
+% The latest version can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
+% The documentation is contained in the cite.sty file itself.
+
+% *** GRAPHICS RELATED PACKAGES ***
+%
+\ifCLASSINFOpdf
+  % \usepackage[pdftex]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../pdf/}{../jpeg/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
+\else
+  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
+  % will default to the driver specified in the system graphics.cfg if no
+  % driver is specified.
+  % \usepackage[dvips]{graphicx}
+  % declare the path(s) where your graphic files are
+  % \graphicspath{{../eps/}}
+  % and their extensions so you won't have to specify these with
+  % every instance of \includegraphics
+  % \DeclareGraphicsExtensions{.eps}
+\fi
+% graphicx was written by David Carlisle and Sebastian Rahtz. It is
+% required if you want graphics, photos, etc. graphicx.sty is already
+% installed on most LaTeX systems. The latest version and documentation can
+% be obtained at: 
+% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
+% Another good source of documentation is "Using Imported Graphics in
+% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
+% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
+%
+% latex, and pdflatex in dvi mode, support graphics in encapsulated
+% postscript (.eps) format. pdflatex in pdf mode supports graphics
+% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
+% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
+% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
+% which can result in "jaggedy"/blurry rendering of lines and letters as
+% well as large increases in file sizes.
+%
+% You can find documentation about the pdfTeX application at:
+% http://www.tug.org/applications/pdftex
+
+% *** MATH PACKAGES ***
+%
+\usepackage[cmex10]{amsmath}
+% A popular package from the American Mathematical Society that provides
+% many useful and powerful commands for dealing with mathematics. If using
+% it, be sure to load this package with the cmex10 option to ensure that
+% only type 1 fonts will utilized at all point sizes. Without this option,
+% it is possible that some math symbols, particularly those within
+% footnotes, will be rendered in bitmap form which will result in a
+% document that can not be IEEE Xplore compliant!
+%
+% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
+% thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500
+% after loading amsmath to restore such page breaks as IEEEtran.cls normally
+% does. amsmath.sty is already installed on most LaTeX systems. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
+
+% *** SPECIALIZED LIST PACKAGES ***
+%
+%\usepackage{algorithmic}
+% algorithmic.sty was written by Peter Williams and Rogerio Brito.
+% This package provides an algorithmic environment fo describing algorithms.
+% You can use the algorithmic environment in-text or within a figure
+% environment to provide for a floating algorithm. Do NOT use the algorithm
+% floating environment provided by algorithm.sty (by the same authors) or
+% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
+% algorithm float types and packages that provide these will not provide
+% correct IEEE style captions. The latest version and documentation of
+% algorithmic.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
+% There is also a support site at:
+% http://algorithms.berlios.de/index.html
+% Also of interest may be the (relatively newer and more customizable)
+% algorithmicx.sty package by Szasz Janos:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
+
+% *** ALIGNMENT PACKAGES ***
+%
+%\usepackage{array}
+% Frank Mittelbach's and David Carlisle's array.sty patches and improves
+% the standard LaTeX2e array and tabular environments to provide better
+% appearance and additional user controls. As the default LaTeX2e table
+% generation code is lacking to the point of almost being broken with
+% respect to the quality of the end results, all users are strongly
+% advised to use an enhanced (at the very least that provided by array.sty)
+% set of table tools. array.sty is already installed on most systems. The
+% latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/required/tools/
+
+%\usepackage{mdwmath}
+%\usepackage{mdwtab}
+% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
+% especially mdwmath.sty and mdwtab.sty which are used to format equations
+% and tables, respectively. The MDWtools set is already installed on most
+% LaTeX systems. The lastest version and documentation is available at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
+
+% IEEEtran contains the IEEEeqnarray family of commands that can be used to
+% generate multiline equations as well as matrices, tables, etc., of high
+% quality.
+
+%\usepackage{eqparbox}
+% Also of notable interest is Scott Pakin's eqparbox package for creating
+% (automatically sized) equal width boxes - aka "natural width parboxes".
+% Available at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
+
+% *** SUBFIGURE PACKAGES ***
+%\usepackage[tight,footnotesize]{subfigure}
+% subfigure.sty was written by Steven Douglas Cochran. This package makes it
+% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
+% work, it is a good idea to load it with the tight package option to reduce
+% the amount of white space around the subfigures. subfigure.sty is already
+% installed on most LaTeX systems. The latest version and documentation can
+% be obtained at:
+% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
+% subfigure.sty has been superceeded by subfig.sty.
+
+%\usepackage[caption=false]{caption}
+%\usepackage[font=footnotesize]{subfig}
+% subfig.sty, also written by Steven Douglas Cochran, is the modern
+% replacement for subfigure.sty. However, subfig.sty requires and
+% automatically loads Axel Sommerfeldt's caption.sty which will override
+% IEEEtran.cls handling of captions and this will result in nonIEEE style
+% figure/table captions. To prevent this problem, be sure and preload
+% caption.sty with its "caption=false" package option. This is will preserve
+% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
+% (recommended due to many improvements over 1.2) of subfig.sty supports
+% the caption=false option directly:
+%\usepackage[caption=false,font=footnotesize]{subfig}
+%
+% The latest version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
+% The latest version and documentation of caption.sty can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
+
+% *** FLOAT PACKAGES ***
+%
+%\usepackage{fixltx2e}
+% fixltx2e, the successor to the earlier fix2col.sty, was written by
+% Frank Mittelbach and David Carlisle. This package corrects a few problems
+% in the LaTeX2e kernel, the most notable of which is that in current
+% LaTeX2e releases, the ordering of single and double column floats is not
+% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
+% single column figure to be placed prior to an earlier double column
+% figure. The latest version and documentation can be found at:
+% http://www.ctan.org/tex-archive/macros/latex/base/
+
+%\usepackage{stfloats}
+% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
+% the ability to do double column floats at the bottom of the page as well
+% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
+% LaTeX2e). It also provides a command:
+%\fnbelowfloat
+% to enable the placement of footnotes below bottom floats (the standard
+% LaTeX2e kernel puts them above bottom floats). This is an invasive package
+% which rewrites many portions of the LaTeX2e float routines. It may not work
+% with other packages that modify the LaTeX2e float routines. The latest
+% version and documentation can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
+% Documentation is contained in the stfloats.sty comments as well as in the
+% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
+% does not allow \baselineskip to stretch. Authors submitting work to the
+% IEEE should note that IEEE rarely uses double column equations and
+% that authors should try to avoid such use. Do not be tempted to use the
+% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
+% not format its papers in such ways.
+
+% *** PDF, URL AND HYPERLINK PACKAGES ***
+%
+%\usepackage{url}
+% url.sty was written by Donald Arseneau. It provides better support for
+% handling and breaking URLs. url.sty is already installed on most LaTeX
+% systems. The latest version can be obtained at:
+% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
+% Read the url.sty source comments for usage information. Basically,
+% \url{my_url_here}.
+
+% *** Do not adjust lengths that control margins, column widths, etc. ***
+% *** Do not use packages that alter fonts (such as pslatex).         ***
+% There should be no need to do such things with IEEEtran.cls V1.6 and later.
+% (Unless specifically asked to do so by the journal or conference you plan
+% to submit to, of course. )
+
+% ---------- my stuff ----------
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+\usetikzlibrary{shapes}
+\usetikzlibrary{positioning}
+\usetikzlibrary{backgrounds}
+\usetikzlibrary{calc}
+\usepackage{3dplot}
+\usetikzlibrary{trees}
+\usetikzlibrary{chains,fit,shapes,intersections}
+
+\usepackage[plain]{algorithm}
+%\usepackage[noend]{algorithmic}
+\usepackage{algorithmic}
+
+%\usepackage[]{algorithm2e}
+%%\usepackage[hang]{subfigure}
+\usepackage[nooneline]{subfigure}
+\usepackage{subfigure}
+
+%\usepackage{amsmath}
+
+\newcommand{\ga}{{\gamma}}
+
+%\newcommand{\xv}{\mbox{\boldmath $x$}} % funker ikke lenger...!?
+
+\newcommand{\bv}{\mathbf{b}}
+\newcommand{\cv}{\mathbf{c}}
+\newcommand{\dv}{\mathbf{d}}
+\newcommand{\ev}{\mathbf{e}}
+\newcommand{\mv}{\mathbf{m}}
+\newcommand{\pv}{\mathbf{p}}
+\newcommand{\sv}{\mathbf{s}}
+\newcommand{\uv}{\mathbf{u}}
+\newcommand{\vv}{\mathbf{v}}
+\newcommand{\xv}{\mathbf{x}}		% Works, but seems more Roman than italics!
+
+\newcommand{\ub}{\bar{u}}
+
+\newcommand{\evh}{\hat{\mathbf{e}}}
+
+\newcommand{\Ev}{\mathbf{E}}
+\newcommand{\Fv}{\mathbf{F}}
+\newcommand{\Tv}{\mathbf{T}}
+\newcommand{\Nv}{\mathbf{N}}
+\newcommand{\Bv}{\mathbf{B}}
+
+\newcommand{\Tvh}{\hat{\mathbf{T}}}
+\newcommand{\Nvh}{\hat{\mathbf{N}}}
+\newcommand{\Bvh}{\hat{\mathbf{B}}}
+
+\newcommand{\Tvb}{\bar{\mathbf{T}}}
+\newcommand{\Nvb}{\bar{\mathbf{N}}}
+\newcommand{\Bvb}{\bar{\mathbf{B}}}
+
+%\newcommand{\Tvd}{\dot{\mathbf{T}}}
+\newcommand{\Tvd}{\mathbf{T}'}
+
+\newcommand{\zev}{\mathbf{0}}
+
+\newcommand{\ra}{\rightarrow}
+\newcommand{\la}{\leftarrow}
+
+%\newcommand{\gav}{\mathbf{\gamma}} % funker heller ikke!
+\newcommand{\gav}{\boldsymbol{\gamma}} % Works only when bm.sty is used. If it is not, we get an ordinary symbol, but no warning/error!
+\newcommand{\nuv}{\boldsymbol{\nu}}
+\newcommand{\nuvh}{\boldsymbol{\nu}}
+
+\newcommand{\xvv}{\boldsymbol{x}}	% Strange. This looks like a ``poor mans bold'', i.e., a couple of superimposed
+	                                % and slightly offset x'es...
+
+\newcommand{\ie}{{\em i.e.}}
+\newcommand{\eg}{{\em e.g.}}
+\newcommand{\etal}{{\em et~al.}}
+
+% ------------------------------
+
+% correct bad hyphenation here
+\hyphenation{op-tical net-works semi-conduc-tor}
+
+%\parskip 6pt plus 2pt minus 1pt
+\parskip 3pt plus 2pt minus 1pt
+
+\pagestyle{empty}
+\begin{document}
+\pagenumbering{gobble}
+%
+% paper title
+% can use linebreaks \\ within to get better formatting as desired
+\title{\textbf{\Large Hybrid Client/Server Rendering with Automatic Proxy Model Generation}\\[0.2ex]}
+
+% author names and affiliations
+% use a multiple column layout for up to three different
+% affiliations
+
 %\author[J.\,O. Nygaard \& J. Mikkelsen]
@@
 
-% ------------------------------------------------------------------------
+\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
+\IEEEauthorblockA{SINTEF Digital\\
+Applied Mathematics\\
+Norway\\
+Email: {\tt Jens.O.Nygaard@sintef.no}}
+\and
+\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen\\[0.3ex]\normalsize}
+\IEEEauthorblockA{SINTEF Digital\\
+Applied Mathematics\\
+Norway\\
+Email: {\tt Jon.Mikkelsen@sintef.no}}}
 
-% if the Editors-in-Chief have given you the data, you may uncomment
-% the following five lines and insert it here
+% conference papers do not typically use \thanks and this command
+% is locked out in conference mode. If really needed, such as for
+% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
+% after \documentclass
+
+% for over three affiliations, or if they all won't fit within the width
+% of the page, use this alternative format:
 % 
-% \volume{27}   % the volume in which the issue will be published;
-% \issue{1}     % the issue number of the publication
-% \pStartPage{1}      % set starting page
+%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
+%Homer Simpson\IEEEauthorrefmark{2},
+%James Kirk\IEEEauthorrefmark{3}, 
+%Montgomery Scott\IEEEauthorrefmark{3} and
+%Eldon Tyrell\IEEEauthorrefmark{4}}
+%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
+%Georgia Institute of Technology,
+%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
+%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
+%Email: homer@thesimpsons.com}
+%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212, Fax: (888) 555--1212}
+%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
 
-%-------------------------------------------------------------------------
-\begin{document}
+% use for special paper notices
+%\IEEEspecialpapernotice{(Invited Paper)}
 
-\teaser{
+% make the title area
+\maketitle
+
+%\teaser
+{
   \centering
@@
 
-\maketitle
+\begin{abstract}
+%\boldmath
+We describe an algorithm and implementation of automatic proxy model generation
+for a client/server remote rendering setup. A {\em proxy model} is a
+light-weight version of a main 3D model that is cheaper to render and
+transfer. We automatically derive this from the main model. The client is
+typically a tablet, for which we only assume availability of WebGL and
+Javascript. The server makes use of OpenGL and a web server. When the rate of
+received server-rendered images deteriorate, the client renders the proxy model,
+which is computed from depth buffers bundled with rendered images from the
+server. This algorithm requires little modification of the application itself.
+\end{abstract}
+% IEEEtran.cls defaults to using nonbold math in the Abstract.
+% This preserves the distinction between vectors and scalars. However,
+% if the conference you are submitting to favors bold math in the abstract,
+% then you can use LaTeX's standard command \boldmath at the very start
+% of the abstract to achieve this. Many IEEE journals/conferences frown on
+% math in the abstract anyway.
 
-\begin{abstract}
-   We describe an algorithm and implementation of automatic proxy model
-   generation for a client/server remote rendering setup. A {\em proxy model} is
-   a light-weight version of a main 3D model that is cheaper to render and transfer. We
-   automatically derive this from the main model. The client is typically a
-   tablet, for which we only assume availability of WebGL and Javascript. The
-   server makes use of OpenGL and a web server. When the rate of received
-   server-rendered images deteriorate, the client renders the proxy model, which
-   is computed from depth buffers bundled with rendered images from the
-   server. This algorithm requires little modification of the application
-   itself.
-  % contribution sentence:
+% no keywords
+
+\begin{IEEEkeywords}
+Client-server; remote rendering; high latency; low bandwidth
+\end{IEEEkeywords}
+
+% For peer review papers, you can put extra information on the cover
+% page as needed:
+% \ifCLASSOPTIONpeerreview
+% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
+% \fi
 %
-  % The contribution of the work is the novel application of the depth buffer
-  % by-product of a rendering process to generate a simplified model of the
-  % rendered scene.
-  % Husk: Skal vaere en blank linje nedenfor
-
-\begin{classification} % according to http://www.acm.org/class/1998/
-\CCScat{Computer Graphics}{I.3.5}{Computational Geometry and Object Modeling}{Geometric algorithms, languages, and systems}
-% http://www.acm.org/about/class/ccs98-html
-\end{classification}
-
-\end{abstract}
+% For peerreview papers, this IEEEtran command inserts a page break and
+% creates the second title. It will be ignored for other modes.
+\IEEEpeerreviewmaketitle
 
@@
 
-%\bibliographystyle{eg-alpha}
-\bibliographystyle{eg-alpha-doi}
+\bibliographystyle{IEEEtran}
 
+% Sample references
+%\bibliography{bibtemplate_samples}
+
+% Our references
 \bibliography{egbibsample}
@@
 
+% that's all folks
 \end{document}

### 2017-09-25 tex/full_2017/paper2.tex
Tok inn et par paragrafer om de siste paperne Ritschel nevnte, unntatt den som heter Deep Screen Space

@@
 %
-We are not aware of anybody else using depth maps exactly for our purpose, most
+\textbf{[OLD: We are not aware of anybody else using depth maps exactly for our purpose, most
 research seems to be focused on the retrieval of the depth map from images,
-while we use an existing depth map to distill and render temporary geometry.
+while we use an existing depth map to distill and render temporary geometry.]}
+\textbf{[NEW: Our approach differs from many similar ones, in that we use existing depth maps 
+to distill and render temporary geometry, rather than retrieveing the depth maps from images.]}
 %
@@
 }
+%
+\textbf{In~\cite{CGF:CGF1871}, Pajak \etal considers remote rendering and
+  streaming of frames rendered from a dynamical 3D model, we are quite agnostic
+  to the source of imagery. Their setup requires more powerful clients than ours
+  (OpenGL vs. WebGL), but they will also have higher fidelity.  }
+%
+\textbf{A special case is provided for in the rendering of stereo images. In
+  this case, the 3D disparity is limited, while the rendering cost is doubled,
+  since two views per frame is needed. In~\cite{DidykERMS2010}, this is
+  exploited to make a solution tailored to such stereo synthesis; performance
+  approaches that of rendering non-stereo, with a minimization of depth
+  disparity artifacts. Occlusion and disocclusion holes are avoided by warping quads rather than
+  pixel and filling in with previous images.}
+%
+% Er denne egentlig relevant?
+%\textbf{Nalbach \etal,~\cite{Nalbach:2014:DSS:2556700.2556708}, ...
+%}
 

### 2017-10-03 tex/full_2017/paper2.tex
Fixed the teaser image

@@
 % make the title area
-\maketitle
+%\maketitle
 
 %\teaser
-{
+%{
+%  \centering
+%  \subfigure[Server-rendered (OpenGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
+%  }
+%  \subfigure[Client-rendered (WebGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
+%  }
+%  \subfigure[Client-rendered (WebGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
+%    % The trim params: left lower right upper
+%  }
+%  \caption{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
+%  (name withheld for the double-blind review process) showing one
+%  server-rendered and two client-rendered images of automatically generated and
+%  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
+%  best possible image, and in~(c) we want to highlight artifacts and
+%  implementational details. See the text for further discussion.}
+%}
+
+\twocolumn[{%
+\renewcommand\twocolumn[1][]{#1}%
+\maketitle
+\begin{center}
     \centering
-  \subfigure[Server-rendered (OpenGL)]{
-    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
-  }
-  \subfigure[Client-rendered (WebGL)]{
-    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
-  }
-  \subfigure[Client-rendered (WebGL)]{
-    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
-    % The trim params: left lower right upper
-  }
-  \caption{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
+		\begin{tabular}[b]{ccc}
+    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
+    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
+    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
+		a) & b) & c)
+		\end{tabular}
+		\captionof{figure}{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
   (name withheld for the double-blind review process) showing one
@@
   implementational details. See the text for further discussion.}
-}
+%    \protect \label{fig:wfedesign} }
+%    \captionof{figure}{Test caption}
+\end{center}%
+}]
+
+%\begin{figure*}
+%    % The star in \begin{figure*} makes the figure stretch across both columns.
+%    \centering
+%    \includegraphics[width=0.99\linewidth]{frviewb1.png}
+%    \caption{
+%        %The graphical user interface of WFE, where dotted lines represent flow of services, while solid lines connect parameters from outputs to inputs.
+%        The graphical user interface of Workflow Editor displays buttons to append services and to add code snippets for conditional branches, loops, parallel service execution (split), etc., as well as options for service %filtering.
+It also lets users add inputs and outputs to the workflow with appropriate buttons shown as I and O.
+The workflow is shown as a directed graph, where blue boxes represent single CloudFlow services, and dark green% boxes are sub-workflows.
+The execution order is represented by the dotted arrows, while data flow is visualized by solid lines, which connect output parameters from services to be input parameters of others.}
+%    \protect \label{fig:wfedesign}
+%\end{figure*}
+
+%\teaser
+%{
+%  \centering
+%  \subfigure[Server-rendered (OpenGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png}
+%  }
+%  \subfigure[Client-rendered (WebGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png}
+%  }
+%  \subfigure[Client-rendered (WebGL)]{
+%    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png}
+%    % The trim params: left lower right upper
+%  }
+%  \caption{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
+%  (name withheld for the double-blind review process) showing one
+%  server-rendered and two client-rendered images of automatically generated and
+%  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
+%  best possible image, and in~(c) we want to highlight artifacts and
+%  implementational details. See the text for further discussion.}
+%}
 

### 2017-10-04 tex/full_2017/paper2.tex
*fixed author *removed anonemous stuff

@@
 \and
-\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen\\[0.3ex]\normalsize}
+\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
 \IEEEauthorblockA{SINTEF Digital\\
@@
 Norway\\
-Email: {\tt Jon.Mikkelsen@sintef.no}}}
+Email: {\tt Jon.M.Hjelmervik@sintef.no}}}
 
@@
 		\end{tabular}
-		\captionof{figure}{\label{fig:FRView} The oil reservoir viewer YYYY~\cite{cloudviz},
-  (name withheld for the double-blind review process) showing one
+		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
   server-rendered and two client-rendered images of automatically generated and
@@
 requirements on computational accuracy, power efficiency, data sizes, scaling
-properties, etc. This is also reinforced by new cloud-based approaches and
+properties, etc. This is also reinforced by popular cloud-based approaches and
 wireless usage patterns. An effect of this is that interactivity still is a
@@
 effective server time available for each one. Suitable scaling may still be
-achieved using our solution. This is currently being commercialized in an
-ongoing XXXX (name withheld for the double-blind review process) project.
+achieved using our solution. This is currently being commercialized as a part of
+the result of  a recently finished EU-project called CloudFlow.
 
@@
 %
-\textbf{[OLD: We are not aware of anybody else using depth maps exactly for our purpose, most
-research seems to be focused on the retrieval of the depth map from images,
-while we use an existing depth map to distill and render temporary geometry.]}
-\textbf{[NEW: Our approach differs from many similar ones, in that we use existing depth maps 
-to distill and render temporary geometry, rather than retrieveing the depth maps from images.]}
+Our approach differs from many similar ones, in that we use existing depth maps 
+to distill and render temporary geometry, rather than retrieveing the depth maps from images.
 %
@@
 
-We have tested the algorithms discussed in the ZZZZ (name withheld for the
-double-blind review process) framework, which is a programming framework for
+We have tested the algorithms proposed algorithms through the Tinia 
+framework~\cite{tinia}, which is a programming framework for
 setting up and managing client/server based interactive visualization
-applications, see~\cite{tinia}. As client we use Google Chrome, code is written
-in standard Javascript/WebGL.  The implementation in ZZZZ is invisible to the
-application, so all existing ZZZZ-applications will have the feature
+applications. As client we use Google Chrome, code is written
+in standard Javascript/WebGL.  The auto-proxy implementation in is invisible to the
+application, so all existing Tinia-applications will have the feature
 available. The algorithm is minimally intrusive in that only the depth buffer
 will have to be added to the rendering output of the application. We have tested
-several smaller test cases, but also on a larger oil reservoir viewer,
-YYYY~\cite{cloudviz}.
+several smaller test cases, but also on a larger oil reservoir viewer.
 
-YYYY utilizes several GLSL shaders to render reservoir cells, boundaries,
+The reservoir viewer utilizes several GLSL shaders to render reservoir cells, boundaries,
 tubular wells etc., and visualizes a 3D model that is not trivial to reduce in

### 2017-10-18 tex/full_2017/paper2.tex
Attempt at accommodating reviewers and their comments/suggestions.

@@
 
-\newcommand{\ie}{{\em i.e.}}
-\newcommand{\eg}{{\em e.g.}}
-\newcommand{\etal}{{\em et~al.}}
+%\newcommand{\ie}{{\em i.e.}}
+%\newcommand{\eg}{{\em e.g.}}
+%\newcommand{\etal}{{\em et~al.}}
+\newcommand{\ie}{{i.e.}}
+\newcommand{\eg}{{e.g.}}
+\newcommand{\etal}{{et~al.}}
 
@@
 
-\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
+%\author{\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\[0.3ex]\normalsize}
+%\IEEEauthorblockA{SINTEF Digital\\
+%Applied Mathematics\\
+%Norway\\
+%Email: {\tt Jens.O.Nygaard@sintef.no}}
+%\and
+%\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
+%\IEEEauthorblockA{SINTEF Digital\\
+%Applied Mathematics\\
+%Norway\\
+%Email: {\tt Jon.M.Hjelmervik@sintef.no}}}
+
+\author{
+\IEEEauthorblockN{~\\[-0.4ex]\large Jens Olav Nygaard\\ and Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
 \IEEEauthorblockA{SINTEF Digital\\
@@
 Norway\\
-Email: {\tt Jens.O.Nygaard@sintef.no}}
-\and
-\IEEEauthorblockN{~\\[-0.4ex]\large Jon Mikkelsen Hjelmervik\\[0.3ex]\normalsize}
-\IEEEauthorblockA{SINTEF Digital\\
-Applied Mathematics\\
-Norway\\
-Email: {\tt Jon.M.Hjelmervik@sintef.no}}}
+Email: {\tt Jens.O.Nygaard@sintef.no},\\ and {\tt Jon.M.Hjelmervik@sintef.no}}
+}
+
+%\author{
+%\IEEEauthorblockN{~\\[-0.4ex]\large Michael Shell\\[0.3ex]\normalsize}
+%\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
+%Georgia Institute of Technology\\
+%Atlanta, Georgia 30332--0250\\
+%Email: {\tt ...}}
+%\and
+%\IEEEauthorblockN{~\\[-0.4ex]\large Homer Simpson\\[0.3ex]\normalsize}
+%\IEEEauthorblockA{Twentieth Century Fox\\
+%Springfield, USA\\
+%Email: {\tt homer@thesimpsons.com}}
+%\and
+%\IEEEauthorblockN{~\\[-0.4ex]\large James Kirk\\ and Montgomery Scott\\[0.3ex]\normalsize}
+%\IEEEauthorblockA{Starfleet Academy\\
+%San Francisco, California 96678-2391\\
+%Telephone: (800) 555--1212\\
+%Fax: (888) 555--1212}
+%}
 
@@
 \maketitle
-\begin{center}
-    \centering
-		\begin{tabular}[b]{ccc}
-    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
-    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
-    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
-		a) & b) & c)
-		\end{tabular}
-		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
-  server-rendered and two client-rendered images of automatically generated and
-  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
-  best possible image, and in~(c) we want to highlight artifacts and
-  implementational details. See the text for further discussion.}
-%    \protect \label{fig:wfedesign} }
-%    \captionof{figure}{Test caption}
-\end{center}%
+%\begin{center}
+%    \centering
+%		\begin{tabular}[b]{ccc}
+%    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
+%    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
+%    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
+%		a) & b) & c)
+%		\end{tabular}
+%		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
+%  server-rendered and two client-rendered images of automatically generated and
+%  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
+%  best possible image, and in~(c) we want to highlight artifacts and
+%  implementational details. See the text for further discussion.}
+%%    \protect \label{fig:wfedesign} }
+%%    \captionof{figure}{Test caption}
+%\end{center}%
 }]
@@
 %\boldmath
-We describe an algorithm and implementation of automatic proxy model generation
-for a client/server remote rendering setup. A {\em proxy model} is a
-light-weight version of a main 3D model that is cheaper to render and
-transfer. We automatically derive this from the main model. The client is
-typically a tablet, for which we only assume availability of WebGL and
-Javascript. The server makes use of OpenGL and a web server. When the rate of
-received server-rendered images deteriorate, the client renders the proxy model,
-which is computed from depth buffers bundled with rendered images from the
-server. This algorithm requires little modification of the application itself.
+%We describe an algorithm and implementation of automatic proxy model generation
+%for a client/server remote rendering setup. A {\em proxy model} is a
+%light-weight version of a main 3D model that is cheaper to render and
+%transfer. We automatically derive this from the main model. The client is
+%typically a tablet, for which we only assume availability of WebGL and
+%Javascript. The server makes use of OpenGL and a web server. When the rate of
+%received server-rendered images deteriorate, the client renders the proxy model,
+%which is computed from depth buffers bundled with rendered images from the
+%server. This algorithm requires little modification of the application itself.
+%
+%Revisit the Abstract to introduce the topic, as for a regular reader; currently the abstract starts too abruptly into the matter.  I suggest:
+%a) What is the problem in a given domain?
+%b) Why is the problem interesting? (a few sentences, shortly)
+%c) What other did/failed on, and what you are dealing with
+%d) What is your solution? (one sentence, very short)
+%e) End the abstract with one sentence on what is the main output of the paper
+%
+%The algorithm is very well described and the computation-flow formulas for generating the proxy model are accurately described.
+%
+% New version after comments from reviewers
+%
+A common problem in remote rendering setups, is that of temporarily insufficient
+bandwidth and latency. For a proper experience of immersiveness, at least some
+rendering should be presented to the user, and it should appear to be responsive
+to user input, even in the presence of connection glitches. The all too familiar
+spinning hourglass symbol, or equivalent, will degrade such an experience. With
+the advent of VR and AR, solutions to this become important even though
+connectedness in general is improving. We are dealing with a remote rendering of
+3D geometry being pushed to a lesser client, and what we in essence do is to
+replace a spinning hourglass symbol with an automatically client-generated
+approximation of the 3D geometry rendered on the client, responding to
+client-recorded user input. We call this a {\em proxy model}.  Our main result
+is a system for automatically producing such proxy models on the client, from
+received images and depth buffers only, for showing on the client when remotely
+rendered frames do not arrive sufficiently fast.
 \end{abstract}
@@
 
+\begin{figure*}
+\begin{center}
+    \centering
+		\begin{tabular}[b]{ccc}
+    \includegraphics[width=0.32\linewidth, trim={1cm 1cm 1cm 1cm}, clip]{frviewb1.png} &
+    \includegraphics[width=0.32\linewidth, trim={0 1cm 0 1.5cm}, clip]{frviewb2.png} &
+    \includegraphics[width=0.32\linewidth, trim={0 1.5cm 0 2cm}, clip]{frviewb3.png} \\
+		a) & b) & c)
+		\end{tabular}
+		\captionof{figure}{\label{fig:FRView} Oil reservoir viewer showing one
+  server-rendered and two client-rendered images of automatically generated and
+  slightly rotated proxy models.  For~(b), parameters are chosen to produce the
+  best possible image, and in~(c) we want to highlight artifacts and
+  implementational details. See the text for further discussion.}
+\end{center}%
+\end{figure*}
+
 Hand in hand with increasingly powerful rendering engines comes ever increasing
@@
 latency, bandwidth and scaling problems in a novel way.
-
+\\
 We introduce a {\em proxy model}, defined as a temporary model to be shown and
@@
 connected server. Producing such proxy models can be difficult for many kinds of
-3D data, like in our main case in which we have an oil reservoir viewer
-(YYYY,~\cite{cloudviz}) that renders large {\em corner point grids}, together
-with faults, oil wells, and more. See Figure~\ref{fig:FRView} above for an
+3D data, like in our main case in which we have an oil reservoir
+viewer~\cite{cloudviz} that renders large {\em corner point grids}, together
+with faults, oil wells, and more; see Figure~\ref{fig:FRView} for an
 example of both a full server-side rendering, and automatically generated proxy
@@
 achieved using our solution. This is currently being commercialized as a part of
-the result of  a recently finished EU-project called CloudFlow.
-
-%-------------------------------------------------------------------------
-\subsection{Previous work}
+the result of  a recently finished EU-project called CloudFlow~\cite{cloudflow}.
+\\
+In Section~\ref{sec:prevWork}, we review some previous and related work, before
+we consider our contribution in Section~\ref{sec:autoproxy}. This is further
+split into three subsections, in Subsection~\ref{subsec:server} we consider the
+server side part of the system, in Subsection~\ref{sec:client}, the client side
+and in Subsection~\ref{sec:autoTuning} we consider automatic parameter tuning
+during use of the system. We discuss results in Section~\ref{sec:results},
+before we finally sum up in Subsection~\ref{sec:conc}.
+
+%-------------------------------------------------------------------------
+\section{Previous work}
 \label{sec:prevWork}
 
-Our approach has some similarities to {\em image based rendering} (IBR)
-techniques, with the difference that an important IBR problem would be the
-reconstruction of a depth map from images, while we have access to the full
-depth map from a rendering pass.  Another way to use depth maps similar to what
-we do, is for ``immersive streaming'', see \eg,~\cite{ibr}, where focus is on
-depth map compression, an issue that we also consider. Another work
-focused on similar streaming and geometry compression,
-is~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
+Our approach has some similarities to {\em Image Based Rendering} (IBR)
+techniques~\cite{journalsritaOliveira02}, with the difference that an
+important IBR problem would be the reconstruction of a depth map from images,
+while we have access to the full depth map from a rendering pass.  Another way
+to use depth maps similar to what we do, is for ``immersive streaming'', see,
+\eg,~\cite{ibr}, where focus is on depth map compression, an issue that we also
+consider. Another work focused on similar streaming and geometry compression,
+is Teler~\cite{teler}. Also,~\cite{220764} contains some of the same ideas as our
 work, with respect to image-based rendering acceleration.
 %
-\textbf{A very early work aiming at the same kind of ``inter-frame rendering'', is
-\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped, and
-subsequently combined, in order to avoid large unpainted areas caused by
+A very early work aiming at the same kind of ``inter-frame rendering'', is
+Mark \etal~\cite{Mark:1997:PW:253284.253292}. Here, several frames are warped,
+and subsequently combined, in order to avoid large unpainted areas caused by
 occlusion. This corresponds to our use of several proxy models, each from a pair
@@
 slightly complicated meshing, discontinuity detection and final image
-composition stage.}
+composition stage.
 %
@@
 %
-\textbf{Another approach is that of~\cite{altenhofen16rixels}. They use websockets
+Another approach is that of~\cite{altenhofen16rixels}. They use websockets
 where we use the http protocol, a more significant difference is that we get
@@
 %
-}
-%
-\textbf{In~\cite{CGF:CGF1871}, Pajak \etal considers remote rendering and
+In~\cite{CGF:CGF1871}, Pajak \etal~considers remote rendering and
 streaming of frames rendered from a dynamical 3D model, we are quite agnostic
 to the source of imagery. Their setup requires more powerful clients than ours
-  (OpenGL vs. WebGL), but they will also have higher fidelity.  }
-%
-\textbf{A special case is provided for in the rendering of stereo images. In
-  this case, the 3D disparity is limited, while the rendering cost is doubled,
-  since two views per frame is needed. In~\cite{DidykERMS2010}, this is
-  exploited to make a solution tailored to such stereo synthesis; performance
-  approaches that of rendering non-stereo, with a minimization of depth
-  disparity artifacts. Occlusion and disocclusion holes are avoided by warping quads rather than
-  pixel and filling in with previous images.}
+(OpenGL vs. WebGL), but they will also have higher fidelity.
+%
+A special case is provided for in the rendering of stereo images. In this case,
+the 3D disparity is limited, while the rendering cost is doubled, since two
+views per frame is needed. In~\cite{DidykERMS2010}, this is exploited to make a
+solution tailored to such stereo synthesis; performance approaches that of
+rendering non-stereo, with a minimization of depth disparity
+artifacts. Occlusion and disocclusion holes are avoided by warping quads rather
+than pixel and filling in with previous images.
 %
@@
 \section{The auto proxy algorithm}
+\label{sec:autoproxy}
 
@@
   \caption{\label{fig:2DheightmapRotated}
-Bird's view of 3D model (solid lines) and proxy model renderings.
+Bird's view of 3D model (solid lines) and proxy\\model renderings.
+}
+\end{figure}
+
 %           The depth map and image sent from the server enables the client to
@@
 %           seen from above.
-}
-\end{figure}
 
@@
 \subsection{The server side}
+\label{subsec:server}
 
 The server renders the 3D model into a framebuffer, and in the process generates
-a depth image that we also send to the client.
-Since this adds to the data being sent, it must be kept to a minimum. We have
-found that reducing the spatial resolution of the depth image (for instance by a
-factor of $1/16$) only degrades the proxy model imperceptibly.  We also encode
-each depth value in the range $[0, 1]$ as a 16 bit fixed point number, and the
-bundling of the depth image with the RGB image then imposes just a small data
-overhead.
-Further compression may bring this down even more, but the cost of
-compression/decompression must also be
-considered.
-One proposed solution is to be found in~\cite{DBLP:journals/tvcg/Lindstrom14},
-which promises to be fast both for the compression and decompression stages.
+a depth image that we also send to the client.  Since this adds to the data
+being sent, it must be kept to a minimum. We have found that reducing the
+spatial resolution of the depth image (for instance by a factor of $1/16$) only
+degrades the proxy model imperceptibly.  We also encode each depth value in the
+range $[0, 1]$ as a 16 bit fixed point number, and the bundling of the depth
+image with the RGB image then imposes just a small data overhead.  Further
+compression may bring this down even more, but the cost of
+compression/decompression must also be considered.  One proposed solution is to
+be found in~\cite{DBLP:journals/tvcg/Lindstrom14}, which promises to be fast
+both for the compression and decompression stages.
 
@@
 the client already has in store, see Section~\ref{sec:proxyModelReplacement}.
-
+\\
 As indicated by Figure~\ref{fig:2DheightmapRotated}~(b), the received height map
@@
 Figure~\ref{fig:2DheightmapRotated}~(b).
-
+\\
 Rendering a 3D point for each depth fragment available may tax a thin client,
@@
 reconstruct topology, but we risk getting more ``holes'' in our models.
-
+\\
 We can render each splat as a fixed geometry, \eg, a disk or rectangle, with the
@@
 approximation, so we should not make them too large either.
-
+\\
 This leaves us $\pv$, $\pv_{x+\Delta x}$ and $\pv_{y+\Delta y}$ in clip
@@
 that the geometry is locally planar does not hold, \eg, if the splats are very
-large, or they originate from a curved or non-smooth part, this may look rather
-odd, see Figure~\ref{fig:LargeSplatsOnCorners}.
+large, or they originate from a curved or non-smooth part, this may look
+slightly odd, see Figure~\ref{fig:LargeSplatsOnCorners}. The figure shows splats
+for which depths are sampled on different planar regions at an angle to one
+another, causing contortions when large splats cover more than one such planar
+region.
 
@@
   \caption{\label{fig:LargeSplatsOnCorners}
-           Splats for which depths are sampled on a planar region look nice on
-  this region, but possibly contorted elsewhere. Notice the corner and edges
-  here, where we have chosen parameters to display this effect.}
+% Reviewer said ``too much text''...
+%  Splats for which depths are sampled on a planar region look nice on
+%  this region, but possibly contorted elsewhere. Notice the corner and edges
+%  here, where we have chosen parameters to display this effect.}
+  Notice the corner and edges, parameters are chosen to display effects of
+  different planes meeting at edges.}
 \end{figure}
@@
 removes the problem visualized in Figure~\ref{fig:LargeSplatsOnCorners}.
-
+\\
 \textbf{Splat depth fragments}
-For larger splats it makes sense to also compute and use depth fragments on the
+For larger splats, it makes sense to also compute and use depth fragments on the
 client. The ``intra-splat'' depth values can easily be fetched from the depth
@@
 WebGL extension.
-
+\\
 \textbf{Splat set replacement algorithms}
@@
 to cover larger ranges of client-side transformations. One can imagine a plethora of
-{\em splat set replacement algorithms}, we have tested three approaches that all
+{\em splat set replacement algorithms.} We have tested three approaches that all
 retain a constant number of proxy models. The first is to replace the
@@
 more than necessary. These demands are not always compatible.
-
+\\
 We have adopted an adaptive specification of proxy model data from the server
@@
 \section{Results and discussion}
-
-We have tested the algorithms proposed algorithms through the Tinia 
+\label{sec:results}
+
+We have tested the proposed algorithms through the Tinia 
 framework~\cite{tinia}, which is a programming framework for
 setting up and managing client/server based interactive visualization
-applications. As client we use Google Chrome, code is written
-in standard Javascript/WebGL.  The auto-proxy implementation in is invisible to the
+applications. As client, we use Google Chrome, code is written
+in standard Javascript/WebGL.  The auto-proxy implementation is invisible to the
 application, so all existing Tinia-applications will have the feature
@@
 several smaller test cases, but also on a larger oil reservoir viewer.
-
-The reservoir viewer utilizes several GLSL shaders to render reservoir cells, boundaries,
-tubular wells etc., and visualizes a 3D model that is not trivial to reduce in
-complexity. It is typically also very large, so rendering it on a thin client is
-prohibitive. With the automatic proxy model, we obtain interactive frame rates
-with limited connection from a lightweight client. For a comparison of a
-server-rendered image and a client-rendered proxy model that is slightly rotated
-on the client, see again Figure~\ref{fig:FRView}. In
-Figure~\ref{fig:FRView}~(a), the full server-rendered image is shown, and in~(b)
-and~(c), a slightly (about 10 degrees) rotated proxy model is rendered on the
-client. In~(b), parameters are chosen for best possible results, while in~(c),
-we want to highlight effects, so it uses a small number of non-overlapping large
-splats. One can easily spot areas not well covered, and areas where the texture
-coordinate transform $T$ does not produce optimal results.
+\\
+The reservoir viewer utilizes several OpenGL Shading Language (GLSL) shaders to
+render reservoir cells, boundaries, tubular wells, etc., and visualizes a 3D
+model that is not trivial to reduce in complexity. It is typically also very
+large, so rendering it on a thin client is prohibitive. With the automatic proxy
+model, we obtain interactive frame rates with limited connection from a
+lightweight client. For a comparison of a server-rendered image and a
+client-rendered proxy model that is slightly rotated on the client, see again
+Figure~\ref{fig:FRView}. In Figure~\ref{fig:FRView}~(a), the full
+server-rendered image is shown, and in~(b) and~(c), a slightly (about 10
+degrees) rotated proxy model is rendered on the client. In~(b), parameters are
+chosen for best possible results, while in~(c), we want to highlight effects, so
+it uses a small number of non-overlapping large splats. One can easily spot
+areas not well covered, and areas where the texture coordinate transform $T$
+does not produce optimal results.
 
@@
 \section{Conclusion and future work}
+\label{sec:conc}
 
@@
 \item \textbf{Depth compression} We would like to ivestigate other approaches
-than just truncation, see \eg,~\cite{DBLP:journals/tvcg/Lindstrom14}.
+than just truncation, see, \eg,~\cite{DBLP:journals/tvcg/Lindstrom14}.
 

### 2017-11-07 tex/full_2017/paper2.tex
Initial fixing as requested by Sue

@@
 %
-A common problem in remote rendering setups, is that of temporarily insufficient
+A common problem in remote rendering setups is that of temporarily insufficient
 bandwidth and latency. For a proper experience of immersiveness, at least some
@@
 spinning hourglass symbol, or equivalent, will degrade such an experience. With
-the advent of VR and AR, solutions to this become important even though
-connectedness in general is improving. We are dealing with a remote rendering of
-3D geometry being pushed to a lesser client, and what we in essence do is to
-replace a spinning hourglass symbol with an automatically client-generated
-approximation of the 3D geometry rendered on the client, responding to
-client-recorded user input. We call this a {\em proxy model}.  Our main result
-is a system for automatically producing such proxy models on the client, from
-received images and depth buffers only, for showing on the client when remotely
-rendered frames do not arrive sufficiently fast.
+the advent of Virtual Reality (VR) and Augmented Reality (AR), solutions to this
+become important even though connectedness in general is improving. We are
+dealing with a remote rendering of three-dimensional (3D) geometry being pushed
+to a lesser client, and what we in essence do is to replace a spinning hourglass
+symbol with an automatically client-generated approximation of the 3D geometry
+rendered on the client, responding to client-recorded user input. We call this a
+{\em proxy model}.  Our main result is a system for automatically producing such
+proxy models on the client, from received images and depth buffers only, for
+showing on the client when remotely rendered frames do not arrive sufficiently
+fast.
 \end{abstract}
@@
 \IEEEpeerreviewmaketitle
+
+\parskip 0pt
 
@@
 latency, bandwidth and scaling problems in a novel way.
-\\
+
 We introduce a {\em proxy model}, defined as a temporary model to be shown and
@@
 Our solution is to pass depth information from the server along with ordinary
-rendered images. From this, a rudimentary 3D model is built, and with the RGB
-image as a texture, this model can be manipulated and rendered on the client
-while waiting for the next update from the server. If the client does not change
-the position or orientation of the model too much, this proxy model rendering
-integrates seamlessly with the slower stream of server-rendered frames.
+rendered images. From this, a rudimentary 3D model is built, and with the Red,
+Green and Blue (RGB) image as a texture, this model can be manipulated and
+rendered on the client while waiting for the next update from the server. If the
+client does not change the position or orientation of the model too much, this
+proxy model rendering integrates seamlessly with the slower stream of
+server-rendered frames.
 %
@@
 achieved using our solution. This is currently being commercialized as a part of
-the result of  a recently finished EU-project called CloudFlow~\cite{cloudflow}.
+the result of a recently finished European Union (EU) project called
+CloudFlow~\cite{cloudflow}.
 \\
@@
 %
-In~\cite{CGF:CGF1871}, Pajak \etal~considers remote rendering and
-streaming of frames rendered from a dynamical 3D model, we are quite agnostic
-to the source of imagery. Their setup requires more powerful clients than ours
-(OpenGL vs. WebGL), but they will also have higher fidelity.
+In~\cite{CGF:CGF1871}, Pajak \etal~considers remote rendering and streaming of
+frames rendered from a dynamical 3D model, we are quite agnostic to the source
+of imagery. Their setup requires more powerful clients than ours, due to the use
+of the Open Graphics Library (OpenGL) vs.~Web Graphics Library (WebGL), but they
+will also have higher fidelity.
 %
@@
 We have adopted an adaptive specification of proxy model data from the server
-involving a more light-weight image (lossy JPG-compression with adaptive quality
-control) while interaction is ongoing.  Also, reduced resolution of the depth
-buffer sent from the server is used. Another possibility is to let the client
-dynamically set the number of splats, number of proxy models, etc.
+involving a more light-weight image (lossy Joint Photographic Experts Group
+(JPEG) compression with adaptive quality control) while interaction is ongoing.
+Also, reduced resolution of the depth buffer sent from the server is
+used. Another possibility is to let the client dynamically set the number of
+splats, number of proxy models, etc.
 
